---
title: "Spatial Analysis of Obesity"
subtitle: "Interactive Plots"
author: "Colin, Lucas, Nina, Qasim"
date: "11 April 2025"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Spatial Distribution of Obesity in Singapore

```{r}
start.time = Sys.time()
pkgs = c("dplyr", "ggplot2", "tidyr", "readr", "crayon", "tibble", "knitr", "gt", "broom", "tidyverse", "sf", "tmap", "aspace", "spdep", "corrplot", "readxl", "spatialreg")
invisible(lapply(pkgs, function(pkg) {
  suppressPackageStartupMessages(library(pkg, character.only = TRUE))
}))
rm(pkgs)
if (requireNamespace("rstudioapi", quietly = TRUE)) {
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}
```

```{r}
# saveRDS(sg_crs, "InteractivePlots/sg_crs.rds")
obesity_data_sf_2023_OSO <-
  readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/obesity_data_sf_2023_OSO.rds")
sg_crs <-
  readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/sg_crs.rds")
```

```{r}
### by Primary School

tmap_mode("view")

# filter out by educational level and remove NA values to calculate SDE
obesity_pri_sch <- obesity_data_sf_2023_OSO %>% 
  filter(LEVEL == "PRIMARY", !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
pri_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_pri_sch))
pri_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_pri_sch))
pri_sch_mean_median <- rbind(pri_sch_mean$LOCATIONS, pri_sch_median$LOCATIONS) %>%
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
pri_sch_mean_median$id <- c("Spatial mean", "Spatial median")

pri_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_pri_sch))
pri_sch_sdd_line <- pri_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_pri_sch))
# pri_sch_sde$ATTRIBUTES[,10] # theta = 92.55625
pri_sch_sde_line <- pri_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_pri_sch$PREVALENCE,
  points = st_coordinates(obesity_pri_sch))
pri_sch_sde_weighted_line <- pri_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_map <- 
  tm_shape(obesity_pri_sch, name = "Overweight prevalence") +
  tm_symbols(
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30), 
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence", 
      position = c("TOP", "LEFT"), 
      item.height = 0.5, 
      item.width = 2.1
      )
    )

pri_sch_map_summ <- pri_sch_map +
  # tm_shape(pri_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) +
  tm_shape(pri_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(pri_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(pri_sch_mean_median, name = "Summary measures") + 
  tm_dots(size = 0.5, shape = "id", fill = "white", shape.legend = tm_legend("Summary measures")) + 
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  # tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_title("Overweight prevalence in Primary Schools")

### by Secondary School

# filter out by educational level and remove NA values to calculate SDE
obesity_sec_sch <- obesity_data_sf_2023_OSO %>% 
  filter(LEVEL == "SECONDARY", !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
sec_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_sec_sch))
sec_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_sec_sch))
sec_sch_mean_median <- rbind(sec_sch_mean$LOCATIONS, sec_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
sec_sch_mean_median$id = c("Spatial mean", "Spatial median")

sec_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_sec_sch))
sec_sch_sdd_line <- sec_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_sec_sch))
# sec_sch_sde$ATTRIBUTES[,10] # theta = 92.3718
sec_sch_sde_line <- sec_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_sec_sch$PREVALENCE,
  points = st_coordinates(obesity_sec_sch))
sec_sch_sde_weighted_line <- sec_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_map <- 
  tm_shape(obesity_sec_sch, name = "Overweight prevalence") +
  tm_symbols(
    shape = 22,
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      position = c("TOP", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    )

sec_sch_map_summ <- sec_sch_map +
  # tm_shape(sec_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) +
  tm_shape(sec_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(sec_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(sec_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  # tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_title("Overweight prevalence in Secondary Schools")

tmap_arrange(pri_sch_map_summ, sec_sch_map_summ, nrow = 2)
```

The standard deviational ellipse, when accounting for school obesity
(red band), is being "pulled" towards the North. (spatial clustering of
high obesity schools)

# Outside-of-School Food Options

```{r}
hawker <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/hawker.rds")
hawker_full <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/hawker_full.rds")
macs <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/macs.rds")
macs_full <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/macs_full.rds")
seven_eleven_full <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/seven_eleven_full.rds")
seven_eleven <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/seven_eleven.rds")
obesity_pri_sch <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/obesity_pri_sch.rds")
ura_regions_map <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/ura_regions_map.rds")
obesity_sec_sch <- readRDS("~/Desktop/RStudio/Spatial Data DSA303 codes/dsa303/InteractivePlots/obesity_sec_sch.rds")
```

## No. of Food Outlets near Schools

```{r}
tmap_mode("view")

hawker = hawker_full["geometry"]
macs = macs_full[,"geometry"] 
seven_eleven = seven_eleven_full[,"geometry"]

#Combine data into single SF object
seven_eleven_sf = mutate(seven_eleven, outlet = "7-11")
macs_sf = mutate(macs, outlet = "mcdonalds")
hawker_sf = mutate(hawker, outlet = "hawker")
food = rbind(seven_eleven_sf, macs_sf, hawker_sf)

#Primary school analysis
obesity_pri_food = obesity_pri_sch 
pri_schools_buffer1 <- st_buffer(obesity_pri_food, dist = 1000) %>% st_geometry()
pri_intersections <- st_intersects(pri_schools_buffer1, food)

#Calculate the number of food outlets within a 1km radius of each school
obesity_pri_food$num_food_outlets <- lengths(pri_intersections)

#Show density via choropleth map
pri_food_map = ura_regions_map + 
  tm_shape(obesity_pri_food, name = "No. of food outlets") +
  tm_symbols(
    size = 0.5,
    fill = "num_food_outlets", fill.scale = tm_scale(values = "brewer.yl_or_rd",
                                                     breaks = c(0,5,10,15,20,25,30,35)),
    fill.legend = tm_legend(title = "No. of food outlets", position = c("TOP", "LEFT"))
  ) +
  tm_title("Food Outlet Density by Primary School")


#Secondary school analysis
obesity_sec_food = obesity_sec_sch 
sec_schools_buffer1 <- st_buffer(obesity_sec_food, dist = 1000) %>% st_geometry()
sec_intersections <- st_intersects(sec_schools_buffer1, food)

#Calculate the number of food outlets within a 1km radius of each school
obesity_sec_food$num_food_outlets <- lengths(sec_intersections)

#Show density via choropleth map
sec_food_map = ura_regions_map + 
  tm_shape(obesity_sec_food, name = "No. of food outlets") +
  tm_symbols(
    size = 0.5, 
    fill = "num_food_outlets", fill.scale = tm_scale(values = "brewer.yl_or_rd",
                                                     breaks = c(0,5,10,15,20,25,30,35)),
    fill.legend = tm_legend(title = "No. of food outlets", position = c("TOP", "LEFT"))
  ) +
  tm_title("Food Outlet Density by Secondary School")

tmap_arrange(pri_food_map, sec_food_map, nrow = 2)
```

The maps above show food outlet density at the school level, with darker
colours representing more food options.

# Changes in Obesity over Time

```{r}
oso_pri_hist <- readRDS("oso_pri_hist.rds")
obesity_data_sf_2023_OSO <- readRDS("obesity_data_sf_2023_OSO.rds")
ura_regions_map <- readRDS("ura_regions_map.rds")
oso_hist_geom <- readRDS("oso_hist_geom.rds")
pri_sch_map <- readRDS("pri_sch_map.rds")
obesity_pri_sch <- readRDS("obesity_pri_sch.rds")
```

```{r}
###Primary School Z-Score Change Map for Model 1
library(tidyverse)
library(imputeTS)  # For interpolation
library(fpp2)
library(gridExtra)
library(seasonal)
library(urca)
library(vars)
library(trend)
library(readxl)
library(dplyr)

oso_pri_hist_years = oso_pri_hist[, c(1,2, 18, 20, 22, 24, 26, 28)] %>%
  drop_na()
  

oso_pri_hist_long <- oso_pri_hist_years %>%
  pivot_longer(
    cols = matches("^\\d"),   # Select columns where names start with a number
    names_to = "year",
    values_to = "oso"
  ) %>%
  mutate(year = as.integer(paste0("20", str_extract(year, "^\\d{2}"))))  


oso_hist_int = oso_pri_hist_long %>%
  group_by(SCHOOL) %>%
  complete(year = full_seq(year, 1)) %>%  # Adds missing years with NA values
  mutate(oso = na_interpolation(oso, option = "spline"),
         oso_int = na_interpolation(oso, option = "spline")) %>%
  ungroup() 

# Showing general distribution of OSO in a histogram
# oso_hist_int %>% 
#   filter(year == 2023) %>% 
#   ggplot(aes(x = oso_int)) +
#   geom_histogram() # generally follows a normal distribution

#Get the Z-scores
# Compute z-scores before and after 2019
oso_hist_int = oso_hist_int %>%
  group_by(year) %>%  # Group by year
  mutate(
    mean_oso = mean(oso_int, na.rm = TRUE),
    sd_oso = sd(oso_int, na.rm = TRUE),
    z_score = (oso_int - mean_oso) / sd_oso
  ) %>%
  ungroup()

# Label the periods
oso_hist_int_2 <- oso_hist_int %>%
  mutate(period = ifelse(year <= 2019, "Before 2019", "After 2019"))

z_scores_summary <- oso_hist_int_2 %>%
  group_by(SCHOOL, period) %>%
  summarise(avg_z = mean(z_score, na.rm = TRUE)) %>%
  ungroup()

z_scores_change= z_scores_summary %>%
  pivot_wider(names_from = period, values_from = avg_z) %>%
  mutate(z_change =`After 2019` - `Before 2019`) %>%
  dplyr::select(c(1,4))


# Merge with spatial data
oso_hist_geom_full = obesity_data_sf_2023_OSO %>%
  inner_join(z_scores_change, by = "SCHOOL")

tmap_mode("view")

oso_hist_geom = oso_hist_geom_full %>%
  dplyr::select(c("PT_ID", "SCHOOL", "geometry", "z_change"))


# School obesity map plot (z-score change)
ura_regions_map +
  tm_shape(oso_hist_geom, name = "Overweight prevalence") +
  tm_symbols(size = 0.5, fill = "z_change", fill.scale = tm_scale(values = "brewer.yl_or_rd")) +
  tm_title("Graph of Z-score Changes in Obesity Prevalence for Primary Schools")
```

The plot contains the Z-score changes for each Primary School, with
darker red colours representing worsening obesity levels and lighter
colours representing improving obesity levels.

## Obesity in Schools: Improving or Worsening?

```{r}
# by Primary schools
# obesity_pri_sch_jitter <- st_jitter(obesity_pri_sch) 
del_nb_pri_z <- tri2nb(st_coordinates(oso_hist_geom), 
                     row.names = oso_hist_geom$PT_ID) 
pri_map <- tm_shape(nb2lines(del_nb_pri_z, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Primary schools
if (require(dbscan, quietly=TRUE)) {
  soi_nb_pri_z <- graph2nb(soi.graph(del_nb_pri_z, st_geometry(oso_hist_geom)))
} else {
  soi_nb = NULL
}
pri_map_z <- tm_shape(nb2lines(soi_nb_pri_z, coords = st_geometry(oso_hist_geom))) +
  tm_lines(lwd = 2) + pri_sch_map


# by Primary schools
soi_lw_pri_z <- nb2listw(soi_nb_pri_z, style = "W")
soi_lw_B_pri_z <- nb2listw(soi_nb_pri_z, style = "B")

# Setting an overweight cutoff point
Z_CHANGE_CUTOFF_INCREASE <- 1 #increase in obesity levels before and after 2019
Z_CHANGE_CUTOFF_DECREASE <- -1 #decrease in obesity levels before and after 2019

# by Primary school - with increased obesity levels
oso_hist_geom1 = oso_hist_geom %>%
  filter(z_change > 0)
oso_hist_geom1$z_change_high_increase <- 
  as.factor(oso_hist_geom1$z_change > Z_CHANGE_CUTOFF_INCREASE)
pri_map_z_increase <- ura_regions_map + oso_hist_geom1 %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "z_change_high_increase", 
    fill.legend = tm_legend(title = "High increase in obesity", position = c("TOP", "LEFT")),
    size = "z_change", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Primary Schools with Worsening Obesity Levels")

# by Primary school - Decrease in Obesity Levels
oso_hist_geom3 = oso_hist_geom %>%
  filter(z_change < 0) %>%
  mutate(z_change = abs(z_change))

Z_CHANGE_CUTOFF_DECREASE_1 = 1
oso_hist_geom3$z_change_high_decrease <- 
  as.factor(oso_hist_geom3$z_change > Z_CHANGE_CUTOFF_DECREASE_1)
pri_map_z_decrease1 <- ura_regions_map + oso_hist_geom3 %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "z_change_high_decrease", 
    fill.legend = tm_legend(title = "High decrease in obesity", position = c("TOP", "LEFT")),
    size = "z_change", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Primary Schools with Improving Obesity Levels")


#Maps:
tmap_arrange(pri_map_z_decrease1, pri_map_z_increase, nrow = 2)
```

The plots provide a descriptive view of Primary Schools with improving
and worsening obesity levels over time. The blue circles represent
extreme cases.

## Change in Obesity: Before vs. After COVID

```{r}
#Model 2: Contingency Table of Primary Schools with Worsening Obesity Levels, controlled for COVID

oso_pri_hist_years = oso_pri_hist[, c(1,2, 18, 20, 22, 24, 26, 28)] %>%
  drop_na()
  
oso_pri_hist_long <- oso_pri_hist_years %>%
  pivot_longer(
    cols = matches("^\\d"),   # Select columns where names start with a number
    names_to = "year",
    values_to = "oso"
  ) %>%
  mutate(year = as.integer(paste0("20", str_extract(year, "^\\d{2}"))))  


oso_hist_int_1 = oso_pri_hist_long  


#Get the Z-scores
# Compute z-scores before and after 2019
oso_hist_int_1 = oso_hist_int_1 %>%
  group_by(year) %>%  # Group by year
  mutate(
    mean_oso = mean(oso, na.rm = TRUE),
    sd_oso = sd(oso, na.rm = TRUE),
    z_score = (oso - mean_oso) / sd_oso
  ) %>%
  ungroup()


oso_covid = oso_hist_int_1 %>%
  mutate(covid_period = case_when(
    year < 2022 ~ "BEFORE_COVID",
    year == 2022 ~ "DURING_COVID",
    year == 2023 ~ "AFTER_COVID",
    TRUE ~ NA_character_
  )) %>%
  group_by(SCHOOL, covid_period) %>%
  summarise(avg_z_score = mean(z_score, na.rm = TRUE), .groups = "drop_last") %>%
  pivot_wider(names_from = covid_period, values_from = avg_z_score) %>%
  mutate(
    after_covid_change = `AFTER_COVID` - `DURING_COVID`,
    before_covid_change = `DURING_COVID` - `BEFORE_COVID`
  )

# Merge with spatial data
oso_covid_geom_full = obesity_data_sf_2023_OSO %>%
  left_join(oso_covid, by = "SCHOOL") %>%
  filter(LEVEL == "PRIMARY") %>%
  drop_na()


oso_covid_geom = oso_covid_geom_full %>%
  dplyr::select(c("PT_ID", "SCHOOL", "geometry", "after_covid_change", "before_covid_change"))


#Classify
oso_covid_geom_classes <- oso_covid_geom %>%
  mutate(
    before_covid_range = case_when(
      before_covid_change < quantile(before_covid_change, 0.33) ~ "Low",
      before_covid_change < quantile(before_covid_change, 0.66) ~ "Medium",
      TRUE ~ "High"
    ),
    after_covid_range = case_when(
      after_covid_change < quantile(after_covid_change, 0.33) ~ "Low",
      after_covid_change < quantile(after_covid_change, 0.66) ~ "Medium",
      TRUE ~ "High"
    )
  )

oso_covid_hh = oso_covid_geom_classes %>%
  filter(before_covid_range == "High" & after_covid_range == "High")

oso_covid_lh = oso_covid_geom_classes %>%
  filter(before_covid_range == "Low" & after_covid_range == "High")

#Scatterplot of schools

scatterplot_of_schools <- ggplot(oso_covid_geom_classes, aes(x = before_covid_change, y = after_covid_change, label = SCHOOL)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Change in Obesity: Before vs. After COVID",
    x = "Before COVID",
    y = "After COVID"
  ) +
  theme_minimal()

#Plot of Contingency Table for Primary School Obesity Rates Controlled for COVID
plotly::ggplotly(scatterplot_of_schools)
```

The plot above provides a contingency table for Primary School obesity
rates when controlling for COVID effects.
