# Data loading, cleaning and transformation

## Loading the data

Change the file paths accordingly

```{r}
# Loading libraries to work with spatial data
library(tidyverse); library(sf); library(tmap); library(aspace); library(spdep)

# Loading Singapore URA subzone boundary data
sg_ura_boundaries <- 
  st_read("MasterPlan2014SubzoneBoundaryWebSHP/MP14_SUBZONE_WEB_PL.shp") %>% 
  st_make_valid()

# Saving CRS provided by SG boundary map
sg_crs <- st_crs(sg_ura_boundaries)

# Loading cleaned obesity data from HPB and giving spatial awareness
obesity_data <- 
  # "#N/A" values converted to "NA" for R to be able to read it
  read.csv("obesity.csv", na.strings = "#N/A") %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  st_transform(crs = sg_crs) %>% 
  st_make_valid()

# Loading cleaned HMSP data from HPB and giving spatial awareness
hmsp_data_sf <- 
  read.csv("hmsp_tier_pri_sec.csv") %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  st_transform(crs = sg_crs) %>% 
  st_make_valid() 

# Checking that all CRS are matching
isTRUE(all.equal(st_crs(sg_ura_boundaries), st_crs(obesity_data), st_crs(hmsp_data_sf)))

# Checking CRS units
st_crs(sg_ura_boundaries)$units
st_crs(obesity_data)$units
st_crs(hmsp_data_sf)$units
```

## Transforming data to be usable for analysis

```{r}
obesity_data_sf <- obesity_data %>% 
  pivot_longer(
    # pivot_longer to separate obesity data columns into a tidy format
    cols = starts_with("X"),
    names_to = c("YEAR", "CATEGORY"),
    names_prefix = "X",
    names_sep = "_",
    names_transform = list(YEAR = as.integer),
    values_to = "COUNT"
    ) %>% 
  mutate(
    # Transform obesity count and school enrollment into an obesity prevalence figure
    PREVALENCE = COUNT / ENROLMENT
    )
```

# Exploratory data analysis (by Prof)

```{r}
tmap_mode("plot")
tm_shape(sg_ura_boundaries) + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.4, fill = "TIER", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "HMSP Tiers"))

tm_shape(sg_ura_boundaries) + tm_fill() + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.4, fill = "ZONE", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "School zones"))

tm_shape(sg_ura_boundaries) + tm_fill() + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.3, fill = "CLUSTER", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "School clusters")) +
  tm_facets_wrap(by = "ZONE")
```

# Obesity profile in Singapore (Colin)

## Singapore maps

```{r}
tmap_mode("view")

# URA regions map
ura_regions <- sg_ura_boundaries %>% group_by(REGION_N) %>% 
  summarise(geometry = st_union(geometry))
ura_region_map <- tm_shape(ura_regions, name = "URA regions") + tm_polygons(
    col = "grey", fill = "REGION_N", fill_alpha = 0.25, 
    fill.legend = tm_legend(title = "URA regions"))

# URA planning areas map
ura_planning_areas <- sg_ura_boundaries %>% group_by(PLN_AREA_N) %>% 
  summarise(geometry = st_union(geometry))
ura_planning_area_map <- tm_shape(ura_planning_areas, name = "URA planning areas") + 
  tm_polygons(
    col = "grey", fill = "PLN_AREA_N", fill_alpha = 0.25, 
    fill.legend = tm_legend_hide())

# URA subzones map
ura_subzones <- sg_ura_boundaries %>% group_by(SUBZONE_N) %>% 
  summarise(geometry = st_union(geometry))
ura_subzone_map <- tm_shape(ura_subzones, name = "URA subzones") + tm_polygons(
  col = "grey", fill = "SUBZONE_N", fill_alpha = 0.25, 
  fill.legend = tm_legend_hide())

# MOE zones map
moe_zones <- obesity_data %>% group_by(ZONE) %>% 
  summarise(geometry = st_union(geometry)) %>% st_convex_hull()
moe_zone_map <- tm_shape(moe_zones, name = "MOE zones") + tm_polygons(
  col = "grey", fill = "ZONE", fill_alpha = 0.25, 
  fill.legend = tm_legend(title = "MOE School Zones"))

# MOE clusters map
moe_clusters <- obesity_data %>% group_by(CLUSTER) %>% 
  summarise(geometry = st_union(geometry)) %>% st_convex_hull()
moe_cluster_map <- tm_shape(moe_clusters, name = "MOE clusters") + tm_polygons(
  col = "grey", fill = "CLUSTER", fill_alpha = 0.25, 
  fill.legend = tm_legend_hide())
```

## Obesity prevalence overview map

```{r}
tmap_mode("view")

# for now, filter for 2023 data and overweight and severely overweight
obesity_filtered <- obesity_data_sf %>% 
  filter(YEAR == 2023, CATEGORY == "OSO")

# Overview of overweight prevalence levels in Singapore schools
ura_region_map + ura_planning_area_map + ura_subzone_map + 
  moe_zone_map + moe_cluster_map +
  tm_shape(obesity_filtered, name = "Overweight prevalence") +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "CENTER"),
      title = "Educational level"
      ),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape",
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overview of overweight prevalence in Singapore schools")
```

## Obesity prevalence by educational level

### by Primary School
```{r}
tmap_mode("view")

# filter out by educational level and remove NA values to calculate SDE
obesity_pri_sch <- obesity_filtered %>% 
  filter(PRI == 1, !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
pri_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_pri_sch))
pri_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_pri_sch))
pri_sch_mean_median <- rbind(pri_sch_mean$LOCATIONS, pri_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
pri_sch_mean_median$id = c("Spatial mean", "Spatial median")

pri_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_pri_sch))
pri_sch_sdd_line <- pri_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_pri_sch))
pri_sch_sde_line <- pri_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_pri_sch$PREVALENCE,
  points = st_coordinates(obesity_pri_sch))
pri_sch_sde_weighted_line <- pri_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_map <- ura_region_map +
  tm_shape(obesity_pri_sch, name = "Overweight prevalence") +
  tm_symbols(
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  tm_shape(pri_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(pri_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(pri_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(pri_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Overweight prevalence in Primary Schools")

print(pri_sch_map)
```

### by Secondary School
```{r}
tmap_mode("view")

# filter out by educational level and remove NA values to calculate SDE
obesity_sec_sch <- obesity_filtered %>% 
  filter(SEC == 1, !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
sec_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_sec_sch))
sec_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_sec_sch))
sec_sch_mean_median <- rbind(sec_sch_mean$LOCATIONS, sec_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
sec_sch_mean_median$id = c("Spatial mean", "Spatial median")

sec_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_sec_sch))
sec_sch_sdd_line <- sec_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_sec_sch))
sec_sch_sde_line <- sec_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_sec_sch$PREVALENCE,
  points = st_coordinates(obesity_sec_sch))
sec_sch_sde_weighted_line <- sec_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_map <- ura_region_map +
  tm_shape(obesity_sec_sch, name = "Overweight prevalence") +
  tm_symbols(
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  tm_shape(sec_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(sec_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(sec_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(sec_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Overweight prevalence in Secondary Schools")

print(sec_sch_map)

# tmap_arrange(pri_sch_map, sec_sch_map, nrow = 1)
```

## Aggregating overweight counts by different defined regions
```{r}
# Aggregating overweight counts by URA regions
obesity_ura_regions <- st_join(sg_ura_boundaries, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(REGION_N) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_region_map + obesity_ura_regions %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA regions")

# Aggregating overweight counts by URA planning areas
obesity_ura_planning_areas <- st_join(sg_ura_boundaries, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(PLN_AREA_N) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_planning_area_map + obesity_ura_planning_areas %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA planning areas")

# Aggregating overweight counts by URA subzones
obesity_ura_subzones <- st_join(sg_ura_boundaries, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(SUBZONE_N) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_subzone_map + obesity_ura_subzones %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA subzones")

# Aggregating overweight counts by MOE school zones
obesity_moe_zones <- 
  # we left_join instead of st_join due to overlapping boundaries
  left_join(moe_zones, st_drop_geometry(obesity_data_sf)) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(ZONE) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

moe_zone_map +
  tm_legend_hide() + 
  obesity_moe_zones %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill_alpha = 0.95,
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1,
      show = T
      )
    ) +
  tm_title("Overweight prevalence by MOE clusters")

# Aggregating overweight counts by MOE clusters
obesity_moe_clusters <- 
  left_join(moe_clusters, st_drop_geometry(obesity_data_sf)) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(CLUSTER) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

moe_cluster_map + obesity_moe_clusters %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill_alpha = 0.95,
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by MOE clusters")
```

```{r}
# Try visualising obesity data in a chloropleth map
```

## Spatial dependence of obesity in Singapore schools
```{r}
sg_ura_boundaries_nb <- poly2nb(sg_ura_boundaries) # queen contiguity
sg_subzone_coords <- st_point_on_surface(st_geometry(sg_ura_boundaries))
plot(st_geometry(sg_ura_boundaries), border="grey60", axes=T,  main="Queen Contiguity")
plot(sg_ura_boundaries_nb, sg_subzone_coords, pch=19, cex=0.6, add=T)

sg_ura_lw <- nb2listw(sg_ura_boundaries_nb, style = "W", zero.policy = T)
sg_ura_lw_B <- nb2listw(sg_ura_boundaries_nb, style = "B", zero.policy = T)
```


# Income distribution (Lucas)

```{r}
census <- read.csv("census stats.csv") %>% 
  mutate(PLN_AREA_N = toupper(PLN_AREA_N))
combine_data <- left_join(ura_planning_areas, census)
combine_data
summary(census)
areas <- c(
"ANG MO KIO", "BEDOK", "BISHAN", "BUKIT BATOK", "BUKIT MERAH", "BUKIT PANJANG",
"BUKIT TIMAH", "CHOA CHU KANG", "CLEMENTI", "DOWNTOWN CORE", "GEYLANG", "HOUGANG",
"JURONG EAST", "JURONG WEST", "KALLANG", "MARINE PARADE", "OUTRAM", "NOVENA", "PASIR RIS", "PUNGGOL",   "QUEENSTOWN", "RIVER VALLEY", "SEMBAWANG", "SENGKANG", "SERANGOON", "TAMPINES",
"TANGLIN", "TOA PAYOH", "WOODLANDS", "YISHUN"
)
for (area in areas) {
area_var <- gsub(" ", "_", tolower(area))
assign(paste0(area_var, "_set"), subset(combine_data, PLN_AREA_N == area))
assign(area_var, st_union(get(paste0(area_var, "_set"))))
assign(paste0(area_var, "_point"), st_centroid(get(area_var)))
}
st_coordinates(ang_mo_kio_point)
st_coordinates(bedok_point)
st_coordinates(bishan_point)
st_coordinates(bukit_batok_point)
st_coordinates(bukit_merah_point)
st_coordinates(bukit_panjang_point)
st_coordinates(bukit_timah_point)
st_coordinates(choa_chu_kang_point)
st_coordinates(clementi_point)
st_coordinates(downtown_core_point)
st_coordinates(geylang_point)
st_coordinates(hougang_point)
st_coordinates(jurong_east_point)
st_coordinates(jurong_west_point)
st_coordinates(kallang_point)
st_coordinates(marine_parade_point)
st_coordinates(outram_point)
st_coordinates(novena_point)
st_coordinates(pasir_ris_point)
st_coordinates(punggol_point)
st_coordinates(queenstown_point)
st_coordinates(river_valley_point)
st_coordinates(sembawang_point)
st_coordinates(sengkang_point)
st_coordinates(serangoon_point)
st_coordinates(tampines_point)
st_coordinates(tanglin_point)
st_coordinates(toa_payoh_point)
st_coordinates(woodlands_point)
st_coordinates(yishun_point)
points_df = data.frame(PLN_AREA_N = c("ANG MO KIO", "BEDOK", "BISHAN", "BUKIT BATOK", "BUKIT MERAH", "BUKIT PANJANG",
"BUKIT TIMAH", "CHOA CHU KANG", "CLEMENTI", "DOWNTOWN CORE", "GEYLANG", "HOUGANG",
"JURONG EAST", "JURONG WEST", "KALLANG", "MARINE PARADE", "OUTRAM", "NOVENA", "PASIR RIS", "PUNGGOL",   "QUEENSTOWN", "RIVER VALLEY", "SEMBAWANG", "SENGKANG", "SERANGOON", "TAMPINES",
"TANGLIN", "TOA PAYOH", "WOODLANDS", "YISHUN"),
lat=c(40229.12,34032.1,37450.89,37527.65,28662.87,38761.84,34689,40833.42,33319.17,29721.45,33683.82,38409.7,33633.69,35974.53,32515.77,31398.33,29477.49,34356.64,40066.41,43023.56,30164.7,31156.94,48594.64,41466.18,38582.59,36200.47,32298.93,35429.21,46928.63,44086.22),
long = c(28976.88,38582.67,28789.76,19255.42,26865.78,21287.04,23256.76,18415.05,19923.36,30335.19,34255.04,34285.76,17035.36,13708.69,31685.37,35005.86,28972.96,28221.51,40795.84,36538.42,22573.46,28072.91,26387.4,34051.82,31839.38,41459.48,26250.09,31083.53,22923.8,28404.04))
points_df
points_sf <- st_as_sf(points_df, coords = c("long", "lat"), crs = sg_crs)
points_sf
census_merge = left_join(points_sf,census)
ura_planning_area_map+tm_shape(census_merge)+ tm_symbols(size=0.5,fill="Income")
ura_planning_area_map+tm_shape(census_merge)+ tm_symbols(size=0.5,fill="Education")


tm_shape(combine_data) + 
  tm_fill(fill="Income", 
          fill.scale = tm_scale(breaks = c(6925,8396,9532,11317,16197))
          ) +
  tm_title("Breakdown by Household Income")
tm_shape(combine_data) + 
  tm_fill(fill="Education", 
          fill.scale = tm_scale(breaks = c(9.76,10.74,11.25,11.89,14.78))
          ) +
  tm_title("Breakdown by Education Level")
tm_shape(combine_data) + 
  tm_fill(fill="Primary", 
          fill.scale = tm_scale(breaks = c(61,4778,7043,11681,21822))
          ) +
  tm_title("Breakdown by Pri Students")
tm_shape(combine_data) + 
  tm_fill(fill="Secondary", 
          fill.scale = tm_scale(breaks=c(59,3526,4894,8804,13221))
          ) +
  tm_title("Breakdown by Sec Students")

census_filtered <- combine_data %>% filter(!is.na(Income))
census_coords <- st_coordinates(census_filtered)
```

# Spatial Descriptive Summary Measures of Income (Lucas)

```{r}
income_mean <- calc_mnc(id = 1, points = census_coords, weighted = T, weights = census_filtered$Income)
income_median <- calc_mdc(id = 2, points = census_coords, weighted = T, weights = census_filtered$Income)
income_mean_median <- rbind(income_mean$LOCATIONS, income_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
income_mean_median$id = c("Weighted mean", "Weighted median")

income_sdd <- calc_sdd(id = 3, points = census_coords)
income_sdd_line <- income_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

income_sde <- calc_sde(id = 4, points = census_coords)
income_sde_line <- income_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

income_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = census_filtered$Income,
  points = census_coords)
income_sde_weighted_line <- income_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")
```

# Food and Nutrition (Nina)

```{r}
#Hawker centre data
hawker_full = st_read("NEAMarketandFoodCentre.geojson") %>%
  st_transform(crs = sg_crs)

hawker = hawker_full["geometry"]


#Fast Food Outlets (Mcdonald's and 711)
macs_full = st_read("mcdonalds_singapore.geojson") %>%
  st_transform(crs = sg_crs)

  #Check validity
table(st_is_valid(macs_full))

  #Check CRS
st_crs(macs_full)$epsg

  #Extract necessary columns for analysis
macs = macs_full[,"geometry"]

  #7-11 Outlets
seven_eleven_full = st_read("711_singapore.geojson") %>%
  st_transform(crs = sg_crs)


  #Extract necessary columns for analysis
seven_eleven = seven_eleven_full[,"geometry"]
  
  #Check validity
table(st_is_valid(seven_eleven))

  #Check CRS
st_crs(seven_eleven)$epsg

  #Plot of 711 outlets
tm_shape(sg_ura_boundaries) + 
  tm_borders() +
  tm_shape(seven_eleven) + 
  tm_dots(col = "green", size = 0.3)

#Combine data into single SF object
seven_eleven_sf = mutate(seven_eleven, outlet = "7-11")
macs_sf = mutate(macs, outlet = "mcdonalds")
hawker_sf = mutate(hawker, outlet = "hawker")
food = rbind(seven_eleven_sf, macs_sf, hawker_sf)


#Primary school analysis

#duplicate primary school data
obesity_pri_food = obesity_pri_sch

#check units
st_crs(obesity_pri_food)$units
pri_schools_buffer1 <- st_buffer(obesity_pri_food, dist = 1000) %>% st_geometry()
pri_intersections <- st_intersects(pri_schools_buffer1, food)

#calculate the number of food outlets within a 1km radius of each school
obesity_pri_food$num_food_outlets <- lengths(pri_intersections)
obesity_pri_food

#show density via choropleth map
ura_region_map + 
  tm_shape(obesity_pri_food) +
  tm_fill(col = "num_food_outlets", 
          palette = "brewer.yl_or_rd",
          alpha = 0.7) + 
  tm_borders(lwd = 0.5) +
  tm_title("Food Outlet Density by Primary School")


#Secondary school analysis

#duplicate primary school data
obesity_sec_food = obesity_sec_sch

#check units
st_crs(obesity_sec_food)$units
sec_schools_buffer1 <- st_buffer(obesity_sec_food, dist = 1000) %>% st_geometry()
sec_intersections <- st_intersects(sec_schools_buffer1, food)

#calculate the number of food outlets within a 1km radius of each school
obesity_sec_food$num_food_outlets <- lengths(sec_intersections)
obesity_sec_food

#show density via choropleth map
ura_region_map + 
  tm_shape(obesity_sec_food) +
  tm_fill(col = "num_food_outlets", 
          palette = "brewer.yl_or_rd",
          alpha = 0.7) + 
  tm_borders(lwd = 0.5) +
  tm_title("Food Outlet Density by Secondary School")
```

# Physical activity (Qasim)

```{r}
# Loading school co-curricular activities data
school_sports_ccas <- 
  read.csv("CocurricularactivitiesCCAs.csv") %>% 
  group_by(school_name) %>% 
  summarise(
    total = n(),
    physical_sports_count = sum(cca_grouping_desc == "PHYSICAL SPORTS")) %>% 
    select(school_name, physical_sports_count)


#2023 OSO school obesity data combined with 2024-25 school physical activity cca data
#gives spatial awareness to physical activity cca by joining to obesity data set by school name
obesity_joined_school_sports_data <- 
  obesity_filtered %>% 
  left_join(school_sports_ccas, by = join_by(SCHOOL == school_name)) %>%
  mutate(sports_cca_per_100 = physical_sports_count / ENROLMENT * 100, #ratio of sports cca per 100 students
         LEVEL = case_when(
           LEVEL == "PRIMARY + SECONDARY" & PRI == 1 ~ "PRIMARY",
           LEVEL == "PRIMARY + SECONDARY" & PRI == 0 ~ "SECONDARY",
           TRUE ~ LEVEL
         )
        ) 


#Visualising physical sports participation by school
#uses same visual style as obesity map for continuity
#scale colour is reversed (green is high, red is low) to indicate higher physical sports participation
sports_ccas_map <- tm_shape(obesity_joined_school_sports_data) +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "RIGHT"),
      title = "Educational level"
      ),
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(
      values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Ratio of Sports CCAs per 100 Students by School",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    )
sports_ccas_map


#same plot but displaying each education level separately
school_level_sports_map <- tm_shape(obesity_joined_school_sports_data) +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "RIGHT"),
      title = "Educational level",
      ),
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(
      values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Ratio of Sports CCAs per 100 Students by School",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + tm_facets_wrap(by = "LEVEL") #facet by education level

school_level_sports_map


tm_shape(obesity_joined_school_sports_data) +
  tm_bubbles(
    shape = "LEVEL",                
    size = "sports_cca_per_100",   #(size.scale = "PREVALENCE" - size too big)    
    col = "PREVALENCE",    
    palette = "brewer.rd_yl_gn",               
    border.col = "black",               
    title.size = "Ratio of Sports CCAs per 100 Students",
    title = "Physical Sports Participation (%)"
  ) +
  tm_layout(
    title = "Obesity Prevalence vs Physical Sports CCAs",
    legend.outside = TRUE 
  ) +tm_facets_wrap(by = "LEVEL") 

#bivariate chloropleth for obesity prevalence and physical sports participation?



```

# Other Useful stuff

```{r}
tmap_mode("view") #interactive map
tmap_mode("plot") #static map
# tmap_save(tmap_object, 'xx.jpg') # for saving maps as images
cols4all::c4a_gui() # colour scales
RColorBrewer::display.brewer.all()
```
