# Data loading, cleaning and transformation

## Loading the data

Change the file paths accordingly

```{r}
# Loading libraries to work with spatial data
library(tidyverse); library(sf); library(tmap); library(aspace)

# Loading Singapore URA subzone boundary data
sg_ura_boundaries <- 
  st_read("MasterPlan2014SubzoneBoundaryWebSHP/MP14_SUBZONE_WEB_PL.shp") %>% 
  st_make_valid()

# Saving CRS provided by SG boundary map
sg_crs <- st_crs(sg_ura_boundaries)

# Loading cleaned obesity data from HPB and giving spatial awareness
obesity_data <- 
  # "#N/A" values converted to "NA" for R to be able to read it
  read.csv("obesity.csv", na.strings = "#N/A") %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  st_transform(crs = sg_crs) %>% 
  st_make_valid()

# Loading cleaned HMSP data from HPB and giving spatial awareness
hmsp_data_sf <- 
  read.csv("hmsp_tier_pri_sec.csv") %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  st_transform(crs = sg_crs) %>% 
  st_make_valid() 

# Checking that all CRS are matching
isTRUE(all.equal(st_crs(sg_ura_boundaries), st_crs(obesity_data), st_crs(hmsp_data_sf)))
```

## Transforming data to be usable for analysis

```{r}
obesity_data_sf <- obesity_data %>% 
  pivot_longer(
    # pivot_longer to separate obesity data columns into a tidy format
    cols = starts_with("X"),
    names_to = c("YEAR", "CATEGORY"),
    names_prefix = "X",
    names_sep = "_",
    names_transform = list(YEAR = as.integer),
    values_to = "COUNT"
    ) %>% 
  mutate(
    # Transform obesity count and school enrollment into an obesity prevalence figure
    PREVALENCE = COUNT / ENROLMENT
    )
```

# Exploratory data analysis (by Prof)

```{r}
tmap_mode("plot")
tm_shape(sg_ura_boundaries) + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.4, fill = "TIER", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "HMSP Tiers"))

tm_shape(sg_ura_boundaries) + tm_fill() + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.4, fill = "ZONE", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "School zones"))

tm_shape(sg_ura_boundaries) + tm_fill() + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.3, fill = "CLUSTER", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "School clusters")) +
  tm_facets_wrap(by = "ZONE")
```

# Obesity profile in Singapore (Colin)

## Singapore maps

```{r}
tmap_mode("view")

# URA regions map
ura_region_map <- tm_shape(sg_ura_boundaries, name = "URA regions") + tm_polygons(
    col = "grey", fill = "REGION_N", fill_alpha = 0.25, 
    fill.legend = tm_legend(title = "URA regions"))

# URA planning areas map
ura_planning_area_map <- tm_shape(sg_ura_boundaries, name = "URA planning areas") + 
  tm_polygons(
    col = "grey", fill = "PLN_AREA_N", fill_alpha = 0.25, 
    fill.legend = tm_legend_hide())

# URA subzones map
ura_subzone_map <- tm_shape(sg_ura_boundaries, name = "URA subzones") + tm_polygons(
  col = "grey", fill = "SUBZONE_N", fill_alpha = 0.25, 
  fill.legend = tm_legend_hide())

# MOE zones map
moe_zones <- obesity_data %>% group_by(ZONE) %>% 
  summarise(geometry = st_union(geometry)) %>% st_convex_hull()
moe_zone_map <- tm_shape(moe_zones, name = "MOE zones") + tm_polygons(
  col = "grey", fill = "ZONE", fill_alpha = 0.25, 
  fill.legend = tm_legend(title = "MOE School Zones"))

# MOE clusters map
moe_clusters <- obesity_data %>% group_by(CLUSTER) %>% 
  summarise(geometry = st_union(geometry)) %>% st_convex_hull()
moe_cluster_map <- tm_shape(moe_clusters, name = "MOE clusters") + tm_polygons(
  col = "grey", fill = "CLUSTER", fill_alpha = 0.25, 
  fill.legend = tm_legend_hide())
```

## Obesity prevalence overview map

```{r}
tmap_mode("view")

# for now, filter for 2023 data and overweight and severely overweight
obesity_filtered <- obesity_data_sf %>% 
  filter(YEAR == 2023, CATEGORY == "OSO")

# Overview of overweight prevalence levels in Singapore schools
ura_region_map + ura_planning_area_map + ura_subzone_map + 
  moe_zone_map + moe_cluster_map +
  tm_shape(obesity_filtered, name = "Overweight prevalence") +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "CENTER"),
      title = "Educational level"
      ),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape",
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overview of overweight prevalence in Singapore schools")
```

## Obesity prevalence by educational level

### by Primary School
```{r}
tmap_mode("view")

# filter out by educational level and remove NA values to calculate SDE
obesity_pri_sch <- obesity_filtered %>% 
  filter(PRI == 1, !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
pri_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_pri_sch))
pri_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_pri_sch))
pri_sch_mean_median <- rbind(pri_sch_mean$LOCATIONS, pri_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
pri_sch_mean_median$id = c("Spatial mean", "Spatial median")

pri_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_pri_sch))
pri_sch_sdd_line <- pri_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_pri_sch))
pri_sch_sde_line <- pri_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_pri_sch$PREVALENCE,
  points = st_coordinates(obesity_pri_sch))
pri_sch_sde_weighted_line <- pri_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_map <- ura_region_map +
  tm_shape(obesity_pri_sch, name = "Overweight prevalence") +
  tm_symbols(
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  tm_shape(pri_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(pri_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(pri_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(pri_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Overweight prevalence in Primary Schools")

print(pri_sch_map)
```

### by Secondary School
```{r}
tmap_mode("view")

# filter out by educational level and remove NA values to calculate SDE
obesity_sec_sch <- obesity_filtered %>% 
  filter(SEC == 1, !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
sec_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_sec_sch))
sec_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_sec_sch))
sec_sch_mean_median <- rbind(sec_sch_mean$LOCATIONS, sec_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
sec_sch_mean_median$id = c("Spatial mean", "Spatial median")

sec_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_sec_sch))
sec_sch_sdd_line <- sec_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_sec_sch))
sec_sch_sde_line <- sec_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_sec_sch$PREVALENCE,
  points = st_coordinates(obesity_sec_sch))
sec_sch_sde_weighted_line <- sec_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_map <- ura_region_map +
  tm_shape(obesity_sec_sch, name = "Overweight prevalence") +
  tm_symbols(
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  tm_shape(sec_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(sec_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(sec_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(sec_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Overweight prevalence in Secondary Schools")

print(sec_sch_map)

# tmap_arrange(pri_sch_map, sec_sch_map, nrow = 1)
```

## Aggregating overweight counts by different defined regions
```{r}
# Aggregating overweight counts by URA regions
obesity_ura_regions <- st_join(sg_ura_boundaries, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(REGION_N) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_region_map + obesity_ura_regions %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA regions")

# Aggregating overweight counts by URA planning areas
obesity_ura_planning_areas <- st_join(sg_ura_boundaries, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(PLN_AREA_N) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_planning_area_map + obesity_ura_planning_areas %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA planning areas")

# Aggregating overweight counts by URA subzones
obesity_ura_subzones <- st_join(sg_ura_boundaries, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(SUBZONE_N) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_subzone_map + obesity_ura_subzones %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA subzones")

# Aggregating overweight counts by MOE school zones
obesity_moe_zones <- 
  # we left_join instead of st_join due to overlapping boundaries
  left_join(moe_zones, st_drop_geometry(obesity_data_sf)) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(ZONE) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

moe_zone_map +
  tm_legend_hide() + 
  obesity_moe_zones %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill_alpha = 0.95,
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1,
      show = T
      )
    ) +
  tm_title("Overweight prevalence by MOE clusters")

# Aggregating overweight counts by MOE clusters
obesity_moe_clusters <- 
  left_join(moe_clusters, st_drop_geometry(obesity_data_sf)) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(CLUSTER) %>% 
  summarise(
    PREVALENCE_AGG = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

moe_cluster_map + obesity_moe_clusters %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGG",
    fill_alpha = 0.95,
    fill.scale = tm_scale_continuous(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by MOE clusters")
```

```{r}
# Try visualising obesity data in a chloropleth map
```

# Income distribution (Lucas)

```{r}
census <- read.csv("census stats.csv") %>% 
  mutate(PLN_AREA_N = toupper(PLN_AREA_N))
combine_data <- left_join(sg_ura_boundaries, census)
View(combine_data)
summary(census)
tm_shape(combine_data) + tm_fill(fill="Income", fill.scale = tm_scale(breaks=c(6925,8396,9532,11317,16197)))+tm_title("Breakdown by Household Income")
tm_shape(combine_data) + tm_fill(fill="Education", fill.scale = tm_scale(breaks=c(9.76,10.74,11.25,11.89,14.78)))+tm_title("Breakdown by Education Level")
tm_shape(combine_data) + tm_fill(fill="Primary", fill.scale = tm_scale(breaks=c(61,4778,7043,11681,21822)))+tm_title("Breakdown by Pri Students")
tm_shape(combine_data) + tm_fill(fill="Secondary", fill.scale = tm_scale(breaks=c(59,3526,4894,8804,13221)))+tm_title("Breakdown by Sec Students")
census_filtered <- combine_data %>% filter(Income != NA)
census_coords <- st_coordinates(census_filtered)

```

# Spatial Descriptive Summary Measures of Income (Lucas)

```{r}
income_mean <- calc_mnc(id = 1, points = st_coordinates(census_coords), weighted = T, weights = census_filtered$Income)
sec_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_sec_sch))
sec_sch_mean_median <- rbind(sec_sch_mean$LOCATIONS, sec_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
sec_sch_mean_median$id = c("Spatial mean", "Spatial median")

sec_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_sec_sch))
sec_sch_sdd_line <- sec_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_sec_sch))
sec_sch_sde_line <- sec_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_sec_sch$PREVALENCE,
  points = st_coordinates(obesity_sec_sch))
sec_sch_sde_weighted_line <- sec_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")
```

# Food and Nutrition (Nina)

```{r}

```

# Physical activity (Qasim)

```{r}
# Loading school co-curricular activities data
school_sports_ccas <- 
  read.csv("CocurricularactivitiesCCAs.csv") %>% 
  group_by(school_name) %>% 
  summarise(
    total = n(),
    physical_sports_count = sum(cca_grouping_desc == "PHYSICAL SPORTS")) %>% 
    select(school_name, physical_sports_count)


#2023 OSO school obesity data combined with 2024-25 school physical activity cca data
#gives spatial awareness to physical activity cca by joining to obesity data set by school name
obesity_joined_school_sports_data <- 
  obesity_filtered %>% 
  left_join(school_sports_ccas, by = join_by(SCHOOL == school_name)) %>%
  mutate(sports_cca_per_100 = physical_sports_count / ENROLMENT * 100, #ratio of sports cca per 100 students
         LEVEL = case_when(
           LEVEL == "PRIMARY + SECONDARY" & PRI == 1 ~ "PRIMARY",
           LEVEL == "PRIMARY + SECONDARY" & PRI == 0 ~ "SECONDARY",
           TRUE ~ LEVEL
         )
        ) 


#Visualising physical sports participation by school
#uses same visual style as obesity map for continuity
#scale colour is reversed (green is high, red is low) to indicate higher physical sports participation
sports_ccas_map <- tm_shape(obesity_joined_school_sports_data) +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "RIGHT"),
      title = "Educational level"
      ),
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(
      values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Ratio of Sports CCAs per 100 Students by School",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    )
sports_ccas_map


#same plot but displaying each education level separately
school_level_sports_map <- tm_shape(obesity_joined_school_sports_data) +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "RIGHT"),
      title = "Educational level",
      ),
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(
      values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Ratio of Sports CCAs per 100 Students by School",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + tm_facets_wrap(by = "LEVEL") #facet by education level

school_level_sports_map


tm_shape(obesity_joined_school_sports_data) +
  tm_bubbles(
    shape = "LEVEL",                
    size = "sports_cca_per_100",   #(size.scale = "PREVALENCE" - size too big)    
    col = "PREVALENCE",    
    palette = "brewer.rd_yl_gn",               
    border.col = "black",               
    title.size = "Ratio of Sports CCAs per 100 Students",
    title = "Physical Sports Participation (%)"
  ) +
  tm_layout(
    title = "Obesity Prevalence vs Physical Sports CCAs",
    legend.outside = TRUE 
  ) +tm_facets_wrap(by = "LEVEL") 

#bivariate chloropleth for obesity prevalence and physical sports participation?



```

# Other Useful stuff

```{r}
tmap_mode("view") #interactive map
tmap_mode("plot") #static map
# tmap_save(tmap_object, 'xx.jpg') # for saving maps as images
cols4all::c4a_gui() # colour scales
RColorBrewer::display.brewer.all()
```
