---
title: Childhood Obesity in Singapore: A Spatial Analysis of Obesity Prevalence and Post-COVID Trends in Schools
author: Colin, Lucas, Nina, Qasim
output: 
  html_document:
    toc: true
    number_sections: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Data loading, cleaning and transformation

## Spatial data libraries

```{r}
library(tidyverse); library(sf); library(tmap); library(aspace); library(spdep); library(corrplot); library(readxl); library(spatialreg)
```

To support spatial and statistical analysis, several R libraries were loaded. tidyverse was used for data manipulation, sf for spatial data handling, and tmap for thematic mapping. aspace and spdep enabled spatial analysis, while corrplot visualised correlations. readxl facilitated reading Excel files, and spatialreg supported spatial regression modelling. These libraries ensure a robust analysis of childhood obesity trends in Singapore.

## Loading and joining HPB and relevant datasets

```{r}
# Loading Singapore URA subzone boundary data
sg_ura_boundaries <- 
  st_read("MasterPlan2014SubzoneBoundaryWebSHP/MP14_SUBZONE_WEB_PL.shp") %>% 
  st_make_valid()

# Saving CRS provided by SG boundary map
sg_crs <- st_crs(sg_ura_boundaries) # SVY21

# Loading cleaned obesity data from HPB and giving spatial awareness
obesity_data <- 
  read.csv("obesity.csv", na.strings = "#N/A") %>% # "#N/A" values converted to "NA" for R to be able to read it
  select(-PT_ID.1) %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  st_transform(crs = sg_crs) %>% 
  st_make_valid()

# Loading cleaned HMSP data from HPB and giving spatial awareness
hmsp_data_sf <- 
  read.csv("hmsp_tier_pri_sec.csv") %>% 
  rename(HMSP_Tier = TIER, HMSP_Score = TIER_SCORE) %>% 
  st_as_sf(coords = c("longitude","latitude"), crs = 4326) %>% 
  st_transform(crs = sg_crs) %>% 
  st_make_valid()

# Loading General Information of Schools dataset
school_info <- 
  read_excel("Generalinformationofschools.xlsx", na = c("na", "NA")) %>% 
  select(school_name, dgp_code:mothertongue3_code, -zone_code, -mainlevel_code) # dropped variables are repeated in other data frames or not useful for our analysis

# Loading previous year enrolment data
school_enrolment <-
  read_excel("enrolment.xlsx", sheet = "primary") %>% 
  select(-starts_with("intake"), -ENROLMENT24) # drop enrolment24 as obesity_data contains 2024 enrolment data for the full set of schools

# Checking that all CRS are matching
isTRUE(all.equal(st_crs(sg_ura_boundaries), st_crs(obesity_data), st_crs(hmsp_data_sf)))

# Checking CRS units 
st_crs(sg_ura_boundaries)$units # metres
st_crs(obesity_data)$units # metres
st_crs(hmsp_data_sf)$units # metres

# Joining data
obesity_data <- obesity_data %>% 
  left_join(school_info, by = join_by(SCHOOL == school_name)) %>% # data was previously cleaned to ensure that school names match 
  left_join(select(st_drop_geometry(hmsp_data_sf), PT_ID, HMSP_Score, HMSP_Tier), by = join_by(PT_ID))
```

The [Urban Redevelopment Authority (URA) Master Plan 2014 Subzone Boundary (No Sea)](https://data.gov.sg/collections/1707/view) dataset to serve as our base geometry map of Singapore. To facilitate urban planning, the Urban Redevelopment Authority (URA) divides Singapore into regions, planning areas and subzones. This would aid our investigation of the regional patterns of childhood obesity.

The coordinate reference system of the URA boundary data was extracted and stored for consistency across our spatial datasets. ([SVY21](https://app.sla.gov.sg/sirent/About/PlaneCoordinateSystem): based on the Transverse Mercator projection from geographical coordinates referenced to the WGS 84 ellipsoid) For further details on the SVY21 CRS, please view the linked Singapore Land Authority (SLA) website .

Next, we imported the cleaned obesity and Healthy Meals in Schools Programme (HMSP) datasets provided by the Health Promotion Board (HPB). This data consists of yearly student obesity counts by schools, and HMSP Tiers and Scores for school canteen stalls. 

The [General Information of Schools](https://data.gov.sg/datasets/d_688b934f82c1059ed0a6993d2a829089/view) dataset contains key information for primary, secondary and pre-university schools, like whether they were government or co-ed schools or had special programmes (SAP, gifted, IP, etc.). These variables would add further nuance to our analysis.

Past year enrolment data was also joined with our obesity datasets. School enrolment data was only provided for 2024, whereas obesity counts were provided from 2016 to 2023 (excluding 2021). Given Singapore's low birthrates and the shrinking student population over the years, an obesity prevalence measure would be inaccurate if solely calculated using 2024 enrolment data. Past year enrollment data was thus estimated using yearly intake data (for Primary schools only, due to intake data availability), e.g. the 2023 enrolment data in any school was estimated by removing the net intake in 2024, where the net intake = no. of students entering Primary 1 - no. of students graduating from Primary 6 (estimated from the P1 intake from 6 years ago).

All datasets with coordinate geometries were transformed into a spatial object using the st_as_sf() function, with longitude and latitude as coordinates, and reprojected to match the CRS of the URA boundary data. Spatial validity was checked and made valid using st_make_valid() to ensure topological correctness. Finally, the CRS units for each dataset were checked to confirm uniformity, ensuring accurate spatial mapping and analysis. These steps were essential for integrating multiple data sources within a consistent spatial framework, allowing for precise geographic visualisation and interpretation of childhood obesity patterns in Singapore.

## Additional data transformations for exploratory data analysis

```{r}
# Changing "PRIMARY + SECONDARY" level label to match PRI and SEC classification
obesity_data <- obesity_data %>% mutate(
  LEVEL = case_when(
    LEVEL == "PRIMARY + SECONDARY" & PRI == 1 ~ "PRIMARY",
    LEVEL == "PRIMARY + SECONDARY" & PRI == 0 ~ "SECONDARY",
    TRUE ~ LEVEL
    )
  ) %>% select(-c("PRI", "SEC"))

# Classifying schools with special programmes with a "special" variable for further analysis
obesity_data <- obesity_data %>% mutate(
  special = case_when(
    sap_ind == "Yes" | autonomous_ind == "Yes" | gifted_ind == "Yes" | ip_ind == "Yes" ~ "Yes",
    TRUE ~ "No"
    )
  )

# Converting school_enrolment from wide to long format
school_enrolment_long <- obesity_data %>% 
  rename(ENROLMENT24 = ENROLMENT) %>% 
  select(PT_ID, SCHOOL, ENROLMENT24) %>% 
  st_drop_geometry() %>% 
  # join 2024 enrollment data from obesity_data with school_enrolment data
  left_join(school_enrolment, by = join_by(PT_ID)) %>% 
  select(-SCHOOL.y) %>% 
  rename(SCHOOL = SCHOOL.x) %>% 
  pivot_longer(cols = starts_with("ENROLMENT"), names_to = "YEAR", values_to = "ENROLMENT") %>%
  # Extract year as numeric
  mutate(YEAR = as.numeric(gsub("ENROLMENT", "", YEAR)) + 2000) %>% 
  group_by(PT_ID) %>% 
  arrange(YEAR, .by_group = T)

# For schools where enrolment data is only available for 2024, replace the NA value in 2023 with the 2024 value. (This is done as most of our spatial analysis is done for the 2023 obesity prevalence data)
for (i in 1:nrow(school_enrolment_long)) {
  if (school_enrolment_long[i,3] == 2023 & is.na(school_enrolment_long[i,4])) {
    school_enrolment_long[i,4] = school_enrolment_long[i+1,4]
    }
  }

# Convert obesity_data from wide to long format
obesity_data_long <- obesity_data %>%
  pivot_longer(cols = starts_with("X"), 
               names_to = "YEAR_CATEGORY", 
               values_to = "COUNT") %>%
  separate(YEAR_CATEGORY, into = c("YEAR", "CATEGORY"), sep = "_", extra = "merge") %>%
  # Extract year as numeric
  mutate(YEAR = as.numeric(gsub("X", "", YEAR))) 

# Creating obesity prevalence measure from obesity count and school enrolment 
obesity_data_sf <- obesity_data_long %>%
  select(-ENROLMENT) %>% 
  full_join(school_enrolment_long, by = c("PT_ID", "SCHOOL", "YEAR")) %>%
  # Calculate overweight prevalence rate
  mutate(PREVALENCE = COUNT / ENROLMENT) %>% 
  relocate(PREVALENCE, .after = SCHOOL)

# For our analysis, filter for 2023 data and overweight and severely overweight
obesity_data_sf_2023_OSO <- obesity_data_sf %>% 
  filter(YEAR == 2023, CATEGORY == "OSO", !(SCHOOL %in% c("FUCHUN SECONDARY SCHOOL", "TANGLIN SECONDARY SCHOOL","GUANGYANG PRIMARY SCHOOL")), !is.na(PREVALENCE))

# # Combining prof's data with ours (delete this line pls)
# oso23 = read_csv("Prof's Codes/oso23.csv", show_col_types = FALSE)
# oso23 = oso23 |> drop_na(OSO_pct) 
# 
# # Re-classify
# oso23[,c(3, 4, 8, 13:26)] = lapply(oso23[,c(3, 4, 8, 13:26)], as.factor)
# oso23_new = oso23 %>% select(c("PT_ID", "HMSP_Tier", "HMSP_Score", "special"))
# obesity_data_sf_2023_OSO_new = left_join(obesity_data_sf_2023_OSO, oso23_new, by = "PT_ID")
# obesity_data_sf_2023_OSO = obesity_data_sf_2023_OSO_new #we can delete this line if necessary and just use obesity_data_sf_2023_oso_new
```

The data was further cleaned and transformed into a format that is ready for spatial data analysis. 

The key variable of interest is *Obesity Prevalence* (PREVALENCE) in schools: calculated as the ratio of Overweight or Severely Overweight (OSO) students in the most recent year (2023).

For our analysis of childhood obesity levels in Primary and Secondary Schools in Singapore, we will focus on the *2023 obesity levels* (obesity_data_sf_2023_OSO) as this is the most recent data available to us that would reflect the childhood obesity landscape today. This would also be the figure that HPB would look at when planning obesity policies. 

For our post-Covid analysis of childhood obesity levels in Singapore schools, the time series data consisting of obesity prevalence for all years will be used to present the effect that the COVID-19 pandemic and Circuit Breaker had on obesity levels (for Primary schools, due to intake data availability).

In addition to the prior data cleaning done by Professor Liu, we decided to remove "FUCHUN SECONDARY SCHOOL" as the enrolment data may be inaccurate due to school mergers. For our analysis for 2023, we remove "TANGLIN SECONDARY SCHOOL" and "GUANGYANG PRIMARY SCHOOL" as these schools have merged and no longer have overweight counts. Schools with no recorded obesity data in 2023 were also removed (only "TANJONG KATONG GIRLS' SCHOOL").

# Exploratory Data Analysis 

```{r}
tmap_mode("plot")
tm_shape(sg_ura_boundaries) + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.4, fill = "HMSP_Tier", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "HMSP Tiers"))

tm_shape(sg_ura_boundaries) + tm_fill() + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.4, fill = "ZONE", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "School zones"))

tm_shape(sg_ura_boundaries) + tm_fill() + tm_borders() +
  tm_shape(hmsp_data_sf) +
  tm_symbols(size = 0.3, fill = "CLUSTER", 
             fill.scale = tm_scale(values = "brewer.set1"), 
             fill.legend = tm_legend(title = "School clusters")) +
  tm_facets_wrap(by = "ZONE")
```

## OSO rate by school level

```{r}
ggplot(obesity_data_sf_2023_OSO, aes(x = LEVEL, y = PREVALENCE, fill = LEVEL)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of prevalence rate by School Level",
       x = "Level",
       y = "Prevalence Rate") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates to make it horizontal 
```

We observe that the median obesity prevalence rate for secondary schools is higher than primary schools. The inter-quartile range for obesity prevalence in secondary schools is considerably larger than that of primary schools, with a larger overall data spread as well. The sole outlier for secondary school obesity prevalence is Crest Secondary School, with an extremely high obesity prevalence of 0.268613.

## OSO rate by Zone

```{r}
ggplot(obesity_data_sf_2023_OSO, aes(x = ZONE, y = PREVALENCE, fill = ZONE)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of prevalence rate by Zone",
       x = "Zone",
       y = "Prevalence Rate") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates to make it horizontal 
```

We observe that across all four zones, the median obesity prevalence is comparable at between 10% to 15%. Ranking the median obesity prevalence for the zones, we get North \> West \> South \> East. The inter-quartile ranges and overall data spreads for all four zones are approximately equivalent, though the North zone is the only among the four for the data spread to stretch beyond an obesity prevalence of 0.20 (outliers exclusive). The two outliers for the East zone are East Spring Secondary School (0.193481) and Loyang View Secondary School (0.221809). It is worth noting that East Spring Secondary School being a data outlier can be attributed to the East zone having majority of its schools at lower obesity prevalence rates, for its obesity value would not be considered an outlier if placed in the North or West zones. The sole outlier for the West zone is Crest Secondary School with an extremely high obesity prevalence of 0.268613.

## OSO rate by School Type

```{r}
sub = obesity_data_sf_2023_OSO |>
  filter(type_code %in% c("GOVERNMENT SCHOOL", "GOVERNMENT-AIDED SCH"))
ggplot(sub, aes(x = type_code, y = PREVALENCE, fill = type_code)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of prevalence rate by School Type",
       x = "School Type",
       y = "Prevalence Rate") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates to make it horizontal 
```

We observe that the median obesity prevalence for government schools is over 2% higher than government-aided schools. The inter-quartile range for government-aided schools is larger, especially for Q3 which is spread over a larger range. The lower bound outliers for government schools are National Junior College (0.056235), Raffles Girls' School (Primary) (0.057807) and Raffles' Girls School (Secondary) (0.061249). The upper bound outliers for government schools are East Spring Secondary School (0.193481), Loyang View Secondary School (0.221809) and Crest Secondary School (0.268613).

## OSO rate by Gender Mix

```{r}
ggplot(obesity_data_sf_2023_OSO, aes(x = nature_code, y = PREVALENCE, fill = nature_code)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of prevalence rate by Gender Mix",
       x = "Gender Mix",
       y = "Prevalence Rate") +
  theme_minimal() +
  scale_x_discrete(drop = TRUE) +
  coord_flip() # Flip the coordinates to make it horizontal 
```

We observe that the median obesity prevalence for coed schools is noticeably higher than that for boys' and girls' schools. Median obesity prevalence in boys' schools is also higher than that for girls' schools. The interquartile range for all three types of schools is comparable, but the overall data range of obesity prevalence is much larger for coed schools than girls' or boys' schools. For girls' school, the sole upper bound data outlier is CHIJ St. Joseph's Convent (0.1552). We further observe that this is a North zone secondary school, but among the North zone girls' schools it is a significant outlier as other counterparts have obesity prevalence of around 0.05-0.06 range. For coed schools, the sole lower bound data outlier is Anglican High School (0.049065). The three upper bound data outliers are Loyang View Secondary School (0.221809), Pioneer Primary School (0.226236) and Crest Secondary School (0.268613). It is interesting that the opposite trend (higher obesity in single-gender schools than coed schools) has been observed in other countries such as [South Korea](https://pmc.ncbi.nlm.nih.gov/articles/PMC4430348/). Possible explanations include a stronger desire for students of both genders to manage their appearance (and in turn, weight) to increase their attractiveness to the opposite gender.

## OSO rate by Specialised Programs

```{r}
ggplot(obesity_data_sf_2023_OSO, aes(x = special, y = PREVALENCE, fill = special)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of prevalence rate by Specialised Programs",
       x = "Special Program",
       y = "Prevalence Rate") +
  theme_minimal() +
  scale_x_discrete(drop = TRUE) +
  coord_flip() # Flip the coordinates to make it horizontal 
```

We observe that the median obesity prevalence for non-specialised schools is noticeably higher than that for specialised schools, at almost 5 percentage points. The interquartile range for both types of schools is comparable, but the overall data range of obesity prevalence is much larger for non-specialised than specialised schools at both upper and lower ends of the spectrum. For specialised schools, the sole upper bound data outlier is Dunamn Secondary School (0.155722). It is worth noting that this obesity prevalence value would not be an outlier in the non-specialised school data and is fairly close to the Q3 range for non-specialised schools. For non-specialised schools, the sole lower bound data outlier is Anglican High School (0.049065). The three upper bound data outliers are Loyang View Secondary School (0.221809), Pioneer Primary School (0.226236) and Crest Secondary School (0.268613).

## OSO rate by HMSP Tier

```{r}
ggplot(obesity_data_sf_2023_OSO, aes(x = HMSP_Tier, y = PREVALENCE, fill = HMSP_Tier)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by HMSP Tier",
       x = "HMSP Tier",
       y = "OSO Rate") +
  theme_minimal() +
  scale_x_discrete(drop = TRUE) +
  coord_flip() # Flip the coordinates to make it horizontal 

cor(obesity_data_sf_2023_OSO$PREVALENCE, obesity_data_sf_2023_OSO$HMSP_Score)
```

We observe that the median obesity prevalence for the B, C, D and E HSMP tiers are within one percentage point of one another (0.12-0.13 range) while the median value of the A tier is much lower at approximately 0.1. The A tier also has a significantly smaller IQR range and total range than other tiers, showing greater consistency in obesity prvalence within HSMP A tier schools (though this is likely also because there are only 8 A tier schools). The other HSMP tiers have similar IQR ranges and total data ranges, though tier B has a much wider Q1 compared to tiers C, D and E despite having similar median values. For C tier, the two upper bound outliers are Loyang View Secondary School (0.221809) and Crest Secondary School (0.268613). For D tier, the sole outlier is Pioneer Primary School (0.226236). For A tier, the sole outlier is Admiralty Secondary School (0.16422), a coed secondary school in the North zone.

# Singapore maps

These are maps that define Singapore's regions based on URA defined boundaries (region, planning area, subzone) and MOE defined boundaries (zone, cluster). Subsequently, the defined regions can be used together with aggregated data so that we can examine overweight prevalence and other contributing factors by these differently delimited regions.

We will use these maps as a basemap for our subsequent analysis, and to facilitate our visualisation of obesity in Singapore schools

```{r}
tmap_mode("view")

# URA regions map
ura_regions <- sg_ura_boundaries %>% group_by(REGION_N) %>% 
  summarise(geometry = st_union(geometry))
ura_regions_map <- tm_shape(ura_regions, name = "URA regions") + tm_polygons(
    col = "grey", fill = "REGION_N", fill_alpha = 0.25, 
    fill.legend = tm_legend(title = "URA regions"))

# URA planning areas map
ura_planning_areas <- sg_ura_boundaries %>% group_by(PLN_AREA_N) %>% 
  summarise(geometry = st_union(geometry))
ura_planning_areas_map <- tm_shape(ura_planning_areas, name = "URA planning areas") + 
  tm_polygons(
    col = "grey", fill = "PLN_AREA_N", fill_alpha = 0.25, 
    fill.legend = tm_legend_hide())

# URA subzones map
ura_subzones <- sg_ura_boundaries %>% group_by(SUBZONE_N) %>% 
  summarise(geometry = st_union(geometry))
ura_subzones_map <- tm_shape(ura_subzones, name = "URA subzones") + tm_polygons(
  col = "grey", fill = "SUBZONE_N", fill_alpha = 0.25, 
  fill.legend = tm_legend_hide())

# MOE zones map
moe_zones <- obesity_data %>% group_by(ZONE) %>% 
  summarise(geometry = st_union(geometry)) %>% st_convex_hull()
moe_zones_map <- tm_shape(moe_zones, name = "MOE zones") + tm_polygons(
  col = "grey", fill = "ZONE", fill_alpha = 0.25, 
  fill.legend = tm_legend(title = "MOE School Zones"))

# MOE clusters map
moe_clusters <- obesity_data %>% group_by(CLUSTER) %>% 
  summarise(geometry = st_union(geometry)) %>% st_convex_hull()
moe_clusters_map <- tm_shape(moe_clusters, name = "MOE clusters") + tm_polygons(
  col = "grey", fill = "CLUSTER", fill_alpha = 0.25, 
  fill.legend = tm_legend_hide())
```

# Spatial overview of childhood obesity (Colin)

The following sections provide us with a spatial overview of childhood obesity in Singapore schools. We provide spatial descriptive summary measures, particularly spatial measures of central tendency (mean, median) and dispersion (standard deviation distance and standard deviation ellipse)

## Obesity prevalence overview map

The following map is an overview of overweight prevalence in Singapore schools. You may use the different layers provided in the map to view the individual school data against URA regions, planning areas and subzones, or MOE zones and clusters. Note that MOE zones and clusters have overlapping areas.

```{r}
tmap_mode("view")

# School obesity map plot
school_obesity_map <- tm_shape(obesity_data_sf_2023_OSO, name = "Overweight prevalence") +
  tm_symbols(
    size = 0.5,
    shape = "LEVEL",
    shape.legend = tm_legend(
      position = c("BOTTOM", "CENTER"),
      title = "Educational level"
      ),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape",
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    )

# Overview of overweight prevalence levels in Singapore schools
ura_regions_map + ura_planning_areas_map + ura_subzones_map + 
  moe_zones_map + moe_clusters_map +
  school_obesity_map +
  tm_title("Overview of overweight prevalence in Singapore schools")
```

## Obesity prevalence by educational level

Given the difference in obesity levels in primary schools versus secondary schools, we separately examine overweight prevalence in by educational level. Primary school and secondary school students may have different characteristics that might affect overweight prevalence.

The spatial mean centre (unweighted) and spatial variance (using the standard deviation distance and standard deviation ellipse) are calculated and plotted for each educational level.


```{r}
### by Primary School

tmap_mode("view")

# filter out by educational level and remove NA values to calculate SDE
obesity_pri_sch <- obesity_data_sf_2023_OSO %>% 
  filter(LEVEL == "PRIMARY", !is.na(PREVALENCE))

obesity_data_sf_2023_OSO %>% summary()

# Mean centre, standard deviation distance, standard deviation ellipse
pri_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_pri_sch))
pri_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_pri_sch))
pri_sch_mean_median <- rbind(pri_sch_mean$LOCATIONS, pri_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
pri_sch_mean_median$id = c("Spatial mean", "Spatial median")

pri_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_pri_sch))
pri_sch_sdd_line <- pri_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_pri_sch))
pri_sch_sde_line <- pri_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_pri_sch$PREVALENCE,
  points = st_coordinates(obesity_pri_sch))
pri_sch_sde_weighted_line <- pri_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_sch_map <- 
  tm_shape(obesity_pri_sch, name = "Overweight prevalence") +
  tm_symbols(
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    )

pri_sch_map_summ <- pri_sch_map +
  tm_shape(pri_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(pri_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(pri_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(pri_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Overweight prevalence in Primary Schools")

### by Secondary School

# filter out by educational level and remove NA values to calculate SDE
obesity_sec_sch <- obesity_data_sf_2023_OSO %>% 
  filter(LEVEL == "SECONDARY", !is.na(PREVALENCE))

# Mean centre, standard deviation distance, standard deviation ellipse
sec_sch_mean <- calc_mnc(id = 1, points = st_coordinates(obesity_sec_sch))
sec_sch_median <- calc_mdc(id = 2, points = st_coordinates(obesity_sec_sch))
sec_sch_mean_median <- rbind(sec_sch_mean$LOCATIONS, sec_sch_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
sec_sch_mean_median$id = c("Spatial mean", "Spatial median")

sec_sch_sdd <- calc_sdd(id = 3, points = st_coordinates(obesity_sec_sch))
sec_sch_sdd_line <- sec_sch_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde <- calc_sde(id = 4, points = st_coordinates(obesity_sec_sch))
sec_sch_sde_line <- sec_sch_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = obesity_sec_sch$PREVALENCE,
  points = st_coordinates(obesity_sec_sch))
sec_sch_sde_weighted_line <- sec_sch_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_sch_map <- 
  tm_shape(obesity_sec_sch, name = "Overweight prevalence") +
  tm_symbols(
    shape = 22,
    size = "PREVALENCE",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "PREVALENCE", 
    # a green-red diverging scale to indicate severity of overweight prevalence
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence by school",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    )

sec_sch_map_summ <- sec_sch_map +
  tm_shape(sec_sch_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(sec_sch_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(sec_sch_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(sec_sch_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Overweight prevalence in Secondary Schools")

tmap_arrange(pri_sch_map_summ, sec_sch_map_summ, ncol = 1)
```

For both primary and secondary schools, we observe that the spatial mean and median lie close to middle of Singapore, while the standard deviation distance spans the North-South length of Singapore. The standard deviation ellipse also generally follows the shape of Singapore: longer along its major axis and shorter along its minor axis, with a very small angle of rotation. This is what we would expect as schools are spread across Singapore evenly. This is with the exception of no schools in industrial zones, military areas or airports in the extreme East and West parts of Singapore, as well as no schools within the Bukit Timah Nature Reserve in the centre of Singapore. 

Interestingly, when while the standard deviational ellipse 


```{r}

```

Generally, we observe greater overweight prevalence levels for schools in the North and lower overweight prevalence levels in the South and Central regions of Singapore.

## Obesity prevalence by region

To facilitate urban planning, the Urban Redevelopment Authority (URA) divides Singapore into regions, planning areas and subzones.

The Planning Regions are divided into smaller Planning Areas. Each Planning Area have a population of about 150,000 and served by a town centre and several neighbourhood commercial/shopping centres. https://data.gov.sg/collections/2104/view

We aggregate overweight counts and school enrolment data by the different delimited regions, and fill in the regions to get visualise overweight prevalence by region.

```{r}
# Aggregating overweight counts by URA regions
obesity_ura_regions <- st_join(ura_regions, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(REGION_N, LEVEL) %>% 
  summarise(
    PREVALENCE_AGGREGATE = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_regions_map + obesity_ura_regions %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    # fill.scale = tm_scale_continuous(values = "brewer.yl_or_rd"), 
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      # orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA regions") +
  tm_facets_wrap(by = "LEVEL")

# Aggregating overweight counts by URA planning areas
obesity_ura_planning_areas <- st_join(ura_planning_areas, obesity_data_sf) %>%
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(PLN_AREA_N, LEVEL) %>% 
  summarise(
    PREVALENCE_AGGREGATE = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_planning_areas_map + obesity_ura_planning_areas %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA planning areas") +
  tm_facets_wrap(by = "LEVEL")

# Aggregating overweight counts by URA subzones
obesity_ura_subzones <- st_join(ura_subzones, obesity_data_sf) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(SUBZONE_N, LEVEL) %>% 
  summarise(
    PREVALENCE_AGGREGATE = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

ura_subzones_map + obesity_ura_subzones %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Overweight prevalence by URA subzones") +
  tm_facets_wrap(by = "LEVEL")

# Aggregating overweight counts by MOE school zones
obesity_moe_zones <- 
  # we left_join instead of st_join due to overlapping boundaries
  left_join(moe_zones, st_drop_geometry(obesity_data_sf)) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(ZONE, LEVEL) %>% 
  summarise(
    PREVALENCE_AGGREGATE = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

moe_zones_map +
  tm_legend_hide() + 
  obesity_moe_zones %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    fill_alpha = 0.95,
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1,
      show = T
      )
    ) +
  tm_title("Overweight prevalence by MOE clusters") +
  tm_facets_wrap(by = "LEVEL")

# Aggregating overweight counts by MOE clusters
obesity_moe_clusters <- 
  left_join(moe_clusters, st_drop_geometry(obesity_data_sf)) %>% 
  filter(YEAR == 2023, CATEGORY == "OSO") %>% 
  group_by(CLUSTER, LEVEL) %>% 
  summarise(
    PREVALENCE_AGGREGATE = sum(COUNT, na.rm = T) / sum(ENROLMENT, na.rm = T),
    geometry = st_union(geometry)
    )

moe_clusters_map + obesity_moe_clusters %>% 
  tm_shape(name = "Overweight prevalence") +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    fill_alpha = 0.95,
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1,
      show = T
      )
    ) +
  tm_title("Overweight prevalence by MOE clusters") +
  tm_facets_wrap(by = "LEVEL")
```

# Spatial clustering of obesity

Given the evidence of a spatial 

To consider spatial clustering of obesity in schools, we need to define how schools are connected spatially (which school is considered a neighbour to which school). 

We analysed spatial autocorrelation in administrative regions (URA regions, planning areas, subzones, MOE clusters), graph neighbours (Delauney triangulation, Sphere of influence), and nearest neighbours (nearest 1, 2 and 3) in the sections below.

## Region based analysis (Contiguous region neighbours)

There are too few contiguous neighbours by URA region to analyse meaningfully (only 5 URA regions). When analysing by URA subzone, there are too many non-contiguous subzones. We recommend to only analyse by URA planning area, which has some non-contiguous neighbours in the central regions. These non-contiguous neighbours by planning area are then manually edited (in the .gal file) to create neighbours with nearest planning areas to ensure contiguity for spatial statistical tests.

```{r}
nb_pval_pri <- data.frame("Neighbour object" = c(), "Moran.I.p.value" = c())
nb_pval_sec <- data.frame("Neighbour object" = c(), "Moran.I.p.value" = c())
```


### Contiguous neighbours by URA regions

```{r}
# by Primary school
obesity_ura_regions_pri <- obesity_ura_regions %>% filter(LEVEL == "PRIMARY")
ura_regions_nb_pri <- obesity_ura_regions_pri %>% poly2nb(queen = T)
ura_regions_nb_line_pri <- ura_regions_nb_pri %>% 
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_ura_regions_pri)))
pri_map <- ura_regions_map +
  ura_regions_nb_line_pri %>% tm_shape() + tm_lines(lwd = 2) + 
  st_point_on_surface(st_geometry(obesity_ura_regions_pri)) %>% tm_shape() + tm_symbols(size = 0.5)

# by Secondary school
obesity_ura_regions_sec <- obesity_ura_regions %>% filter(LEVEL == "SECONDARY")
ura_regions_nb_sec <- obesity_ura_regions_sec %>% poly2nb(queen = T)
ura_regions_nb_line_sec <- ura_regions_nb_sec %>% 
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_ura_regions_sec)))
sec_map <- ura_regions_map +
  ura_regions_nb_line_sec %>% tm_shape() + tm_lines(lwd = 2) + 
  st_point_on_surface(st_geometry(obesity_ura_regions_sec)) %>% tm_shape() + tm_symbols(size = 0.5)

tmap_arrange(pri_map, sec_map)
```

### Contiguous neighbours by URA planning areas

```{r}
# by Primary school
obesity_ura_planning_areas_pri <- obesity_ura_planning_areas %>%
  filter(LEVEL == "PRIMARY") %>%
  rowid_to_column("index")

# # Manually connecting 1 observation with no neighbours with its nearest neighbours
# ura_planning_areas_nb_pri <- obesity_ura_planning_areas_pri %>% poly2nb(queen = T)
# ura_planning_areas_nb_pri %>%
#   write.nb.gal(file = "ura_planning_areas_nb_pri.gal",
#                ind = ura_planning_areas_nb_pri$index)

ura_planning_areas_nb_pri <- read.gal("ura_planning_areas_nb_pri.gal")
ura_planning_areas_nb_line_pri <- ura_planning_areas_nb_pri %>% 
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_ura_planning_areas_pri)))
pri_map <- ura_planning_areas_map +
  obesity_ura_planning_areas_pri %>% tm_shape() +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      # orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  ura_planning_areas_nb_line_pri %>% tm_shape() + tm_lines(lwd = 2) +
  st_point_on_surface(st_geometry(obesity_ura_planning_areas_pri)) %>% 
  tm_shape() + tm_symbols(size = 0.5)

# by Secondary school
obesity_ura_planning_areas_sec <- obesity_ura_planning_areas %>% 
  filter(LEVEL == "SECONDARY")  %>%
  rowid_to_column("index")

# # Manually connecting 1 observation with no neighbours with its nearest neighbour
# ura_planning_areas_nb_sec <- obesity_ura_planning_areas_sec %>% poly2nb(queen = T)
# ura_planning_areas_nb_sec %>%
#   write.nb.gal(file = "ura_planning_areas_nb_sec.gal",
#                ind = ura_planning_areas_nb_sec$index)

ura_planning_areas_nb_sec <- read.gal("ura_planning_areas_nb_sec.gal")
ura_planning_areas_nb_line_sec <- ura_planning_areas_nb_sec %>% 
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_ura_planning_areas_sec)))
sec_map <- ura_planning_areas_map +
  obesity_ura_planning_areas_sec %>% tm_shape() +
  tm_polygons(
    fill = "PREVALENCE_AGGREGATE",
    fill.scale = tm_scale(
      breaks = c(0, 0.05, 0.10, 0.15, 0.20, 0.25, 0.30),
      values = "brewer.yl_or_rd"
      ), 
    fill.legend = tm_legend(
      title = "Overweight prevalence",
      # orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  ura_planning_areas_nb_line_sec %>% tm_shape() + tm_lines(lwd = 2) +
  st_point_on_surface(st_geometry(obesity_ura_planning_areas_sec)) %>% 
  tm_shape() + tm_symbols(size = 0.5)

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(ura_planning_areas_lw_pri <- nb2listw(ura_planning_areas_nb_pri, style = "W"))
ura_planning_areas_lw_B_pri <- nb2listw(ura_planning_areas_nb_pri, style = "B")
(moran <- moran.test(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, ura_planning_areas_lw_pri))
moran.plot(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, ura_planning_areas_lw_pri,
           main="Moran scatterplot")
geary.test(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, ura_planning_areas_lw_pri)
globalG.test(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE,
             ura_planning_areas_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "Planning Area", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(ura_planning_areas_lw_sec <- nb2listw(ura_planning_areas_nb_sec, style = "W"))
ura_planning_areas_lw_B_sec <- nb2listw(ura_planning_areas_nb_sec, style = "B")
(moran <- moran.test(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, ura_planning_areas_lw_sec))
moran.plot(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, ura_planning_areas_lw_sec, 
           main="Moran scatterplot")
geary.test(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, ura_planning_areas_lw_sec)
globalG.test(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE,
             ura_planning_areas_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "Planning Area", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 3.428571 \
p-value for Global Moran's I test (Primary school): 0.005857 \
p-value for Global Geary's C test (Primary school): 0.005738 \
p-value for Global Getis-Ord's G test (Primary school): 0.8278

Average number of links (Secondary school): 3.285714 \
p-value for Global Moran's I test (Secondary school): 0.5256 \
p-value for Global Geary's C test (Secondary school): 0.5418 \
p-value for Global Getis-Ord's G test (Secondary school): 0.8072

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_ura_planning_areas_pri$HIGH_PREVALENCE <- 
  as.factor(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_ura_planning_areas_pri %>% 
  tm_shape(name = "Overweight prevalence") + tm_polygons(fill = "HIGH_PREVALENCE") +
  tm_title("Join count analysis")
joincount.multi(obesity_ura_planning_areas_pri$HIGH_PREVALENCE, 
                ura_planning_areas_lw_B_pri) 

# by Secondary school
obesity_ura_planning_areas_sec$HIGH_PREVALENCE <- 
  as.factor(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_ura_planning_areas_sec %>% 
  tm_shape(name = "Overweight prevalence") + tm_polygons(fill = "HIGH_PREVALENCE") +
  tm_title("Join count analysis")
joincount.multi(obesity_ura_planning_areas_sec$HIGH_PREVALENCE, 
                ura_planning_areas_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_ura_planning_areas_pri <- 
  localmoran(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, 
             ura_planning_areas_lw_pri)
obesity_ura_planning_areas_pri$LMORANZ <- localm_ura_planning_areas_pri[,4]
pri_map <- tm_shape(obesity_ura_planning_areas_pri) + tm_polygons(fill = "LMORANZ") + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_ura_planning_areas_sec <- 
  localmoran(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, 
             ura_planning_areas_lw_sec)
obesity_ura_planning_areas_sec$LMORANZ <- localm_ura_planning_areas_sec[,4]
sec_map <- tm_shape(obesity_ura_planning_areas_sec) + tm_polygons(fill = "LMORANZ") + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_ura_planning_areas_pri$GO_STAT <-
  localG(obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, 
         ura_planning_areas_lw_B_pri)
pri_map <- tm_shape(obesity_ura_planning_areas_pri) + tm_polygons("GO_STAT") +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_ura_planning_areas_sec$GO_STAT <-
  localG(obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, 
         ura_planning_areas_lw_B_sec)
sec_map <- tm_shape(obesity_ura_planning_areas_sec) + tm_polygons("GO_STAT") +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram

```{r}
# by Primary school
correl_planning_areas_pri <- sp.correlogram(
  ura_planning_areas_nb_pri, obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, 
  order = 3, method = "corr", style = "W")
correlI_planning_areas_pri <- sp.correlogram(
  ura_planning_areas_nb_pri, obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE, 
  order = 3, method = "I", style = "W")
correlC_planning_areas_pri <- sp.correlogram(
  ura_planning_areas_nb_pri, obesity_ura_planning_areas_pri$PREVALENCE_AGGREGATE,
  order = 3, method = "C", style = "W")
par(mfrow=c(2,3))
plot(correl_planning_areas_pri, main="Contiguity lag orders: correlation")
plot(correlI_planning_areas_pri, main="Contiguity lag orders: Moran's I")
plot(correlC_planning_areas_pri, main="Contiguity lag orders: Geary's C")

# by Secondary school
correl_planning_areas_sec <- sp.correlogram(
  ura_planning_areas_nb_sec, obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, 
  order = 3, method = "corr", style = "W")
correlI_planning_areas_sec <- sp.correlogram(
  ura_planning_areas_nb_sec, obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE, 
  order = 3, method = "I", style = "W")
correlC_planning_areas_sec <- sp.correlogram(
  ura_planning_areas_nb_sec, obesity_ura_planning_areas_sec$PREVALENCE_AGGREGATE,
  order = 3, method = "C", style = "W")
plot(correl_planning_areas_sec, main="Contiguity lag orders: correlation")
plot(correlI_planning_areas_sec, main="Contiguity lag orders: Moran's I")
plot(correlC_planning_areas_sec, main="Contiguity lag orders: Geary's C")
par(mfrow=c(1,1))
```

### Contiguous neighbours by URA subzones

```{r}
obesity_ura_subzones_pri <- obesity_ura_subzones %>% filter(LEVEL == "PRIMARY")
ura_subzones_nb_pri <- obesity_ura_subzones_pri %>% poly2nb(queen = T)
ura_subzones_nb_line_pri <- ura_subzones_nb_pri %>% 
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_ura_subzones_pri)))
pri_map <- ura_subzones_map +
  ura_subzones_nb_line_pri %>% tm_shape() + tm_lines(lwd = 2) + 
  st_point_on_surface(st_geometry(obesity_ura_subzones_pri)) %>% tm_shape() + tm_symbols(size = 0.5)

obesity_ura_subzones_sec <- obesity_ura_subzones %>% filter(LEVEL == "SECONDARY")
ura_subzones_nb_sec <- obesity_ura_subzones_sec %>% poly2nb(queen = T)
ura_subzones_nb_line_sec <- ura_subzones_nb_sec %>% 
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_ura_subzones_sec)))
sec_map <- ura_subzones_map +
  ura_subzones_nb_line_sec %>% tm_shape() + tm_lines(lwd = 2) + 
  st_point_on_surface(st_geometry(obesity_ura_subzones_sec)) %>% tm_shape() + tm_symbols(size = 0.5)

tmap_arrange(pri_map, sec_map)
```

## Cluster based analysis (Social distance)

### Contiguous neighbours by MOE clusters

```{r}
# Function to write a .gal file in the specified format
write_gal <- function(data, filename) {
  fileConn <- file(filename, "w")
  # Write the number of participants
  writeLines(as.character(nrow(data)), fileConn)
  # Iterate through each participant
  for (i in seq_len(nrow(data))) {
    pt_id <- data$PT_ID[i]
    # Find neighbors (same CLUSTER, different PT_ID)
    neighbors <- data$PT_ID[data$CLUSTER == data$CLUSTER[i] & data$PT_ID != pt_id]
    # Write PT_ID followed by number of neighbors
    writeLines(paste(as.character(pt_id), length(neighbors)) , fileConn)
    # Write list of neighbors (if any)
    if (length(neighbors) > 0) {
      writeLines(paste(neighbors, collapse = " "), fileConn)
    }
  }
  close(fileConn)
}

# Write to a .gal file
moe_clusters_pri <- obesity_pri_sch %>%
  st_drop_geometry() %>%
  select(PT_ID, SCHOOL, CLUSTER)
write_gal(moe_clusters_pri, "moe_clusters_pri.gal")

moe_clusters_sec <- obesity_sec_sch %>%
  st_drop_geometry() %>%
  select(PT_ID, SCHOOL, CLUSTER)
write_gal(moe_clusters_sec, "moe_clusters_sec.gal")

obesity_moe_clusters_pri <- obesity_pri_sch %>% filter(LEVEL == "PRIMARY")
moe_clusters_nb_pri <-
  read.gal("moe_clusters_pri.gal", region.id = obesity_pri_sch$PT_ID)
moe_clusters_nb_line_pri <- moe_clusters_nb_pri %>%
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_moe_clusters_pri)))
pri_map <- moe_clusters_map +
  moe_clusters_nb_line_pri %>% tm_shape() + tm_lines(lwd = 0.5) +
  pri_sch_map


obesity_moe_clusters_sec <- obesity_sec_sch %>% filter(LEVEL == "SECONDARY")
moe_clusters_nb_sec <-
  read.gal("moe_clusters_sec.gal", region.id = obesity_sec_sch$PT_ID)
moe_clusters_nb_line_sec <- moe_clusters_nb_sec %>%
  nb2lines(coords = st_point_on_surface(st_geometry(obesity_moe_clusters_sec)))
sec_map <- moe_clusters_map +
  moe_clusters_nb_line_sec %>% tm_shape() + tm_lines(lwd = 0.5) +
  sec_sch_map

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(moe_clusters_lw_pri <- nb2listw(moe_clusters_nb_pri, style = "W"))
moe_clusters_lw_B_pri <- nb2listw(moe_clusters_nb_pri, style = "B")
(moran <- moran.test(obesity_pri_sch$PREVALENCE, moe_clusters_lw_pri))
moran.plot(obesity_pri_sch$PREVALENCE, moe_clusters_lw_pri, main="Moran scatterplot")
geary.test(obesity_pri_sch$PREVALENCE, moe_clusters_lw_pri)
globalG.test(obesity_pri_sch$PREVALENCE, moe_clusters_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "MOE clusters", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(moe_clusters_lw_sec <- nb2listw(moe_clusters_nb_sec, style = "W"))
moe_clusters_lw_B_sec <- nb2listw(moe_clusters_nb_sec, style = "B")
(moran <- moran.test(obesity_sec_sch$PREVALENCE, moe_clusters_lw_sec))
moran.plot(obesity_sec_sch$PREVALENCE, moe_clusters_lw_sec, main="Moran scatterplot")
geary.test(obesity_sec_sch$PREVALENCE, moe_clusters_lw_sec)
globalG.test(obesity_sec_sch$PREVALENCE, moe_clusters_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "MOE clusters", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 5.657459 \
p-value for Global Moran's I test (Primary school): 0.1608 \
p-value for Global Geary's C test (Primary school): 0.1608 \
p-value for Global Getis-Ord's G test (Primary school): 0.02355

Average number of links (Secondary school): 4.347222 \
p-value for Global Moran's I test (Secondary school): 0.1768 \
p-value for Global Geary's C test (Secondary school): 0.1768 \
p-value for Global Getis-Ord's G test (Secondary school): 0.8314

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_pri_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_pri_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_pri_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_pri_sch$HIGH_PREVALENCE, 
                moe_clusters_lw_B_pri) 

# by Secondary school
obesity_sec_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_sec_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_sec_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_sec_sch$HIGH_PREVALENCE, 
                moe_clusters_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_moe_clusters_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             moe_clusters_lw_pri)
obesity_pri_sch$LMORANZ <- localm_moe_clusters_pri[,4]
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_moe_clusters_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             moe_clusters_lw_sec)
obesity_sec_sch$LMORANZ <- localm_moe_clusters_sec[,4]
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_pri_sch$GO_STAT <-
  localG(obesity_pri_sch$PREVALENCE, 
         moe_clusters_lw_B_pri)
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_sec_sch$GO_STAT <-
  localG(obesity_sec_sch$PREVALENCE, 
         moe_clusters_lw_B_sec)
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram (does not work)

```{r}
# # by Primary school
# correl_moe_clusters_pri <- sp.correlogram(
#   moe_clusters_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_moe_clusters_pri <- sp.correlogram(
#   moe_clusters_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_moe_clusters_pri <- sp.correlogram(
#   moe_clusters_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# par(mfrow=c(2,3))
# plot(correl_moe_clusters_pri, main="Contiguity lag orders: correlation")
# plot(correlI_moe_clusters_pri, main="Contiguity lag orders: Moran's I")
# plot(correlC_moe_clusters_pri, main="Contiguity lag orders: Geary's C")
# 
# # by Secondary school
# correl_moe_clusters_sec <- sp.correlogram(
#   moe_clusters_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_moe_clusters_sec <- sp.correlogram(
#   moe_clusters_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_moe_clusters_sec <- sp.correlogram(
#   moe_clusters_nb_sec, obesity_sec_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# plot(correl_moe_clusters_sec, main="Contiguity lag orders: correlation")
# plot(correlI_moe_clusters_sec, main="Contiguity lag orders: Moran's I")
# plot(correlC_moe_clusters_sec, main="Contiguity lag orders: Geary's C")
# par(mfrow=c(1,1))
```

## Graph neighbours based analysis

### Delauney triangulation neighbours

```{r}
# by Primary schools
# obesity_pri_sch_jitter <- st_jitter(obesity_pri_sch) 
del_nb_pri <- tri2nb(st_coordinates(obesity_pri_sch), 
                     row.names = obesity_pri_sch$PT_ID) 
pri_map <- tm_shape(nb2lines(del_nb_pri, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Secondary schools
# obesity_sec_sch_jitter <- st_jitter(obesity_sec_sch) 
del_nb_sec <- tri2nb(st_coordinates(obesity_sec_sch), 
                     row.names = obesity_sec_sch$PT_ID) 
sec_map <- tm_shape(nb2lines(del_nb_sec, coords = st_geometry(obesity_sec_sch))) +
  tm_lines(lwd = 2) + sec_sch_map

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(del_lw_pri <- nb2listw(del_nb_pri, style = "W"))
del_lw_B_pri <- nb2listw(del_nb_pri, style = "B")
(moran <- moran.test(obesity_pri_sch$PREVALENCE, del_lw_pri))
moran.plot(obesity_pri_sch$PREVALENCE, del_lw_pri, main="Moran scatterplot")
geary.test(obesity_pri_sch$PREVALENCE, del_lw_pri)
globalG.test(obesity_pri_sch$PREVALENCE, del_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "Delauney triangles", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(del_lw_sec <- nb2listw(del_nb_sec, style = "W"))
del_lw_B_sec <- nb2listw(del_nb_sec, style = "B")
(moran <- moran.test(obesity_sec_sch$PREVALENCE, del_lw_sec))
moran.plot(obesity_sec_sch$PREVALENCE, del_lw_sec, main="Moran scatterplot")
geary.test(obesity_sec_sch$PREVALENCE, del_lw_sec)
globalG.test(obesity_sec_sch$PREVALENCE, del_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "Delauney triangles", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 5.823204 \
p-value for Global Moran's I test (Primary school): 1.809e-12 \
p-value for Global Geary's C test (Primary school): 3.885e-12 \
p-value for Global Getis-Ord's G test (Primary school): 0.03254

Average number of links (Secondary school): 5.805556 \
p-value for Global Moran's I test (Secondary school): 0.0001695 \
p-value for Global Geary's C test (Secondary school): 9.452e-05 \
p-value for Global Getis-Ord's G test (Secondary school): 0.2914

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_pri_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_pri_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_pri_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_pri_sch$HIGH_PREVALENCE, 
                del_lw_B_pri) 

# by Secondary school
obesity_sec_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_sec_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_sec_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_sec_sch$HIGH_PREVALENCE, 
                del_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_del_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             del_lw_pri)
obesity_pri_sch$LMORANZ <- localm_del_pri[,4]
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_del_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             del_lw_sec)
obesity_sec_sch$LMORANZ <- localm_del_sec[,4]
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_pri_sch$GO_STAT <-
  localG(obesity_pri_sch$PREVALENCE, 
         del_lw_B_pri)
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_sec_sch$GO_STAT <-
  localG(obesity_sec_sch$PREVALENCE, 
         del_lw_B_sec)
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram

```{r}
# by Primary school
correl_del_pri <- sp.correlogram(
  del_nb_pri, obesity_pri_sch$PREVALENCE, 
  order = 3, method = "corr", style = "W")
correlI_del_pri <- sp.correlogram(
  del_nb_pri, obesity_pri_sch$PREVALENCE, 
  order = 3, method = "I", style = "W")
correlC_del_pri <- sp.correlogram(
  del_nb_pri, obesity_pri_sch$PREVALENCE,
  order = 3, method = "C", style = "W")
par(mfrow=c(2,3))
plot(correl_del_pri, main="Contiguity lag orders: correlation")
plot(correlI_del_pri, main="Contiguity lag orders: Moran's I")
plot(correlC_del_pri, main="Contiguity lag orders: Geary's C")

# by Secondary school
correl_del_sec <- sp.correlogram(
  del_nb_sec, obesity_sec_sch$PREVALENCE, 
  order = 3, method = "corr", style = "W")
correlI_del_sec <- sp.correlogram(
  del_nb_sec, obesity_sec_sch$PREVALENCE, 
  order = 3, method = "I", style = "W")
correlC_del_sec <- sp.correlogram(
  del_nb_sec, obesity_sec_sch$PREVALENCE,
  order = 3, method = "C", style = "W")
plot(correl_del_sec, main="Contiguity lag orders: correlation")
plot(correlI_del_sec, main="Contiguity lag orders: Moran's I")
plot(correlC_del_sec, main="Contiguity lag orders: Geary's C")
par(mfrow=c(1,1))
```

### Sphere of influence neighbours

```{r}
# by Primary schools
if (require(dbscan, quietly=TRUE)) {
  soi_nb_pri <- graph2nb(soi.graph(del_nb_pri, st_geometry(obesity_pri_sch)))
} else {
  soi_nb = NULL
}
pri_map <- tm_shape(nb2lines(soi_nb_pri, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Secondary schools
if (require(dbscan, quietly=TRUE)) {
  soi_nb_sec <- graph2nb(soi.graph(del_nb_sec, st_geometry(obesity_sec_sch)))
} else {
  soi_nb = NULL
}
sec_map <- tm_shape(nb2lines(soi_nb_sec, coords = st_geometry(obesity_sec_sch))) +
  tm_lines(lwd = 2) + sec_sch_map

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(soi_lw_pri <- nb2listw(soi_nb_pri, style = "W"))
soi_lw_B_pri <- nb2listw(soi_nb_pri, style = "B")
(moran <- moran.test(obesity_pri_sch$PREVALENCE, soi_lw_pri))
moran.plot(obesity_pri_sch$PREVALENCE, soi_lw_pri, main="Moran scatterplot")
geary.test(obesity_pri_sch$PREVALENCE, soi_lw_pri)
globalG.test(obesity_pri_sch$PREVALENCE, soi_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "Sphere of Influence", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(soi_lw_sec <- nb2listw(soi_nb_sec, style = "W"))
soi_lw_B_sec <- nb2listw(soi_nb_sec, style = "B")
(moran <- moran.test(obesity_sec_sch$PREVALENCE, soi_lw_sec))
moran.plot(obesity_sec_sch$PREVALENCE, soi_lw_sec, main="Moran scatterplot")
geary.test(obesity_sec_sch$PREVALENCE, soi_lw_sec)
globalG.test(obesity_sec_sch$PREVALENCE, soi_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "Sphere of Influence", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 3.082873 \
p-value for Global Moran's I test (Primary school): 6.672e-07 \
p-value for Global Geary's C test (Primary school): 3.173e-07 \
p-value for Global Getis-Ord's G test (Primary school): 0.07193

Average number of links (Secondary school): 2.958333 \
p-value for Global Moran's I test (Secondary school): 0.0004617 \
p-value for Global Geary's C test (Secondary school): 0.0002675 \
p-value for Global Getis-Ord's G test (Secondary school): 0.3102

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_pri_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_pri_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_pri_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_pri_sch$HIGH_PREVALENCE, 
                soi_lw_B_pri) 

# by Secondary school
obesity_sec_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_sec_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_sec_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_sec_sch$HIGH_PREVALENCE, 
                soi_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_soi_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             soi_lw_pri)
obesity_pri_sch$LMORANZ <- localm_soi_pri[,4]
soi_pri_map <- 
  tm_shape(nb2lines(soi_nb_pri, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + 
  pri_sch_map +
  tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")
  
# by Secondary school
localm_soi_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             soi_lw_sec)
obesity_sec_sch$LMORANZ <- localm_soi_sec[,4]
soi_sec_map <- 
  tm_shape(nb2lines(soi_nb_sec, coords = st_geometry(obesity_sec_sch))) +
  tm_lines(lwd = 2) +
  sec_sch_map +
  tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(soi_pri_map, soi_sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_pri_sch$GO_STAT <-
  localG(obesity_pri_sch$PREVALENCE, 
         soi_lw_B_pri)
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_sec_sch$GO_STAT <-
  localG(obesity_sec_sch$PREVALENCE, 
         soi_lw_B_sec)
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram (does not work)

Some sub-graphs have 2 or less neighbours

```{r}
# # by Primary school
# correl_soi_pri <- sp.correlogram(
#   soi_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "corr", style = "W")
# correlI_soi_pri <- sp.correlogram(
#   soi_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "I", style = "W")
# correlC_soi_pri <- sp.correlogram(
#   soi_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# par(mfrow=c(2,3))
# plot(correl_soi_pri, main="Contiguity lag orders: correlation")
# plot(correlI_soi_pri, main="Contiguity lag orders: Moran's I")
# plot(correlC_soi_pri, main="Contiguity lag orders: Geary's C")
# 
# # by Secondary school
# correl_soi_sec <- sp.correlogram(
#   soi_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_soi_sec <- sp.correlogram(
#   soi_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_soi_sec <- sp.correlogram(
#   soi_nb_sec, obesity_sec_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# plot(correl_soi_sec, main="Contiguity lag orders: correlation")
# plot(correlI_soi_sec, main="Contiguity lag orders: Moran's I")
# plot(correlC_soi_sec, main="Contiguity lag orders: Geary's C")
# par(mfrow=c(1,1))
```

## K nearest neighbour based analysis

### 1 neighbour

```{r}
# by Primary schools
knn1_nb_pri <- obesity_pri_sch %>% 
  knearneigh(k = 1) %>% 
  knn2nb(row.names = obesity_pri_sch$PT_ID) # nearest neighbour
pri_map <- tm_shape(nb2lines(knn1_nb_pri, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Secondary schools
knn1_nb_sec <- obesity_sec_sch %>% 
  knearneigh(k = 1) %>% 
  knn2nb(row.names = obesity_sec_sch$PT_ID) # nearest neighbour
sec_map <- tm_shape(nb2lines(knn1_nb_sec, coords = st_geometry(obesity_sec_sch))) +
  tm_lines(lwd = 2) + sec_sch_map

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(knn1_lw_pri <- nb2listw(knn1_nb_pri, style = "W"))
knn1_lw_B_pri <- nb2listw(knn1_nb_pri, style = "B")
(moran <- moran.test(obesity_pri_sch$PREVALENCE, knn1_lw_pri))
moran.plot(obesity_pri_sch$PREVALENCE, knn1_lw_pri, main="Moran scatterplot")
geary.test(obesity_pri_sch$PREVALENCE, knn1_lw_pri)
globalG.test(obesity_pri_sch$PREVALENCE, knn1_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "1 nearest neighbour", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(knn1_lw_sec <- nb2listw(knn1_nb_sec, style = "W"))
knn1_lw_B_sec <- nb2listw(knn1_nb_sec, style = "B")
(moran <- moran.test(obesity_sec_sch$PREVALENCE, knn1_lw_sec))
moran.plot(obesity_sec_sch$PREVALENCE, knn1_lw_sec, main="Moran scatterplot")
geary.test(obesity_sec_sch$PREVALENCE, knn1_lw_sec)
globalG.test(obesity_sec_sch$PREVALENCE, knn1_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "1 nearest neighbour", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 1 \
p-value for Global Moran's I test (Primary school): 0.004193 \
p-value for Global Geary's C test (Primary school): 0.02331 \
p-value for Global Getis-Ord's G test (Primary school): 0.09593

Average number of links (Secondary school): 1 \
p-value for Global Moran's I test (Secondary school): 0.0004819 \
p-value for Global Geary's C test (Secondary school): 0.004126 \
p-value for Global Getis-Ord's G test (Secondary school): 0.234

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_pri_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_pri_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_pri_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_pri_sch$HIGH_PREVALENCE, 
                knn1_lw_B_pri) 

# by Secondary school
obesity_sec_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_sec_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_sec_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_sec_sch$HIGH_PREVALENCE, 
                knn1_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_knn1_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             knn1_lw_pri)
obesity_pri_sch$LMORANZ <- localm_knn1_pri[,4]
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_knn1_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             knn1_lw_sec)
obesity_sec_sch$LMORANZ <- localm_knn1_sec[,4]
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_pri_sch$GO_STAT <-
  localG(obesity_pri_sch$PREVALENCE, 
         knn1_lw_B_pri)
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_sec_sch$GO_STAT <-
  localG(obesity_sec_sch$PREVALENCE, 
         knn1_lw_B_sec)
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram (does not work)

Some sub-graphs have 2 or less neighbours

```{r}
# # by Primary school
# correl_knn1_pri <- sp.correlogram(
#   knn1_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_knn1_pri <- sp.correlogram(
#   knn1_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_knn1_pri <- sp.correlogram(
#   knn1_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# par(mfrow=c(2,3))
# plot(correl_knn1_pri, main="Contiguity lag orders: correlation")
# plot(correlI_knn1_pri, main="Contiguity lag orders: Moran's I")
# plot(correlC_knn1_pri, main="Contiguity lag orders: Geary's C")
# 
# # by Secondary school
# correl_knn1_sec <- sp.correlogram(
#   knn1_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_knn1_sec <- sp.correlogram(
#   knn1_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_knn1_sec <- sp.correlogram(
#   knn1_nb_sec, obesity_sec_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# plot(correl_knn1_sec, main="Contiguity lag orders: correlation")
# plot(correlI_knn1_sec, main="Contiguity lag orders: Moran's I")
# plot(correlC_knn1_sec, main="Contiguity lag orders: Geary's C")
# par(mfrow=c(1,1))
```

### 2 neighbours

```{r}
# by Primary schools
knn2_nb_pri <- obesity_pri_sch %>% 
  knearneigh(k = 2) %>% 
  knn2nb(row.names = obesity_pri_sch$PT_ID) # nearest neighbour
pri_map <- tm_shape(nb2lines(knn2_nb_pri, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Secondary schools
knn2_nb_sec <- obesity_sec_sch %>% 
  knearneigh(k = 2) %>% 
  knn2nb(row.names = obesity_sec_sch$PT_ID) # nearest neighbour
sec_map <- tm_shape(nb2lines(knn2_nb_sec, coords = st_geometry(obesity_sec_sch))) +
  tm_lines(lwd = 2) + sec_sch_map

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(knn2_lw_pri <- nb2listw(knn2_nb_pri, style = "W"))
knn2_lw_B_pri <- nb2listw(knn2_nb_pri, style = "B")
(moran <- moran.test(obesity_pri_sch$PREVALENCE, knn2_lw_pri))
moran.plot(obesity_pri_sch$PREVALENCE, knn2_lw_pri, main="Moran scatterplot")
geary.test(obesity_pri_sch$PREVALENCE, knn2_lw_pri)
globalG.test(obesity_pri_sch$PREVALENCE, knn2_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "2 nearest neighbours", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(knn2_lw_sec <- nb2listw(knn2_nb_sec, style = "W"))
knn2_lw_B_sec <- nb2listw(knn2_nb_sec, style = "B")
(moran <- moran.test(obesity_sec_sch$PREVALENCE, knn2_lw_sec))
moran.plot(obesity_sec_sch$PREVALENCE, knn2_lw_sec, main="Moran scatterplot")
geary.test(obesity_sec_sch$PREVALENCE, knn2_lw_sec)
globalG.test(obesity_sec_sch$PREVALENCE, knn2_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "2 nearest neighbours", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 2 \
p-value for Global Moran's I test (Primary school): 1.474e-05 \
p-value for Global Geary's C test (Primary school): 4.17e-05 \
p-value for Global Getis-Ord's G test (Primary school): 0.01137

Average number of links (Secondary school): 2 \
p-value for Global Moran's I test (Secondary school): 0.003471 \
p-value for Global Geary's C test (Secondary school): 0.03461 \
p-value for Global Getis-Ord's G test (Secondary school): 0.2781

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_pri_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_pri_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_pri_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_pri_sch$HIGH_PREVALENCE, 
                knn2_lw_B_pri) 

# by Secondary school
obesity_sec_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_sec_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_sec_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_sec_sch$HIGH_PREVALENCE, 
                knn2_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_knn2_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             knn2_lw_pri)
obesity_pri_sch$LMORANZ <- localm_knn2_pri[,4]
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_knn2_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             knn2_lw_sec)
obesity_sec_sch$LMORANZ <- localm_knn2_sec[,4]
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_pri_sch$GO_STAT <-
  localG(obesity_pri_sch$PREVALENCE, 
         knn2_lw_B_pri)
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_sec_sch$GO_STAT <-
  localG(obesity_sec_sch$PREVALENCE, 
         knn2_lw_B_sec)
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram (does not work)

Some sub-graphs have 2 or less neighbours

```{r}
# # by Primary school
# correl_knn2_pri <- sp.correlogram(
#   knn2_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_knn2_pri <- sp.correlogram(
#   knn2_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_knn2_pri <- sp.correlogram(
#   knn2_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# par(mfrow=c(2,3))
# plot(correl_knn2_pri, main="Contiguity lag orders: correlation")
# plot(correlI_knn2_pri, main="Contiguity lag orders: Moran's I")
# plot(correlC_knn2_pri, main="Contiguity lag orders: Geary's C")
# 
# # by Secondary school
# correl_knn2_sec <- sp.correlogram(
#   knn2_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_knn2_sec <- sp.correlogram(
#   knn2_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_knn2_sec <- sp.correlogram(
#   knn2_nb_sec, obesity_sec_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# plot(correl_knn2_sec, main="Contiguity lag orders: correlation")
# plot(correlI_knn2_sec, main="Contiguity lag orders: Moran's I")
# plot(correlC_knn2_sec, main="Contiguity lag orders: Geary's C")
# par(mfrow=c(1,1))
```

### 3 neighbours

```{r}
# by Primary schools
knn3_nb_pri <- obesity_pri_sch %>% 
  knearneigh(k = 3) %>% 
  knn2nb(row.names = obesity_pri_sch$PT_ID) # nearest neighbour
pri_map <- tm_shape(nb2lines(knn3_nb_pri, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Secondary schools
knn3_nb_sec <- obesity_sec_sch %>% 
  knearneigh(k = 3) %>% 
  knn2nb(row.names = obesity_sec_sch$PT_ID) # nearest neighbour
sec_map <- tm_shape(nb2lines(knn3_nb_sec, coords = st_geometry(obesity_sec_sch))) +
  tm_lines(lwd = 2) + sec_sch_map

tmap_arrange(pri_map, sec_map)
```

Spatial autocorrelation statistics

```{r}
# by Primary schools
(knn3_lw_pri <- nb2listw(knn3_nb_pri, style = "W"))
knn3_lw_B_pri <- nb2listw(knn3_nb_pri, style = "B")
(moran <- moran.test(obesity_pri_sch$PREVALENCE, knn3_lw_pri))
moran.plot(obesity_pri_sch$PREVALENCE, knn3_lw_pri, main="Moran scatterplot")
geary.test(obesity_pri_sch$PREVALENCE, knn3_lw_pri)
globalG.test(obesity_pri_sch$PREVALENCE, knn3_lw_B_pri)
nb_pval_pri <- rbind(nb_pval_pri, data.frame("Neighbour object" = "3 nearest neighbours", "Moran.I.p.value" = moran$p.value))

# by Secondary schools
(knn3_lw_sec <- nb2listw(knn3_nb_sec, style = "W"))
knn3_lw_B_sec <- nb2listw(knn3_nb_sec, style = "B")
(moran <- moran.test(obesity_sec_sch$PREVALENCE, knn3_lw_sec))
moran.plot(obesity_sec_sch$PREVALENCE, knn3_lw_sec, main="Moran scatterplot")
geary.test(obesity_sec_sch$PREVALENCE, knn3_lw_sec)
globalG.test(obesity_sec_sch$PREVALENCE, knn3_lw_B_sec)
nb_pval_sec <- rbind(nb_pval_sec, data.frame("Neighbour object" = "3 nearest neighbours", "Moran.I.p.value" = moran$p.value))
```

Average number of links (Primary school): 3 \
p-value for Global Moran's I test (Primary school): 7.811e-08 \
p-value for Global Geary's C test (Primary school): 9.495e-08 \
p-value for Global Getis-Ord's G test (Primary school): 0.00337

Average number of links (Secondary school): 3 \
p-value for Global Moran's I test (Secondary school): 0.000427 \
p-value for Global Geary's C test (Secondary school): 0.002858 \
p-value for Global Getis-Ord's G test (Secondary school): 0.3985

#### Join count analysis

```{r}
# Setting an overweight cutoff point
OVERWEIGHT_CUTOFF <- 0.14

# by Primary school
obesity_pri_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_pri_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
pri_map <- ura_regions_map + obesity_pri_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_pri_sch$HIGH_PREVALENCE, 
                knn3_lw_B_pri) 

# by Secondary school
obesity_sec_sch$HIGH_PREVALENCE <- 
  as.factor(obesity_sec_sch$PREVALENCE > OVERWEIGHT_CUTOFF)
sec_map <- ura_regions_map + obesity_sec_sch %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "HIGH_PREVALENCE", 
    
    size = "PREVALENCE", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")
joincount.multi(obesity_sec_sch$HIGH_PREVALENCE, 
                knn3_lw_B_sec)

tmap_arrange(pri_map, sec_map)
```

#### Local Moran's I statistic

```{r}
# by Primary school
localm_knn3_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             knn3_lw_pri)
obesity_pri_sch$LMORANZ <- localm_knn3_pri[,4]
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_knn3_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             knn3_lw_sec)
obesity_sec_sch$LMORANZ <- localm_knn3_sec[,4]
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)
```

#### Local Getis-Ord's G statistics

```{r}
# by Primary school
obesity_pri_sch$GO_STAT <-
  localG(obesity_pri_sch$PREVALENCE, 
         knn3_lw_B_pri)
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

# by Secondary school
obesity_sec_sch$GO_STAT <-
  localG(obesity_sec_sch$PREVALENCE, 
         knn3_lw_B_sec)
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "GO_STAT", size = 0.5) +
  tm_title("Local Getis-Ord's G statistics")

tmap_arrange(pri_map, sec_map)
```

#### Correlogram (does not work)

Some sub-graphs have 2 or less neighbours

```{r}
# # by Primary school
# correl_knn3_pri <- sp.correlogram(
#   knn3_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_knn3_pri <- sp.correlogram(
#   knn3_nb_pri, obesity_pri_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_knn3_pri <- sp.correlogram(
#   knn3_nb_pri, obesity_pri_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# par(mfrow=c(2,3))
# plot(correl_knn3_pri, main="Contiguity lag orders: correlation")
# plot(correlI_knn3_pri, main="Contiguity lag orders: Moran's I")
# plot(correlC_knn3_pri, main="Contiguity lag orders: Geary's C")
# 
# # by Secondary school
# correl_knn3_sec <- sp.correlogram(
#   knn3_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "corr", style = "W")
# correlI_knn3_sec <- sp.correlogram(
#   knn3_nb_sec, obesity_sec_sch$PREVALENCE, 
#   order = 3, method = "I", style = "W")
# correlC_knn3_sec <- sp.correlogram(
#   knn3_nb_sec, obesity_sec_sch$PREVALENCE,
#   order = 3, method = "C", style = "W")
# plot(correl_knn3_sec, main="Contiguity lag orders: correlation")
# plot(correlI_knn3_sec, main="Contiguity lag orders: Moran's I")
# plot(correlC_knn3_sec, main="Contiguity lag orders: Geary's C")
# par(mfrow=c(1,1))
```

## Conclusions from analysis of spatial clustering of obesity

```{r}
nb_pval_pri
nb_pval_sec
```

Based on the Moran's I statistics, Delauney triangulation neigbours reported the lowest p-values, suggesting that there was significant spatial clustering following Delauney triangulation clusters.

However, due to the arbitrary neighbour links introduced by Delauney triangulation (e.g. schools along the borders of Singapore were indicated as neighbours even when they were very far apart), we determined that Sphere of Influence was the better choice to model neighbour links instead.

```{r}
# by Primary school
localm_soi_pri <- 
  localmoran(obesity_pri_sch$PREVALENCE, 
             soi_lw_pri)
obesity_pri_sch$LMORANZ <- localm_soi_pri[,4]
pri_map <- tm_shape(obesity_pri_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

# by Secondary school
localm_soi_sec <- 
  localmoran(obesity_sec_sch$PREVALENCE, 
             soi_lw_sec)
obesity_sec_sch$LMORANZ <- localm_soi_sec[,4]
sec_map <- tm_shape(obesity_sec_sch) + tm_symbols(fill = "LMORANZ", size = 0.5) + 
  tm_title("Local Moran's I Z-Statistic")

tmap_arrange(pri_map, sec_map)

# Identified clusters in Primary schools 
obesity_pri_sch %>%
  st_drop_geometry() %>% 
  filter(LMORANZ >= 3)

# Identified clusters in Secondary schools
obesity_sec_sch %>% 
  st_drop_geometry() %>%
  filter(LMORANZ >= 3) 
```

# Elite metrics (Lucas)

We obtain a Yes/No binary metric based on the popularity of primary & secondary schools in Singapore, which we dub as the "elite" metric. We derive this using the following methodology: Primary schools: We take the Phase 2C balloting information from the 2024 Primary 1 Registration Exercise for all 180 primary schools (database link: <https://www.moe.gov.sg/primary/p1-registration/past-vacancies-and-balloting-data/>). If the number of applicants in Phase 2C is greater than the number of vacancies, the school is marked as popular (Yes) reflecting oversubscribing due to high demand for the school's coveted vacancies. If the number of applicants in Phase 2C is less than or equal to the number of vacancies, the school is marked as not popular (No). Based on this methodology, there are 82 popular schools (Yes) and 98 non-popular schools (No). Secondary schools: We obtain the 2024 secondary school PSLE cutoff points from SG Schooling (<https://sgschooling.com/secondary/cop/all.html>) in the IP/PG3 (Express) streams. We take the cutoff point for each secondary school if given as a single value, or the highest value in a given point range, ignoring all secondary criteria for the Mother Tongue Grade (D/M/P) to get data for 138 secondary schools. We then calculate the top quartile of cutoff points, which ends up as 13, and label all secondary schools with cutoff points less than or equal to this value as popular schools (Y) and all other schools as non-popular (N). Using this methodology, we obtain 35 popular schools (Y) and 103 non-popular schools (N). For the secondary schools with no given PSLE cutoff point for 2024, we conduct the following manual adjustments: - School of the Arts, NUS High School and School of Science and Technology: These schools only accept admissions by Direct School Admission, so there is no PSLE cutoff range. From our research, these three schools are prestigious in their unique domains with many applicants each year. As such, we labelled these three schools as popular (Y). - Crest Secondary School, Assumption Pathway School and Northlight School: These schools offer specialised pathways for experiential and hands-on learning, allowing students to graduate with an ITE Skills certificate. As this is a non-mainstream pathway with only PG1/PG2 streams, we labelled these three schools as non-popular (N). In total, we have 144 secondary schools with 38 as popular (Y) and 106 as non-popular (N).

```{r}
pri_rankings <- read.csv("pri rankings.csv") %>% 
  mutate(School = toupper(School)) %>% 
  rename(is_popular_sch = Balloting.2C) %>% 
  select(School, is_popular_sch)
sec_rankings <- read.csv("sec rankings.csv") %>% 
  mutate(School = toupper(School)) %>% 
  rename(is_popular_sch = Quartile.1) %>% 
  select(School, is_popular_sch)
# obesity_data <- obesity_data %>% 
#   
# obesity_data <- obesity_data %>% 
#   left_join(sec_rankings, by = join_by(SCHOOL == School))

school_rankings <- rbind(pri_rankings, sec_rankings)
```

# Education and Income Census data
Next, we consider the Singapore Census of Population

2020 data and bring in two variables: Income: From Table 110 Resident Households by Planning Area of Residence and Monthly Household Income from Work, we sum up the income for each planning area by multiplying the midpoint of a given income range (for example, \$9,000 - \$9,999 would have a midpoint of \$9,500) with the number of residents with that given monthly income range and totalling everything. The only exception to this is the income bracket "\$20,000 and Over", of which we took the midpoint value as \$25,000. We then obtain the mean income per planning area by dividing this totalled income figure by the total number of residents in each planning area (given in the second column of the data table). Education: From Table 100 Resident Population Aged 15 Years and Over by Planning Area of Residence and Highest Qualification Attained, we sum up the number of years of education in each planning area by multiplying the minimum number of years of education completed (No Qualification = 0, Primary = 6, Lower Secondary = 8 [6+2], Secondary = 10 [6+2+2], Post-Secondary (Non-Tertiary) = 12 [6+2+2+2 assuming JC route], Polytechnic Diploma = 13 [6+2+2+3 assuming average Polytechnic diploma as 3 year programme], Professional Qualification and Other Diploma = 15 [6+2+2+2+3 assuming JC then 3 year other diploma], University =16 [6+2+2+2+4 assuming JC then 4 year university degree]) with the number of residents that fall within the relevant maximum education completion category. We then obtain the mean number of years or education per planning area by dividing this totalled education years figure by the total number of residents in each planning area (given in the second column of the data table).

## Preliminary Data Manipulation and Visuals (Lucas)

```{r}
# missing MUSEUM & NEWTON planning areas, so we will fill in with OTHERS data
census_add <- data.frame(
  PLN_AREA_N = c("MUSEUM", "NEWTON"),
  Total = c(12233, 12233),
  Income = c(12376, 12376),
  Education = c(12.58, 12.58),
  Primary = c(1693, 1693), 
  Secondary = c(1125, 1125))

census <- read.csv("census stats.csv") %>% 
  mutate(PLN_AREA_N = toupper(PLN_AREA_N)) %>% 
  bind_rows(census_add)

census_merge <- ura_planning_areas %>%
  left_join(census) %>%
  drop_na("Income")

ura_planning_areas_map + tm_shape(census_merge) + 
  tm_polygons(fill="Income", 
              fill.scale = tm_scale(breaks = c(6925,8396,9532,11317,16197))
              ) +
  tm_title("Breakdown by Household Income")

ura_planning_areas_map + tm_shape(census_merge) + 
  tm_polygons(fill="Education", 
              fill.scale = tm_scale(breaks = c(9.76,10.74,11.25,11.89,14.78))
              ) +
  tm_title("Breakdown by Education Level")

ura_planning_areas_map + tm_shape(census_merge) + 
  tm_polygons(fill="Primary", 
              fill.scale = tm_scale(breaks = c(61,4778,7043,11681,21822))
              ) +
  tm_title("Breakdown by Pri Students")

ura_planning_areas_map + tm_shape(census_merge) + 
  tm_polygons(fill="Secondary", 
              fill.scale = tm_scale(breaks=c(59,3526,4894,8804,13221))
              ) +
  tm_title("Breakdown by Sec Students")
```

For the income variable, the planning areas with higher incomes are situated in the southern and central regions. The northern and western planning areas appear to have lower incomes while the eastern planning areas have moderately high incomes. For the education variable, planning areas with residents that have more years of education are situated in the southern and central regions. The northern and western planning areas appear to have fewer years of education while the eastern planning areas moderate to high number of years of education.

## Spatial Descriptive Summary Measures of Income (Lucas)

```{r}
census_coords <- st_coordinates(st_point_on_surface(census_merge))
income_mean <- calc_mnc(id = 1, points = census_coords)
income_median <- calc_mdc(id = 2, points = census_coords)
income_mean_median <- rbind(income_mean$LOCATIONS, income_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
income_mean_median$id = c("Weighted mean", "Weighted median")

income_sdd <- calc_sdd(id = 3, points = census_coords)
income_sdd_line <- income_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

income_sde <- calc_sde(id = 4, points = census_coords)
income_sde_line <- income_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

income_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = census_merge$Income,
  points = census_coords)
income_sde_weighted_line <- income_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

income_map <- ura_planning_areas_map +
    tm_shape(census_merge, name = "Income") +
    tm_symbols(
        size = "Income",
        size.scale = tm_scale_continuous(values.scale = 0.8),
        size.legend = tm_legend_hide(),
        fill = "Income", 
        # a green-red diverging scale to indicate severity of overweight prevalence
        fill.scale = tm_scale_continuous(values = "brewer.yl_or_rd"), 
        fill.legend = tm_legend(
            title = "Income Distribution",
            orientation = "landscape", 
            position = c("BOTTOM", "LEFT"),
            item.height = 0.5, item.width = 2.1
        )
    ) + 
    tm_shape(income_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
    tm_shape(income_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
    tm_shape(income_sde_weighted_line, name = "weighted SDE") + 
    tm_lines(col = "red", lwd = 2, lty = 3) +
    tm_shape(income_mean_median, name = "Summary measures") + tm_dots(
        size = 0.5, shape = "id", fill = "white", 
        shape.legend = tm_legend("Summary measures")
    ) + 
    tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
    tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
    tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
    tm_title("Income Distribution across SG Planning Areas")

print(income_map)
```

Analysing the spatial descriptive summary measures for income variable, we observe all spatial measures (spatial mean, spatial median, SDD, SDE and weighted SDE) are directionally skewed towards the southeast, though much more skewed towards the south than the eastern direction. This matches the observation that most of the higher income planning areas are concentrated in the south, with middle range income planning areas in the east. The weighted SDE (red dotted line) being more skewed towards the south than the SDE (black dotted line) further reflects this spatial pattern.

## Spatial Descriptive Summary Measures of Education Years (Lucas)

```{r}
education_mean <- calc_mnc(id = 1, points = census_coords, weighted = T, weights = census_merge$Education)
education_median <- calc_mdc(id = 2, points = census_coords)
education_mean_median <- rbind(education_mean$LOCATIONS, education_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
education_mean_median$id = c("Weighted mean", "Weighted median")

education_sdd <- calc_sdd(id = 3, points = census_coords)
education_sdd_line <- education_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

education_sde <- calc_sde(id = 4, points = census_coords)
education_sde_line <- education_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

education_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = census_merge$Education,
  points = census_coords)
education_sde_weighted_line <- education_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

education_map <- ura_planning_areas_map +
    tm_shape(census_merge, name = "Education") +
    tm_symbols(
        size = "Education",
        size.scale = tm_scale_continuous(values.scale = 0.8),
        size.legend = tm_legend_hide(),
        fill = "Education", 
        # a green-red diverging scale to indicate severity of overweight prevalence
        fill.scale = tm_scale_continuous(values = "brewer.yl_or_rd"), 
        fill.legend = tm_legend(
            title = "Education Years",
            orientation = "landscape", 
            position = c("BOTTOM", "LEFT"),
            item.height = 0.5, item.width = 2.1
        )
    ) + 
    tm_shape(education_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
    tm_shape(education_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
    tm_shape(education_sde_weighted_line, name = "weighted SDE") + 
    tm_lines(col = "red", lwd = 2, lty = 3) +
    tm_shape(education_mean_median, name = "Summary measures") + tm_dots(
        size = 0.5, shape = "id", fill = "white", 
        shape.legend = tm_legend("Summary measures")
    ) + 
    tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
    tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
    tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
    tm_title("Education Years Across SG Planning Areas")

print(education_map)
```

Analysing the spatial descriptive summary measures for education variable, we observe all spatial measures (spatial mean, spatial median, SDD, SDE and weighted SDE) are directionally skewed towards the southeast, though much more skewed towards the south than the eastern direction. Visually comparing with the spatial descriptive measures for the income variable, the spatial mean and median are in approximately the same location (within the same planning area at least) while the SDE and weighted SDE for the education variable are slightly less skewed towards the south than the income variable. This could be because of the nature of the variable since income ranges (max-min = \$8,000) have a much larger spread than education years (max - min = 4 years). This matches the observation that most of the planning areas with best educated residents are concentrated in the south, with planning areas in the east having more years of education than planning areas in the west. The weighted SDE (red dotted line) being more skewed towards the south than the SDE (black dotted line) further reflects this spatial pattern, although there appears to be less of a southbound skewing effect for the education variable than the income variable.

# Food and Nutrition (Nina)

```{r}
#Data transformation

#Hawker Centres
hawker_full = st_read("NEAMarketandFoodCentre.geojson") %>%
  st_transform(crs = sg_crs)
hawker = hawker_full["geometry"]

#Fast Food Outlets 
  #Mcdonald's
macs_full = st_read("mcdonalds_singapore.geojson") %>%
  st_transform(crs = sg_crs) 
table(st_is_valid(macs_full)) #Check validity
st_crs(macs_full)$epsg #Check CRS
macs = macs_full[,"geometry"] #Extract necessary columns for analysis

  #7-11 Outlets
seven_eleven_full = st_read("711_singapore.geojson") %>%
  st_transform(crs = sg_crs)
seven_eleven = seven_eleven_full[,"geometry"] #Extract necessary columns for analysis
table(st_is_valid(seven_eleven)) #Check validity
st_crs(seven_eleven)$epsg #Check CRS

#Combine data into single SF object
seven_eleven_sf = mutate(seven_eleven, outlet = "7-11")
macs_sf = mutate(macs, outlet = "mcdonalds")
hawker_sf = mutate(hawker, outlet = "hawker")
food = rbind(seven_eleven_sf, macs_sf, hawker_sf)


##### Primary school analysis
obesity_pri_food = obesity_pri_sch #duplicate primary school data
st_crs(obesity_pri_food)$units #check units
pri_schools_buffer1 <- st_buffer(obesity_pri_food, dist = 1000) %>% st_geometry()
pri_intersections <- st_intersects(pri_schools_buffer1, food)

#Calculate the number of food outlets within a 1km radius of each school
obesity_pri_food$num_food_outlets <- lengths(pri_intersections)
obesity_pri_food

#Show density via choropleth map
pri_food_map = ura_regions_map + 
  tm_shape(obesity_pri_food) +
  tm_symbols(
    size = 0.5,
    fill = "num_food_outlets"
  ) +
  tm_title("Food Outlet Density by Primary School")

print(pri_food_map)


####Secondary school analysis
obesity_sec_food = obesity_sec_sch #duplicate secondary school data
st_crs(obesity_sec_food)$units #check units
sec_schools_buffer1 <- st_buffer(obesity_sec_food, dist = 1000) %>% st_geometry()
sec_intersections <- st_intersects(sec_schools_buffer1, food)

#calculate the number of food outlets within a 1km radius of each school
obesity_sec_food$num_food_outlets <- lengths(sec_intersections)
obesity_sec_food

#show density via choropleth map
sec_food_map = ura_regions_map + 
  tm_shape(obesity_sec_food) +
  tm_symbols(
    size = 0.5,
    fill = "num_food_outlets"
  ) +
  tm_title("Food Outlet Density by Secondary School")

print(sec_food_map)

###Spatial analysis - Secondary school
st_crs(obesity_sec_food)$proj #Check for metric units
obesity_sec_food_coords = st_coordinates(obesity_sec_food)

mean_sec_food = calc_mnc(id = 1, points = obesity_sec_food_coords)
median_sec_food = calc_mdc(id = 2, points = obesity_sec_food_coords)
sdd_sec_food = calc_sdd(id = 3, calccentre = T, weighted = F, points = obesity_sec_food_coords)

LC_sdd_line_sec_food = sdd_sec_food$LOCATIONS %>% 
    st_as_sf(coords = c("x", "y"), crs=st_crs(obesity_sec_food)) %>% 
    st_combine() %>% 
    st_cast("LINESTRING") # CRS from the parent object.

sde_sec_food = calc_sde(id=4, calccentre = T, weighted=F, points= obesity_sec_food_coords)
ls()
LC_sde_line_sec = sde_sec_food$LOCATIONS %>%
    st_as_sf(coords = c("x", "y"), crs=st_crs(obesity_sec_food)) %>% 
    st_combine() %>% 
    st_cast("LINESTRING") # CRS from the parent object.

(LC_sdd_line_sec = sdd_sec_food$LOCATIONS %>% 
    st_as_sf(coords = c("x", "y"), crs=st_crs(obesity_sec_food)) %>% 
    st_combine() %>% 
    st_cast("LINESTRING")) # CRS from the parent object.

(LC_std_dist_sec = st_sf(sdd_sec_food$ATTRIBUTES, geom=st_geometry(LC_sdd_line_sec))) 

LC_std_ellps_sec = st_sf(sde_sec_food$ATTRIBUTES, geom = st_geometry(LC_sde_line_sec))

#Plot the summary measures
(mean_median_sec = rbind(mean_sec_food$LOCATIONS, median_sec_food$LOCATIONS) %>% 
    data.frame() %>% 
    st_as_sf(coords=c("x","y"), crs=st_crs(obesity_sec_food))) # matrix -> data.frame -> sf points

mean_median_sec$id =  c("Spatial Mean","Spatial Median") # assign names
mean_median_sec

sec_food_map + 
  tm_shape(obesity_sec_food) +
  tm_shape(LC_std_ellps_sec) + tm_lines(lwd=2) + 
  tm_shape(LC_std_dist_sec) + tm_lines(lwd=2, lty=2) + 
  tm_shape(mean_median_sec) + tm_dots(size=0.5, shape="id", fill = "white", shape.legend = tm_legend_hide()) + 
  tm_compass(position = c("left", "top")) +
  tm_add_legend(type="symbols", labels="Spatial mean", shape = 1) +
  tm_add_legend(type="symbols", labels="Spatial median", shape = 0) +
  tm_add_legend(type="lines", labels="SDD", lty=2) + 
  tm_add_legend(type="lines", labels="SDE", lty=1)


##Spatial analysis - Primary school
st_crs(obesity_pri_food)$proj #Check for metric units 
obesity_pri_food_coords = st_coordinates(obesity_pri_food)

mean_pri_food = calc_mnc(id = 1, points = obesity_pri_food_coords)
median_pri_food = calc_mdc(id = 2, points = obesity_pri_food_coords)
sdd_pri_food = calc_sdd(id = 3, calccentre = T, weighted = F, points = obesity_pri_food_coords)

(LC_sdd_line_pri_food = sdd_pri_food$LOCATIONS %>% 
    st_as_sf(coords = c("x", "y"), crs=st_crs(obesity_pri_food)) %>% 
    st_combine() %>% 
    st_cast("LINESTRING")) # CRS from the parent object.

sde_pri_food = calc_sde(id=4, calccentre = T, weighted=F, points= obesity_pri_food_coords)
ls()
LC_sde_line_pri = sde_pri_food$LOCATIONS %>%
    st_as_sf(coords = c("x", "y"), crs=st_crs(obesity_pri_food)) %>% 
    st_combine() %>% 
    st_cast("LINESTRING") # CRS from the parent object.


(LC_sdd_line_pri = sdd_pri_food$LOCATIONS %>% 
    st_as_sf(coords = c("x", "y"), crs=st_crs(obesity_pri_food)) %>% 
    st_combine() %>% 
    st_cast("LINESTRING")) # CRS from the parent object.

(LC_std_dist_pri = st_sf(sdd_pri_food$ATTRIBUTES, geom=st_geometry(LC_sdd_line_pri))) 

LC_std_ellps_pri = st_sf(sde_pri_food$ATTRIBUTES, geom = st_geometry(LC_sde_line_pri))

#Plot the summary measures 
(mean_median_pri = rbind(mean_pri_food$LOCATIONS, median_pri_food$LOCATIONS) %>% 
    data.frame() %>% 
    st_as_sf(coords=c("x","y"), crs=st_crs(obesity_pri_food))) # matrix -> data.frame -> sf points

mean_median_pri$id =  c("Spatial Mean","Spatial Median") # assign names
mean_median_pri

pri_food_map + 
  tm_shape(obesity_pri_food) +
  tm_shape(LC_std_ellps_pri) + tm_lines(lwd=2) + 
  tm_shape(LC_std_dist_pri) + tm_lines(lwd=2, lty=2) + 
  tm_shape(mean_median_pri) + tm_dots(size=0.5, shape="id", fill = "white", shape.legend = tm_legend_hide()) + 
  tm_compass(position = c("left", "top")) +
  tm_add_legend(type="symbols", labels="Spatial mean", shape = 1) +
  tm_add_legend(type="symbols", labels="Spatial median", shape = 0) +
  tm_add_legend(type="lines", labels="SDD", lty=2) + 
  tm_add_legend(type="lines", labels="SDE", lty=1)
#Bivariate Moran's I
obesity_pri_food
coords_pri_sch <- st_coordinates(obesity_pri_food)
knn_nb_pri_sch_food <- knn2nb(knearneigh(coords_pri_sch, k = 3))
wts_pri_sch_food <- nb2listw(knn_nb_pri_sch_food, style = "W")
morans_bi_pri_food <- lm.morantest(lm(PREVALENCE ~ num_food_outlets, data = obesity_pri_food), listw = wts_pri_sch_food)
morans_bi_pri_food  #P-value = 4.071e-08, hence there is a significant relationship between obesity prevalence and number of food outlets


obesity_sec_food
coords_sec_sch <- st_coordinates(obesity_sec_food)
knn_nb_sec_sch_food <- knn2nb(knearneigh(coords_sec_sch, k = 3))
wts_sec_sch_food <- nb2listw(knn_nb_sec_sch_food, style = "W")
# morans_bi_sec_food <- lm.morantest(lm(PREVALENCE ~ num_food_outlets, data = obesity_sec_food), listw = wts_sec_sch_food)
# morans_bi_sec_food  #P-value = 0.0003288, hence there is a significant relationship between obesity prevalence and number of food outlets at a 5% sig level
```

# Distance of the nearest food option to each school

```{r}
#Write function to compute distance to nearest food option, along with outlet_type
library(geosphere)

find_nearest_food_options <- function(schools_sf, food_options_sf) {

  if (st_crs(schools_sf) != st_crs(food_options_sf)) {
    stop("CRS of schools_sf and food_options_sf must be the same.")
  }

  distances <- st_distance(schools_sf, food_options_sf)

  nearest_indices <- apply(distances, 1, which.min)

  nearest_distances <- apply(distances, 1, min)

  nearest_outlet_types <- food_options_sf$outlet[nearest_indices]

    result_sf <- schools_sf %>%
    mutate(
      distance_to_nearest_food_option = as.numeric(nearest_distances),
      outlet_type = nearest_outlet_types
    ) %>%
    select(SCHOOL, distance_to_nearest_food_option, outlet_type, geometry)
  
  print(result_sf)
}


nearest_outlet_distance = find_nearest_food_options(obesity_data_sf_2023_OSO, food)
nearest_outlet_distance

#Combine data into original sf object
food_combined_sf = obesity_data_sf_2023_OSO %>%
  st_join(nearest_outlet_distance[,c(2,3)], by = "SCHOOL")

#Plot graph 
ura_regions_map +
  tm_shape(food_combined_sf) +
  tm_symbols(size = 0.5, fill = "distance_to_nearest_food_option")


food_combined_sf
coords_dist <- st_coordinates(food_combined_sf)
knn_nb_sec_sch_food <- knn2nb(knearneigh(coords_dist, k = 3))
wts_sec_sch_food <- nb2listw(knn_nb_sec_sch_food, style = "W")
# morans_bi_sec_food <- lm.morantest(lm(PREVALENCE ~ num_food_outlets, data = obesity_sec_food), listw = wts_sec_sch_food)
# morans_bi_sec_food  #P-value = 0.0003288, hence there is a significant relationship between obesity prevalence and number of food outlets at a 5% sig level
```

# Sports CCAs in Schools (Qasim)

This section aims to explore the spatial distribution of physical sports participation in schools across Singapore. We will be using the 2023 School Obesity (OSO) data and the 2024-25 school co-curricular activities (CCA) data to visualise the ratio of physical sports CCAs per 100 students in both primary and secondary schools. 

```{r}
# Loading school co-curricular activities data
school_sports_ccas <- 
  read.csv("CocurricularactivitiesCCAs.csv") %>% 
  group_by(school_name,school_section) %>% 
  summarise(
    total = n(),
    physical_sports_count = sum(cca_grouping_desc == "PHYSICAL SPORTS")) %>% 
    select(school_name,school_section, physical_sports_count) %>% 
    filter(school_section != "JUNIOR COLLEGE")
  


#2023 OSO school obesity data combined with 2024-25 school physical activity cca data
#gives spatial awareness to physical activity cca by joining to obesity data set by school name
obesity_joined_school_sports_data <- 
  obesity_data_sf_2023_OSO %>% 
  left_join(school_sports_ccas, by = join_by(SCHOOL == school_name)) %>%
  mutate(sports_cca_per_100 = physical_sports_count / ENROLMENT * 100) #ratio of sports cca per 100 students
```

# Visualising physical sports participation by school level

###########################PRIMARY SCHOOL####################################

```{r}
# Primary school sports participation

#NEED TO AVERAGE VALUES FOR MISSING VALUES
pri_cca_data <- obesity_joined_school_sports_data %>% filter(LEVEL == "PRIMARY", !is.na(sports_cca_per_100))

pri_cca_map <- ura_planning_areas_map+  
  tm_shape(pri_cca_data) +
  tm_symbols(
    size = "sports_cca_per_100",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    shape = 21,  # circle
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(values = "brewer.rd_yl_gn"),
    fill.legend = tm_legend(
      title = "Sports CCAs per 100 Students",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
    )
  ) + 
  tm_layout(
    legend.position = c("BOTTOM", "RIGHT"),
    legend.title.size = 1.2,
    main.title = "Ratio of Sports CCAs per 100 Students by Primary School"
  )

pri_cca_map 



# Spatial descriptive summary measures of physical sports participation by primary school

# Mean centre, standard deviation distance, standard deviation ellipse
pri_cca_mean <- calc_mnc(id = 1, points = st_coordinates(pri_cca_data))
pri_cca_median <- calc_mdc(id = 2, points = st_coordinates(pri_cca_data))
pri_cca_mean_median <- rbind(pri_cca_mean$LOCATIONS, pri_cca_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
pri_cca_mean_median$id = c("Spatial mean", "Spatial median")

pri_cca_sdd <- calc_sdd(id = 3, points = st_coordinates(pri_cca_data))
pri_cca_sdd_line <- pri_cca_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_cca_sde <- calc_sde(id = 4, points = st_coordinates(pri_cca_data))
pri_cca_sde_line <- pri_cca_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_cca_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = pri_cca_data$sports_cca_per_100,
  points = st_coordinates(pri_cca_data))
pri_cca_sde_weighted_line <- pri_cca_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

pri_cca_map_sde <- ura_planning_areas_map +
  tm_shape(pri_cca_data, name = "Ratio of Sports CCAs per 100 Students by School") +
  tm_symbols(
    size = "sports_cca_per_100",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Sports CCAs per 100 Students",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  tm_shape(pri_cca_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(pri_cca_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(pri_cca_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(pri_cca_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Ratio of Sports CCAs per 100 Students by Primary School")

pri_cca_map_sde

```

####################################SECONDARY SCHOOL####################################

```{r}
sec_cca_data <- obesity_joined_school_sports_data %>% filter(LEVEL == "SECONDARY", !is.na(sports_cca_per_100))

sec_cca_map <- ura_planning_areas_map +  
  tm_shape(sec_cca_data) +
  tm_symbols(
    size = "sports_cca_per_100",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    shape = 22,  #square
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(values = "brewer.rd_yl_gn"),
    fill.legend = tm_legend(
      title = "Sports CCAs per 100 Students",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
    )
  ) + 
  tm_layout(
    legend.position = c("BOTTOM", "RIGHT"),
    legend.title.size = 1.2,
    main.title = "Ratio of Sports CCAs per 100 Students by Secondary School"
  )


sec_cca_map



# Mean centre, standard deviation distance, standard deviation ellipse
sec_cca_mean <- calc_mnc(id = 1, points = st_coordinates(sec_cca_data))
sec_cca_median <- calc_mdc(id = 2, points = st_coordinates(sec_cca_data))
sec_cca_mean_median <- rbind(sec_cca_mean$LOCATIONS, sec_cca_median$LOCATIONS) %>% 
  data.frame() %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs)
sec_cca_mean_median$id = c("Spatial mean", "Spatial median")

sec_cca_sdd <- calc_sdd(id = 3, points = st_coordinates(sec_cca_data))
sec_cca_sdd_line <- sec_cca_sdd$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_cca_sde <- calc_sde(id = 4, points = st_coordinates(sec_cca_data))
sec_cca_sde_line <- sec_cca_sde$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_cca_sde_weighted <- calc_sde(
  id = 5, calccentre = T, weighted = T, weights = sec_cca_data$sports_cca_per_100,
  points = st_coordinates(sec_cca_data))
sec_cca_sde_weighted_line <- sec_cca_sde_weighted$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs = sg_crs) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

sec_cca_map_sde <- ura_planning_areas_map +
  tm_shape(sec_cca_data, name = "Ratio of Sports CCAs per 100 Students by School") +
  tm_symbols(
    shape = 22,
    size = "sports_cca_per_100",
    size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide(),
    fill = "sports_cca_per_100", 
    fill.scale = tm_scale_continuous(values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Sports CCAs per 100 Students",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) + 
  tm_shape(sec_cca_sdd_line, name = "SDD") + tm_lines(lwd = 2, lty = 1) + 
  tm_shape(sec_cca_sde_line, name = "SDE") + tm_lines(lwd = 2, lty = 3) +
  tm_shape(sec_cca_sde_weighted_line, name = "weighted SDE") + 
  tm_lines(col = "red", lwd = 2, lty = 3) +
  tm_shape(sec_cca_mean_median, name = "Summary measures") + tm_dots(
    size = 0.5, shape = "id", fill = "white", 
    shape.legend = tm_legend("Summary measures")
    ) + 
  tm_add_legend(type = "lines", labels = "SDD", lty = 1) +
  tm_add_legend(type = "lines", labels = "SDE", lty = 3) +
  tm_add_legend(type = "lines", labels = "weighted SDE", lty = 3, col = "red") +
  tm_title("Ratio of Sports CCAs per 100 Students by Secondary School")

sec_cca_map_sde


###################################################################

```

# Aggregating sports cca ratio counts by URA planning area

```{r}
cca_planning_avg <- st_join(ura_planning_areas, obesity_joined_school_sports_data) %>% 
  group_by(PLN_AREA_N, LEVEL) %>% 
  summarise(
    RATIO_AGG = mean(sports_cca_per_100, na.rm = T),
    geometry = st_union(geometry)
    ) %>%
  filter(!is.na(RATIO_AGG))

cca_planning_avg_map <- ura_planning_areas_map + cca_planning_avg %>% 
  tm_shape(name = "Sports CCA Ratio per 100") +
  tm_polygons(
    fill = "RATIO_AGG",
    fill.scale = tm_scale(values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Sports CCAs per 100 Students",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Average Sports CCAs Ratio by URA planning areas") +
  tm_facets_wrap(by = "LEVEL")

print(cca_planning_avg_map)







#raw cca avg for COMBINED schools
combined_cca_avg <- st_join(ura_planning_areas, obesity_joined_school_sports_data) %>% 
  group_by(PLN_AREA_N) %>% 
  summarise(
    RATIO_AGG = mean(sports_cca_per_100, na.rm = T),
    geometry = st_union(geometry)
    ) %>%
  filter(!is.na(RATIO_AGG))

```

# BIVARIATE CHLOROPLETH MAPS FOR CCA RATIO VS PREVALENCE

```{r}

library(classInt);library(dplyr)

#Select the two variables for the bivariate map
obesity_sports_data <- obesity_joined_school_sports_data %>% 
  filter(!is.na(sports_cca_per_100), !is.na(PREVALENCE)) %>%
  select(SCHOOL,LEVEL, sports_cca_per_100, PREVALENCE, geometry)

#primary school

pri_obesity_sports <- obesity_sports_data %>%
  filter(LEVEL == "PRIMARY") %>%
  select(-LEVEL)

pri_cca_prevalence <- st_join(ura_planning_areas, pri_obesity_sports) %>% 
  group_by(PLN_AREA_N) %>% 
  summarise(
    avg_cca_ratio = mean(sports_cca_per_100, na.rm = T),
    avg_prevalence = mean(PREVALENCE, na.rm = T),
  ) %>%
  filter(!is.na(avg_cca_ratio), !is.na(avg_prevalence))


#Classify both variables into 3 categories (low, medium, high)
pri_bc <- pri_cca_prevalence %>%
  mutate(
    sports_cat = cut(avg_cca_ratio, 
                     breaks = classIntervals(avg_cca_ratio, 3, style = "quantile")$brks, 
                     labels = c("Low", "Medium", "High"), 
                     include.lowest = TRUE),
    obesity_cat = cut(avg_prevalence, 
                      breaks = classIntervals(avg_prevalence, 3, style = "quantile")$brks, 
                      labels = c("Low", "Medium", "High"), 
                      include.lowest = TRUE)
  )

#Combine the two classifications into one
pri_bc <- pri_bc %>%
  mutate(bivariate_class = paste(sports_cat, obesity_cat, sep = "_"))


bivariate_colors <- c(
  "Low_Low" = "#e8e8e8", "Medium_Low" = "#b4d3e4", "High_Low" = "#8bb5d1",
  "Low_Medium" = "#e4b4b4", "Medium_Medium" = "#b4b4b4", "High_Medium" = "#8b8b8b",
  "Low_High" = "#d18b8b", "Medium_High" = "#b58b8b", "High_High" = "#8b8b8b"
)


pri_bivariate_map <-ura_planning_areas_map +
  tm_shape(pri_bc) +
  tm_polygons(fill = "bivariate_class", palette = bivariate_colors, title = "Bivariate Categories (CCA - Obesity)") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_title("Bivariate Map of Sports CCAs and Obesity Prevalence by URA Planning Areas (Primary Schools)")


pri_bivariate_map




# Secondary school bivariate map


sec_obesity_sports <- obesity_sports_data %>%
  filter(LEVEL == "SECONDARY") %>%
  select(-LEVEL)

sec_cca_prevalence <- st_join(ura_planning_areas, sec_obesity_sports) %>% 
  group_by(PLN_AREA_N) %>% 
  summarise(
    avg_cca_ratio = mean(sports_cca_per_100, na.rm = T),
    avg_prevalence = mean(PREVALENCE, na.rm = T),
  ) %>%
  filter(!is.na(avg_cca_ratio), !is.na(avg_prevalence))


#Classify both variables into 3 categories (low, medium, high)
sec_bc <- sec_cca_prevalence %>%
  mutate(
    sports_cat = cut(avg_cca_ratio, 
                     breaks = classIntervals(avg_cca_ratio, 3, style = "quantile")$brks, 
                     labels = c("Low", "Medium", "High"), 
                     include.lowest = TRUE),
    obesity_cat = cut(avg_prevalence, 
                      breaks = classIntervals(avg_prevalence, 3, style = "quantile")$brks, 
                      labels = c("Low", "Medium", "High"), 
                      include.lowest = TRUE)
  )

#Combine the two classifications into one
sec_bc <- sec_bc %>%
  mutate(bivariate_class = paste(sports_cat, obesity_cat, sep = "_"))


bivariate_colors <- c(
  "Low_Low" = "#e8e8e8", "Medium_Low" = "#b4d3e4", "High_Low" = "#8bb5d1",
  "Low_Medium" = "#e4b4b4", "Medium_Medium" = "#b4b4b4", "High_Medium" = "#8b8b8b",
  "Low_High" = "#d18b8b", "Medium_High" = "#b58b8b", "High_High" = "#8b8b8b"
)


sec_bivariate_map <-ura_planning_areas_map +
  tm_shape(sec_bc) +
  tm_polygons(fill = "bivariate_class", palette = bivariate_colors, title = "Bivariate Categories (CCA - Obesity)") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_title("Bivariate Map of Sports CCAs and Obesity Prevalence by URA Planning Areas (Secondary Schools)")


sec_bivariate_map
```


# Sports Fields and Facilities

```{r}
# Loading sports facilities data
sports_facilities_file <- "SportSGSportFacilities.geojson"
sports_facilities_sf <- st_read(sports_facilities_file) %>% 
  st_transform(crs = sg_crs)

st_make_valid(sports_facilities_sf) # Check for invalid geometries
st_as_sf(st_make_valid(sports_facilities_sf)) # Fix invalid geometries


sport_sg_facility_map <- ura_planning_areas_map + tm_shape(sports_facilities_sf) +
  tm_polygons(size = 0.3,
          fill = "slateblue4",
          fill.legend = tm_legend(title = "Sports Fields"))

sport_sg_facility_map



# Loading Sports fields data
fields_shp <- 
  st_read("SportsFieldsSGSHP/PLAYSG.shp") %>% 
  st_make_valid() %>%
  st_as_sf() %>%
  st_transform(crs = sg_crs) 

sports_fields_map <- ura_planning_areas_map + tm_shape(fields_shp) +
    tm_dots(size = 0.3,
          fill = "slateblue2",
          fill.legend = tm_legend(title = "Sports Fields"),
          fill_alpha = 0.5)


sports_fields_map



#DATA SET OF CHOICE
# Loading dual use facilities data

dual_use_facilities <- 
  st_read("DualUseSchoolsSportsFacilitiesSHP/DUS_School_Sports_Facilities.shp") %>% 
  st_make_valid() %>%
  st_as_sf() %>%
  st_transform(crs = sg_crs) 

dus_facility_map <- ura_planning_areas_map + tm_shape(dual_use_facilities) +
    tm_dots(size = 0.3,
          fill = "navy",
          fill.legend = tm_legend(title = "Dual Use School Sports Facilities"),
          fill_alpha = 0.5)

dus_facility_map




#displaying schools and sports facilities on the same map

#primary schools
pri_dus_map <- ura_planning_areas_map + 
  tm_shape(pri_cca_data) +
  tm_symbols(
    size = 0.3,
    fill = "red"
  ) +
  tm_shape(dual_use_facilities) +
  tm_dots(
    size = 0.3,
    fill = "blue"
  ) +
  tm_layout(
    legend.position = c("BOTTOM", "RIGHT"),
    legend.title.size = 1.2,
    main.title = "Primary Schools and Dual Use Sports Facilities"
  )

pri_dus_map


#seconadry schools
sec_dus_map <- ura_planning_areas_map + 
  tm_shape(sec_cca_data) +
  tm_symbols(
    size = 0.3,
    fill = "red"
  ) +
  tm_shape(dual_use_facilities) +
  tm_dots(
    size = 0.3,
    fill = "blue"
  ) +
  tm_layout(
    legend.position = c("BOTTOM", "RIGHT"),
    legend.title.size = 1.2,
    main.title = "Secondary Schools and Dual Use Sports Facilities"
  )

sec_dus_map

```

# K Nearest Neighbours

# NEAREST SPORTS FACILITY TO SCHOOL

## 1 neighbour

```{r}

# Calculate distances
distances <- st_distance(obesity_joined_school_sports_data, dual_use_facilities)

# For each school, find the k nearest facilities
k <- 3  # Change this to your desired k value
nearest_facilities <- list()

for (i in 1:nrow(obesity_joined_school_sports_data)) {
  # Get distances from this school to all facilities
  school_distances <- distances[i,]
  
  # Find indices of k nearest facilities (or fewer if there aren't k facilities)
  nearest_indices <- order(school_distances)[1:min(k, length(school_distances))]
  
  # Store these indices
  nearest_facilities[[i]] <- nearest_indices
}

# Create a data frame with the results
results <- data.frame(
  school_id = obesity_joined_school_sports_data$PT_ID,
  school_name = obesity_joined_school_sports_data$SCHOOL,
  level = obesity_joined_school_sports_data$LEVEL,
  geometry = obesity_joined_school_sports_data$geometry,
  stringsAsFactors = FALSE
)

# Add the nearest facilities for each school
for (j in 1:k) {
  facility_col <- paste0("nearest_facility_", j)
  distance_col <- paste0("distance_", j)
  
  results[[facility_col]] <- sapply(1:nrow(obesity_joined_school_sports_data),
  function(i) {
    if (length(nearest_facilities[[i]]) >= j) {
      # Return the ID of the jth nearest facility using the OBJECTID column
      dual_use_facilities$OBJECTID[nearest_facilities[[i]][j]]
    } else {
      NA
    }
  })
  
  results[[distance_col]] <- sapply(1:nrow(obesity_joined_school_sports_data),
  function(i) {
    if (length(nearest_facilities[[i]]) >= j) {
      # Return the distance to the jth nearest facility in meters
      as.numeric(distances[i, nearest_facilities[[i]][j]])
    } else {
      NA
    }
  })
}

pri_knn_3 <- results %>% filter(level == "PRIMARY") 
sec_knn_3 <- results %>% filter(level == "SECONDARY")



pri_nearest_dus <- st_as_sf(pri_knn_3, crs = sg_crs)
sec_nearest_dus <- st_as_sf(sec_knn_3, crs = sg_crs)


#AVERAGE DISTANCE TO THE NEAREST SPORTS FACILITY FROM SCHOOLS

#Heat map of nearest sports facility to schools - PRIMARY SCHOOLS
pri_dus_avg_distance <- st_join(ura_planning_areas, pri_nearest_dus) %>% 
  group_by(PLN_AREA_N, level) %>% 
  summarise(
    avg_distance = mean(distance_1, na.rm = T),
    geometry = st_union(geometry)
    ) %>%
  filter(!is.na(avg_distance))

pri_dus_avg_distance_map <- ura_planning_areas_map + pri_dus_avg_distance %>% 
  tm_shape(name = "Average Distance to Nearest DUS Sports Facility") +
  tm_polygons(
    fill = "avg_distance",
    fill.scale = tm_scale(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Average Distance in Metres",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Average Distance to Nearest DUS Sports Facility from Primary Schools")

print(pri_dus_avg_distance_map)




#Heat map of nearest sports facility to schools - SECONDARY SCHOOLS 
sec_dus_avg_distance <- st_join(ura_planning_areas, sec_nearest_dus) %>% 
  group_by(PLN_AREA_N, level) %>% 
  summarise(
    avg_distance = mean(distance_1, na.rm = T),
    geometry = st_union(geometry)
    ) %>%
  filter(!is.na(avg_distance))

sec_dus_avg_distance_map <- ura_planning_areas_map + sec_dus_avg_distance %>% 
  tm_shape(name = "Average Distance to Nearest DUS Sports Facility") +
  tm_polygons(
    fill = "avg_distance",
    fill.scale = tm_scale(values = "-brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "Average Distance in Metres",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("Average Distance to Nearest DUS Sports Facility from Secondary Schools")

print(sec_dus_avg_distance_map)


```

# Facility Density by URA Planning Area

```{r}

facility_by_area <- st_join(ura_planning_areas, dual_use_facilities) %>% 
  group_by(PLN_AREA_N) %>% 
  filter(!is.na(OBJECTID)) %>%
  select(PLN_AREA_N, OBJECTID) 

facility_by_area <- facility_by_area %>%
  group_by(PLN_AREA_N) %>%
  summarise(facility_count = n())

facility_by_area$area_km2 <- as.numeric(st_area(facility_by_area)) / 1e6

facility_by_area$facility_density <- facility_by_area$facility_count / facility_by_area$area_km2


facility_density_map <- ura_planning_areas_map + facility_by_area %>% 
  tm_shape(name = "DUS Sports Facility Density by URA Planning Area") +
  tm_polygons(
    fill = "facility_density",
    fill.scale = tm_scale(values = "brewer.rd_yl_gn"), 
    fill.legend = tm_legend(
      title = "DUS Facilities per km",
      orientation = "landscape", 
      position = c("BOTTOM", "LEFT"),
      item.height = 0.5, item.width = 2.1
      )
    ) +
  tm_title("DUS Sports Facility Density by URA Planning Area")


facility_density_map

```

# Modelling obesity in Singapore schools

## Joining data variables

```{r}
# Selecting relevant columns
census_merge_data <- census_merge %>% 
  st_drop_geometry() %>% 
  select(PLN_AREA_N, Income, Education)

nearest_outlet_distance_data <- nearest_outlet_distance %>% 
  st_drop_geometry()

num_food_outlets_data <- rbind(
  select(st_drop_geometry(obesity_pri_food), PT_ID, num_food_outlets),
  select(st_drop_geometry(obesity_sec_food), PT_ID, num_food_outlets)
  )

school_sports_ccas_data <- school_sports_ccas %>%
  select(-school_section)

nearest_dus_data <- 
  rbind(
    select(st_drop_geometry(pri_nearest_dus), school_id, distance_1),
    select(st_drop_geometry(sec_nearest_dus), school_id, distance_1)
    ) %>% 
  rename(distance_to_nearest_dus = distance_1)

# Joining data frames
obesity_overall_data <-
  obesity_data_sf_2023_OSO %>% 
  st_join(ura_planning_areas) %>% 
  left_join(census_merge_data, by = join_by(PLN_AREA_N)) %>%
  left_join(school_rankings, by = join_by(SCHOOL == School)) %>% 
  left_join(nearest_outlet_distance_data, by = join_by(SCHOOL)) %>% 
  left_join(num_food_outlets_data, by = join_by(PT_ID)) %>% 
  left_join(school_sports_ccas_data, by = join_by(SCHOOL == school_name)) %>%
  mutate(sports_cca_per_100 = physical_sports_count / ENROLMENT * 100) %>%
  select(-physical_sports_count) %>% 
  left_join(nearest_dus_data, by = join_by(PT_ID == school_id)) %>% 
  st_drop_geometry() %>% 
  mutate(across(
    c(LEVEL, type_code, nature_code, session_code, sap_ind, autonomous_ind, 
      gifted_ind, ip_ind, special, is_popular_sch, outlet_type), 
    as.factor
    )) 

obesity_overall_data %>% 
  filter(LEVEL == "PRIMARY") %>% 
  summary()

obesity_overall_data %>% 
  filter(LEVEL == "SECONDARY") %>% 
  summary()

# Converting categorical data to dummy variables

obesity_model_data <- obesity_overall_data %>% 
  mutate(
    # government schools = 1, non-govt (govt-aided, independent, specialsed) = 0
    is_govt_sch = ifelse(type_code == "GOVERNMENT SCHOOL", 1, 0),
    # co-ed schools = 1, non-coed (boys or girls schools) = 0
    is_coed_sch = ifelse(nature_code == "CO-ED SCHOOL", 1, 0),
    # single session schools = 1, non-single sess (double or full day) = 0
    is_single_sess = ifelse(session_code == "SINGLE SESSION", 1, 0),
    # SAP schools = 1, non-SAP = 0
    is_sap_sch = ifelse(sap_ind == "Yes", 1, 0),
    # autonomous schools = 1, non-autonomous schools = 0
    is_autonomous_sch = ifelse(autonomous_ind == "Yes", 1, 0),
    # gifted schools = 1, non-gifted schools = 0
    is_gifted_sch = ifelse(gifted_ind == "Yes", 1, 0),
    # integrated programme schools = 1, non-IP schools = 0
    is_ip_sch = ifelse(ip_ind == "Yes", 1, 0),
    # special schools = 1, non-special schools = 0
    is_special_sch = ifelse(special == "Yes", 1, 0),
    # popular schools = 1, non-popular schools = 0
    is_popular_sch = ifelse(is_popular_sch == "Y", 1, 0)
    )

# separating into primary and secondary schools
obesity_model_pri <- obesity_model_data %>% 
  filter(LEVEL == "PRIMARY") %>% 
  select(PREVALENCE, where(is.numeric), -c(PT_ID, GROUP_ID, YEAR)) %>% 
  # no integrated programme and too few autonomous primary schools
  select(-is_ip_sch, -is_autonomous_sch)

obesity_model_sec <- obesity_model_data %>% 
  filter(LEVEL == "SECONDARY") %>% 
  select(PREVALENCE, where(is.numeric), -c(PT_ID, GROUP_ID, YEAR))
```

## Correlation analysis

We must ensure that the predictors are correlated with the dependent variable and uncorrelated with each other (there is no multicollinearity). Correlation among predictors cause problems as interpretations become difficult as everything changes when one predictor changes. In addition collinearity makes inferences less reliable (power of tests are reduced).

```{r}
# for primary schools
rmat_pri <- cor(obesity_model_pri, use = "pairwise.complete.obs")
corrplot.mixed(rmat_pri, upper = "pie", number.cex=0.75, tl.cex = 0.75,
               upper.col = COL2("PiYG"), lower.col = COL2("PiYG"))

# for secondary schoools
rmat_sec <- cor(obesity_model_sec, use = "pairwise.complete.obs") 
corrplot.mixed(rmat_sec, upper = "pie", number.cex=0.75, tl.cex = 0.75,
               upper.col = COL2("PiYG"), lower.col = COL2("PiYG"))
```

We used pairwise complete observations as the sports_cca_per_100 had many missing rows.

**For Primary Schools** Based on the sign of the correlation coefficient values, we can observe a positive linear relationship between overweight prevalence (PREVALENCE) and the

-   Number of food outlets within a 1km radius of school (num_food_outlets),
-   Government school status (is_govt_sch),
-   Co-ed school status (is_coed_sch),
-   Single session school status (is_single_sess).

Furthermore, there is a negative linear relationship between overweight prevalence (PREVALENCE) and the

-   HMSP score (HMSP_Score),
-   Sports CCAs per 100 students (sports_cca_per_100),
-   Distance to nearest dual-use sports facility (distance_to_nearest_dus),
-   Gross Monthly Income from Work by Planning Area (Income),
-   Years of Education by Planning Area (Education),
-   Popular schools by Phase 2C subscription (is_popular_sch),
-   Distance from school to the nearest external food option (distance_to_nearest_food_option),
-   SAP school status (is_sap_sch),
-   GEP school status (is_gifted_sch),
-   Special school status: SAP or GEP (is_special_sch).

**For Secondary Schools** Based on the sign of the correlation coefficient values, we can observe a positive linear relationship between overweight prevalence (PREVALENCE) and the

-   Government school status (is_govt_sch),
-   Co-ed school status (is_coed_sch),
-   Single session school status (is_single_sess).

Furthermore, there is a negative linear relationship between overweight prevalence (PREVALENCE) and the

-   HMSP score (HMSP_Score),
-   Sports CCAs per 100 students (sports_cca_per_100),
-   Distance to nearest dual-use sports facility (distance_to_nearest_dus),
-   Gross Monthly Income from Work by Planning Area (Income),
-   Years of Education by Planning Area (Education),
-   Popular schools by cutoff score (is_popular_sch),
-   Distance from school to the nearest external food option (distance_to_nearest_food_option),
-   Number of food outlets within a 1km radius of school (num_food_outlets),
-   Single session school status (is_single_sess),
-   SAP school status (is_sap_sch),
-   Autonomous school status (is_autonomous_sch),
-   GEP school status (is_gifted_sch),
-   IP school status (is_ip_sch),
-   Special school status: SAP, GEP, Autonomous or IP (is_special_sch).

## Model Creation (LM, BSS, Foward/Backward Selection)

```{r}
obesity_model_analysis_pri <- obesity_model_pri %>%
  select(-COUNT, -ENROLMENT,-is_sap_sch,-is_gifted_sch)
lm.all_pri<- lm(PREVALENCE~.,data=obesity_model_analysis_pri)
summary(lm.all_pri)
obesity_model_analysis_sec <- obesity_model_sec %>%
  select(-COUNT, -ENROLMENT,-is_sap_sch,-is_gifted_sch,-is_ip_sch,-is_autonomous_sch)
lm.all_sec<- lm(PREVALENCE~.,data=obesity_model_analysis_sec)
summary(lm.all_sec)
```

For the primary school variables, the intercept is the most statistically significant with p-value close to 0. Other statistically significant variables include is_coed_sch (p-value \<0.01), is_popular_sch (p-value\<0.05), while moderately significant variables are is_govt_sch, is_special_sch and num_food_outlets (p-value\<0.1) For the secondary school variables, is_special_sch is the most statistically significant with p-value close to 0. Other statistically significant variables include the intercept (p-value\<0.05) and is_popular_sch (p-value\<0.1).

```{r}
#BSS for Pri
library(leaps)
regfitbss.all_pri <- regsubsets(PREVALENCE~., obesity_model_analysis_pri) # run the Best Subset Selection
regbss.summary_pri <- summary(regfitbss.all_pri)
res_bss_pri <- which.max(regbss.summary_pri$adjr2)
res_bss_pri
coef(regfitbss.all_pri, res_bss_pri)
c_bss_pri <- which.min(regbss.summary_pri$bic)
c_bss_pri
coef(regfitbss.all_pri, c_bss_pri)
#BSS for Sec
regfitbss.all_sec <- regsubsets(PREVALENCE~., obesity_model_analysis_sec) # run the Best Subset Selection
regbss.summary_sec <- summary(regfitbss.all_sec)
res_bss_sec <- which.max(regbss.summary_sec$adjr2)
res_bss_sec
coef(regfitbss.all_sec, res_bss_sec)
c_bss_sec <- which.min(regbss.summary_sec$bic)
c_bss_sec
coef(regfitbss.all_sec, c_bss_sec)

#Forward Selection for Pri
regfitfwd.all_pri <- regsubsets(PREVALENCE~., obesity_model_analysis_pri,method="forward") # run the Forward Selection
regfwd.summary_pri <- summary(regfitfwd.all_pri)
res_fwd_pri <- which.max(regfwd.summary_pri$adjr2)
res_fwd_pri
coef(regfitfwd.all_pri, res_fwd_pri)
c_fwd_pri <- which.min(regfwd.summary_pri$bic)
c_fwd_pri
coef(regfitfwd.all_pri, c_fwd_pri)
#Forward Selection for Sec
regfitfwd.all_sec <- regsubsets(PREVALENCE~., obesity_model_analysis_sec,method="forward") # run the Forward Selection
regfwd.summary_sec <- summary(regfitfwd.all_sec)
res_fwd_sec <- which.max(regfwd.summary_sec$adjr2)
res_fwd_sec
coef(regfitfwd.all_sec, res_fwd_sec)
c_fwd_sec <- which.min(regfwd.summary_sec$bic)
c_fwd_sec
coef(regfitfwd.all_sec, c_fwd_sec)

#Backward Selection for Pri
regfitbwd.all_pri <- regsubsets(PREVALENCE~., obesity_model_analysis_pri,method="backward") # run the Backward Selection
regbwd.summary_pri <- summary(regfitbwd.all_pri)
res_bwd_pri <- which.max(regbwd.summary_pri$adjr2)
res_bwd_pri
coef(regfitbwd.all_pri, res_bwd_pri)
c_bwd_pri <- which.min(regbwd.summary_pri$bic)
c_bwd_pri
coef(regfitbwd.all_pri, c_bwd_pri)
#Backward Selection for Sec
regfitbwd.all_sec <- regsubsets(PREVALENCE~., obesity_model_analysis_sec,method="backward") # run the Backward Selection
regbwd.summary_sec <- summary(regfitbwd.all_sec)
res_bwd_sec <- which.max(regbwd.summary_sec$adjr2)
res_bwd_sec
coef(regfitbwd.all_sec, res_bwd_sec)
c_bwd_sec <- which.min(regbwd.summary_sec$bic)
c_bwd_sec
coef(regfitbwd.all_sec, c_bwd_sec)
ols_pri<-lm(PREVALENCE~Education+is_popular_sch+is_govt_sch+is_coed_sch,data=obesity_model_analysis_pri)
summary(ols_pri)
ols_sec<-lm(PREVALENCE~Education+is_special_sch,data=obesity_model_analysis_sec)
summary(ols_sec)
```

Based on the three types of model selection, We observe that BSS and Backward selection recommend the same model for primary school variables with 4 terms based on smallest BIC criterion. This model was chosen because it has the fewest number of variables, facilitating simplicity in interpretation.

The best linear model for obesity prevalence in primary schools is \
PREVALENCE = 0.215854168 - 0.010527116 * Education - 0.008987562 * is_popular_sch + 0.014435592 * is_govt_sch + 0.016975461 * is_coed_sch 

This is corroborated by the variable analysis where is_coed_sch, is_popular_sch and is_govt_sch are recognised as statistically significant variables. Interestingly, the variable for Education is also included in the best linear model despite not being statistically significant in the earlier analysis.

We also observe all 3 approaches recommend the same model for secondary school variables with 2 terms based on smallest BIC criterion.

The best linear model for obesity prevalence in secondary schools is \
PREVALENCE = 0.247219424 - 0.009114996 * Education - 0.043620675 * is_special_sch 

This is corroborated by the variable analysis where Education and is_special_sch are the two most statistically significant variables.

## Residual checks for OLS model

### Residual analysis for Primary Schools

```{r}
# Generating residuals for OLS model
res_pri = residuals(ols_pri)

# Plotting residuals on independent variables used in OLS model
plot(obesity_model_analysis_pri$Education, res_pri, 
     main = "Residual Plot", 
     ylab = "Residuals",
     xlab = "Education")
cor(obesity_model_analysis_pri$Education, res_pri)

plot(obesity_model_analysis_pri$is_popular_sch, res_pri, 
     main = "Residual Plot", 
     ylab = "Residuals",
     xlab = "is_special_sch")
cor(obesity_model_analysis_pri$is_popular_sch, res_pri)

plot(obesity_model_analysis_pri$is_govt_sch, res_pri, 
     main = "Residual Plot", 
     ylab = "Residuals",
     xlab = "is_special_sch")
cor(obesity_model_analysis_pri$is_govt_sch, res_pri)

plot(obesity_model_analysis_pri$is_coed_sch, res_pri, 
     main = "Residual Plot", 
     ylab = "Residuals",
     xlab = "is_special_sch")
cor(obesity_model_analysis_pri$is_coed_sch, res_pri)

length(ols_pri)
summary(ols_pri)
dim(obesity_model_analysis_pri)
```

### Residual analysis for Secondary Schools

```{r}
# Generating residuals for OLS model
res_sec = residuals(ols_sec)

# Plotting residuals on independent variables used in OLS model
plot(obesity_model_analysis_sec$Education, res_sec, 
     main = "Residual Plot", 
     ylab = "Residuals",
     xlab = "Education")
cor(obesity_model_analysis_sec$Education, res_sec)

plot(obesity_model_analysis_sec$is_special_sch, res_sec, 
     main = "Residual Plot", 
     ylab = "Residuals",
     xlab = "is_special_sch")
cor(obesity_model_analysis_sec$is_special_sch, res_sec)

length(ols_sec)
summary(ols_sec)
dim(obesity_model_analysis_sec)
```

## Generating Spatial Models and Spatial Tests

### Spatial Models (Primary Schools)

```{r}
# Generating 4 spatial models for consideration: SAR, SLM, SEM, SARAR
pri_sar <- spautolm(formula = PREVALENCE ~ 1, data = obesity_model_analysis_pri, 
                    listw = soi_lw_pri)
summary(pri_sar) # Lambda: 0.34009, p-value: 7.2031e-06 

pri_slm <- lagsarlm(
  formula = PREVALENCE ~ Education + is_popular_sch + is_govt_sch + is_coed_sch, 
  data = obesity_model_analysis_pri, listw = soi_lw_pri)
summary(pri_slm) # Rho: 0.14194, p-value: 0.04966

pri_sem <- errorsarlm(
  formula = PREVALENCE ~ Education + is_popular_sch + is_govt_sch + is_coed_sch,
  data = obesity_model_analysis_pri, listw = soi_lw_pri)
summary(pri_sem) # Lambda: 0.11517, p-value: 0.17252

pri_sarar <- sacsarlm(
  formula = PREVALENCE ~ Education + is_popular_sch + is_govt_sch + is_coed_sch,
  data = obesity_model_analysis_pri, listw = soi_lw_pri)
summary(pri_sarar) 
# Rho: 0.29499, p-value: 0.11725; Lambda: -0.21663, p-value: 0.16879
```

### Spatial Tests for model comparison (Primary Schools)

```{r}
# Lagrange Multiplier tests

lm.LMtests(ols_pri, listw = soi_lw_pri, test = "all") 
# RSerr = 1.8749, df = 1, p-value = 0.1709
# RSlag = 3.8848, df = 1, p-value = 0.04873
# adjRSerr = 0.4243, df = 1, p-value = 0.5148
# adjRSlag = 2.4342, df = 1, p-value = 0.1187
# SARMA = 4.3091, df = 2, p-value = 0.116

# Likelihood Ratio tests

LR.Sarlm(pri_sem, pri_sarar) # p-value = 0.07437
LR.Sarlm(pri_slm, pri_sarar) # p-value = 0.275
LR.Sarlm(pri_slm, pri_sar) # p-value < 2.2e-16
LR.Sarlm(pri_sem, pri_sar) # p-value < 2.2e-16

LR1.Sarlm(pri_slm) # p-value = 0.04966
LR1.Sarlm(pri_sem) # p-value = 0.1725

# Comparing AIC 
AIC(pri_slm) # -921.7105
AIC(pri_sem) # -919.7185
AIC(ols_pri) # -919.8576

# Comparing BIC
BIC(pri_slm) # -899.321
BIC(pri_sem) # -897.329
BIC(ols_pri) # -900.6666
```

Using the LR tests, we see the tests with SARAR as alternative hypothesis that the large p-values for both tests mean SLM and SEM are more suitable than SARAR. The extremely small p-value for the test between SAR and SLM indicates the alternative SLM is more suitable than SAR. For the tests with OLS as the null, we see that when SLM is the alternative the p-value is small (0.01429) which is sufficient evidence at 5% significance level that the alternative SLM is more suitable. On the other hand, when SEM is the alternative, the p-value is larger (0.05814), so there is insufficient evidence at 5% significance level to reject OLS. This is further supported by the LM test results, where adjRSlag has a much smaller p-value of 0.06073 as compared to adjRSerr p-value of 0.6956. This suggests the SLM is more suitable than SEM, though the significance of SLM relative to OLS may be difficult to determine at 5% significance level from this test alone. Finally, considering the AIC and BIC values, we see pri_slm has marginally smaller AIC and BIC values than both SEM and OLS. This brings us to our conclusion that SLM is the most suitable spatial model for modelling obesity prevalence in primary schools.

### Spatial Models (Secondary Schools)

```{r}
# Generating 4 spatial models for consideration: SAR, SLM, SEM, SARAR
sec_sar <- spautolm(formula = PREVALENCE ~ 1, data = obesity_model_analysis_sec, 
                    listw = soi_lw_sec)
summary(sec_sar) # Lambda: 0.27535, p-value: 0.0018312 

sec_slm <- lagsarlm(
  formula = PREVALENCE ~ Education + is_popular_sch + is_govt_sch + is_coed_sch, 
  data = obesity_model_analysis_sec, listw = soi_lw_sec)
summary(sec_slm) # Rho: 0.13841, p-value: 0.082856

sec_sem <- errorsarlm(
  formula = PREVALENCE ~ Education + is_popular_sch + is_govt_sch + is_coed_sch,
  data = obesity_model_analysis_sec, listw = soi_lw_sec)
summary(sec_sem) # Lambda: 0.1602, p-value: 0.083578

sec_sarar <- sacsarlm(
  formula = PREVALENCE ~ Education + is_popular_sch + is_govt_sch + is_coed_sch,
  data = obesity_model_analysis_sec, listw = soi_lw_sec)
summary(sec_sarar) 
# Rho: 0.076366, p-value: 0.63869; Lambda: 0.084001, p-value: 0.65265
```

### Spatial Tests for model comparison (Secondary Schools)

```{r}
# Lagrange Multiplier tests

lm.LMtests(ols_sec, listw = soi_lw_sec, test = "all") 
# RSerr = 0.91006, df = 1, p-value = 0.3401
# RSlag = 2.2997, df = 1, p-value = 0.1294
# adjRSerr = 0.60109, df = 1, p-value = 0.4382
# adjRSlag = 1.9907, df = 1, p-value = 0.1583
# SARMA = 2.9008, df = 2, p-value = 0.2345

# Likelihood Ratio tests

LR.Sarlm(sec_sem, sec_sarar) # p-value = 0.6651
LR.Sarlm(sec_slm, sec_sarar) # p-value = 0.6771
LR.Sarlm(sec_slm, sec_sar) # p-value = 2.331e-15
LR.Sarlm(sec_sem, sec_sar) # p-value = 2.331e-15

LR1.Sarlm(sec_slm) # p-value = 0.08286
LR1.Sarlm(sec_sem) # p-value = 0.08358

# Comparing AIC 
AIC(sec_slm) # -600.2422
AIC(sec_sem) # -600.2281
AIC(ols_sec) # -601.9956

# Comparing BIC
BIC(sec_slm) # -579.4535
BIC(sec_sem) # -579.4535
BIC(ols_sec) # -590.1164
```

Using the LR tests, we see the tests with SARAR as alternative hypothesis that the large p-values for both tests mean SLM and SEM are more suitable than SARAR. The extremely small p-value for the test between SAR and SLM indicates the alternative SLM is more suitable than SAR. For the tests with OLS as the null, we see in both tests where SLM and SEM are alternatives that the p-value is much larger than 0.05, showing there is insufficient evidence at 5% significance level to reject the null OLS. This is further supported by the LM test results, where adjRSlag and adjRSerr p-values are very large. This suggests the OLS may be a more suitable model than the spatial models SLM and SEM. Finally, considering the AIC and BIC values, we see ols_sec has smaller AIC and BIC values than both SEM and OLS. This brings us to our conclusion that OLS is the most suitable for modelling obesity prevalence in primary schools, more so than any type of spatial model.

## Time series

```{r}
library(tidyverse)
library(imputeTS)  # For interpolation
library(fpp2)
library(gridExtra)
library(seasonal)
library(urca)
library(vars)
library(trend)
library(readxl)
library(dplyr)

oso_pri_hist = read_excel("oso_primary.xlsx")

oso_pri_hist_years = oso_pri_hist[, c(1,2, 18, 20, 22, 24, 26, 28)] %>%
  drop_na()
  

oso_pri_hist_long <- oso_pri_hist_years %>%
  pivot_longer(
    cols = matches("^\\d"),   # Select columns where names start with a number
    names_to = "year",
    values_to = "oso"
  ) %>%
  mutate(year = as.integer(paste0("20", str_extract(year, "^\\d{2}"))))  


oso_pri_hist_long %>%
  group_by(SCHOOL) %>%
  mutate(enrolment = na_interpolation(oso, option = "spline")) %>%
  ungroup()


oso_hist_int = oso_pri_hist_long %>%
  group_by(SCHOOL) %>%
  complete(year = full_seq(year, 1)) %>%  # Adds missing years with NA values
  mutate(oso = na_interpolation(oso, option = "spline"),
         oso_int = na_interpolation(oso, option = "spline")) %>%
  ungroup() 

# Showing general distribution of OSO in a histogram
oso_hist_int %>% 
  filter(year == 2023) %>% 
  ggplot(aes(x = oso_int)) +
  geom_histogram() # generally follows a normal distribution

#Get the Z-scores
# Compute z-scores before and after 2019
oso_hist_int = oso_hist_int %>%
  group_by(year) %>%  # Group by year
  mutate(
    mean_oso = mean(oso_int, na.rm = TRUE),
    sd_oso = sd(oso_int, na.rm = TRUE),
    z_score = (oso_int - mean_oso) / sd_oso
  ) %>%
  ungroup()

# Label the periods
oso_hist_int_2 <- oso_hist_int %>%
  mutate(period = ifelse(year <= 2019, "Before 2019", "After 2019"))

z_scores_summary <- oso_hist_int_2 %>%
  group_by(SCHOOL, period) %>%
  summarise(avg_z = mean(z_score, na.rm = TRUE)) %>%
  ungroup()

z_scores_change= z_scores_summary %>%
  pivot_wider(names_from = period, values_from = avg_z) %>%
  mutate(z_change =`After 2019` - `Before 2019`) %>%
  dplyr::select(c(1,4))


# Merge with spatial data
oso_hist_geom_full = obesity_data_sf_2023_OSO %>%
  inner_join(z_scores_change, by = "SCHOOL")

oso_hist_geom = oso_hist_geom_full %>%
  dplyr::select(c("PT_ID", "SCHOOL", "geometry", "z_change"))


# School obesity map plot (z-score change)
ura_regions_map +
  tm_shape(oso_hist_geom) +
  tm_symbols(size = 0.5, fill = "z_change")

# Schools which saw a 1 standard deviation increase in OSO levels
oso_hist_geom %>% 
  st_drop_geometry() %>% 
  filter(z_change >= 1) %>% 
  arrange(desc(z_change)) # 10 schools identified
  
# Schools which saw a 1 standard deviation decrease in OSO levels
oso_hist_geom %>% 
  st_drop_geometry() %>% 
  filter(z_change <= -1) %>% 
  arrange(desc(z_change)) # 10 schools identified
```

## Sphere of influence neighbours

```{r}
# by Primary schools
# obesity_pri_sch_jitter <- st_jitter(obesity_pri_sch) 
del_nb_pri_z <- tri2nb(st_coordinates(oso_hist_geom), 
                     row.names = oso_hist_geom$PT_ID) 
pri_map <- tm_shape(nb2lines(del_nb_pri_z, coords = st_geometry(obesity_pri_sch))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Primary schools
if (require(dbscan, quietly=TRUE)) {
  soi_nb_pri_z <- graph2nb(soi.graph(del_nb_pri_z, st_geometry(oso_hist_geom)))
} else {
  soi_nb = NULL
}
pri_map_z <- tm_shape(nb2lines(soi_nb_pri_z, coords = st_geometry(oso_hist_geom))) +
  tm_lines(lwd = 2) + pri_sch_map
```

### Spatial autocorrelation statistics

```{r}
# by Primary schools
(soi_lw_pri_z <- nb2listw(soi_nb_pri_z, style = "W"))
soi_lw_B_pri_z <- nb2listw(soi_nb_pri_z, style = "B")
moran.test(oso_hist_geom$z_change, soi_lw_pri_z)
moran.plot(oso_hist_geom$z_change, soi_lw_pri_z, main="Moran scatterplot")
geary.test(oso_hist_geom$z_change, soi_lw_pri_z)



```

Average number of links (Primary school): 3.094972\
p-value for Global Moran's I test (Primary school): 0.5712 p-value for Global Geary's C test (Primary school): 0.3879

Based on the Moran's I and Geary's C tests, there is no significant spatial autocorrelation in Z-score change between Primary Schools. This suggests that the incidence of obesity prevalence when normalized does not affect neighbouring schools before and after 2019.

#### Join count analysis

```{r}
# Setting an overweight cutoff point
Z_CHANGE_CUTOFF_INCREASE <- 1 #increase in obesity levels before and after 2019
Z_CHANGE_CUTOFF_DECREASE <- -1 #decrease in obesity levels before and after 2019

# by Primary school - with increased obesity levels
oso_hist_geom1 = oso_hist_geom %>%
  filter(z_change > 0)
oso_hist_geom1$z_change_high_increase <- 
  as.factor(oso_hist_geom1$z_change > Z_CHANGE_CUTOFF_INCREASE)
pri_map_z_increase <- ura_regions_map + oso_hist_geom1 %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "z_change_high_increase", 
    size = "z_change", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")


# by Primary school - with decreased obesity levels ##TO ASK PROF: here, the bigger circles are smaller decreases. should i change everything to positive values so that it shows up correctly?
oso_hist_geom2 = oso_hist_geom %>%
  filter(z_change < 0)
oso_hist_geom2$z_change_high_decrease <- 
  as.factor(oso_hist_geom2$z_change < Z_CHANGE_CUTOFF_DECREASE)
pri_map_z_decrease <- ura_regions_map + oso_hist_geom2 %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "z_change_high_decrease", 
    size = "z_change", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")


#example with absolute values
Z_CHANGE_CUTOFF_DECREASE_1 = 1

oso_hist_geom3 = oso_hist_geom %>%
  filter(z_change < 0) %>%
  mutate(z_change = abs(z_change))

oso_hist_geom3$z_change_high_decrease <- 
  as.factor(oso_hist_geom3$z_change > Z_CHANGE_CUTOFF_DECREASE_1)
pri_map_z_decrease1 <- ura_regions_map + oso_hist_geom3 %>% 
  tm_shape(name = "Overweight prevalence") + 
  tm_symbols(
    fill = "z_change_high_decrease", 
    size = "z_change", size.scale = tm_scale_continuous(values.scale = 0.8),
    size.legend = tm_legend_hide()) +
  tm_title("Join count analysis")


#Maps:
#pri_map_z_decrease
pri_map_z_decrease1
pri_map_z_increase
```

# Joint count considering only extremes

```{r}
#For increases
del_nb_pri_z_increase <- tri2nb(st_coordinates(oso_hist_geom1), 
                     row.names = oso_hist_geom1$PT_ID) 
pri_map_z_increase <- tm_shape(nb2lines(del_nb_pri_z_increase, coords = st_geometry(oso_hist_geom1))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Primary schools
if (require(dbscan, quietly=TRUE)) {
  soi_nb_pri_z_increase <- graph2nb(soi.graph(del_nb_pri_z_increase, st_geometry(oso_hist_geom1)))
} else {
  soi_nb = NULL
}
#pri_map_z <- tm_shape(nb2lines(soi_nb_pri_z, coords = st_geometry(oso_hist_geom))) +
#  tm_lines(lwd = 2) + pri_sch_map

# by Primary schools
(soi_lw_pri_z_increase <- nb2listw(soi_nb_pri_z_increase, style = "W"))
soi_lw_B_pri_z_increase <- nb2listw(soi_nb_pri_z_increase, style = "B")

#For decreases
del_nb_pri_z_decrease <- tri2nb(st_coordinates(oso_hist_geom3), 
                     row.names = oso_hist_geom3$PT_ID) 
pri_map_z_decrease <- tm_shape(nb2lines(del_nb_pri_z_decrease, coords = st_geometry(oso_hist_geom3))) +
  tm_lines(lwd = 2) + pri_sch_map

# by Primary schools
if (require(dbscan, quietly=TRUE)) {
  soi_nb_pri_z_decrease <- graph2nb(soi.graph(del_nb_pri_z_decrease, st_geometry(oso_hist_geom3)))
} else {
  soi_nb = NULL
}
#pri_map_z <- tm_shape(nb2lines(soi_nb_pri_z, coords = st_geometry(oso_hist_geom))) +
#  tm_lines(lwd = 2) + pri_sch_map

# by Primary schools
(soi_lw_pri_z_decrease <- nb2listw(soi_nb_pri_z_decrease, style = "W"))
soi_lw_B_pri_z_decrease <- nb2listw(soi_nb_pri_z_decrease, style = "B")


#Join count analysis
joincount.multi(oso_hist_geom3$z_change_high_decrease, 
                soi_lw_B_pri_z_decrease) 
joincount.multi(oso_hist_geom1$z_change_high_increase, 
                soi_lw_B_pri_z_increase) 

```

Based on the joint count analysis, there is no significant spatial clustering of any schools which had an extreme decrease in obesity (of \<1)

Based on the joint count analysis, there is no significant spatial clustering of any schools which had an extreme increase in obesity (of \>1)

```{r}
tmap_mode("view") #interactive map
tmap_mode("plot") #static map
# tmap_save(tmap_object, 'xx.jpg') # for saving maps as images
cols4all::c4a_gui() # colour scales
RColorBrewer::display.brewer.all()
```
