---
title: "Obesity in 2023"
author: "S F Liu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of Obesity in 2023

## EDA

Packages

```{r include=F}
pkgs = c("dplyr", "ggplot2", "tidyr", "readr", "crayon", "tibble", "knitr", 
         "gt", "sf")
invisible(lapply(pkgs, function(pkg) {
  suppressPackageStartupMessages(library(pkg, character.only = TRUE))
}))
rm(pkgs)
if (requireNamespace("rstudioapi", quietly = TRUE)) {
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}
```

Read and reclassify data

```{r}
# Read CSV file
oso23 = read_csv("oso23.csv", show_col_types = FALSE)
oso23 = oso23 |> drop_na(OSO_pct) 

# Re-classify
oso23[,c(3, 4, 8, 13:26)] = lapply(oso23[,c(3, 4, 8, 13:26)], as.factor)
```

### Boxplots

```{r}
# OSO rate by school level
ggplot(oso23, aes(x = LEVEL, y = OSO_pct, fill = LEVEL)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by School Level",
       x = "Level",
       y = "OSO Rate") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates to make it horizontal 

# OSO rate by Zone
ggplot(oso23, aes(x = ZONE, y = OSO_pct, fill = ZONE)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by Zone",
       x = "Zone",
       y = "OSO Rate") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates to make it horizontal 

# OSO rate by School Type
sub = oso23 |>
  filter(type_code %in% c("GOVERNMENT SCHOOL", "GOVERNMENT-AIDED SCH"))
ggplot(sub, aes(x = type_code, y = OSO_pct, fill = type_code)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by School Type",
       x = "School Type",
       y = "OSO Rate") +
  theme_minimal() +
  coord_flip()  # Flip the coordinates to make it horizontal 

# OSO rate by Gender Mix
ggplot(oso23, aes(x = nature_code, y = OSO_pct, fill = nature_code)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by Gender Mix",
       x = "Gender Mix",
       y = "OSO Rate") +
  theme_minimal() +
  scale_x_discrete(drop = TRUE) +
  coord_flip() # Flip the coordinates to make it horizontal 
  
# OSO rate by Specialised Programs
ggplot(oso23, aes(x = special, y = OSO_pct, fill = special)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by Specialised Programs",
       x = "Special Program",
       y = "OSO Rate") +
  theme_minimal() +
  scale_x_discrete(drop = TRUE) +
  coord_flip() # Flip the coordinates to make it horizontal 

# OSO rate by HMSP Tier
ggplot(oso23, aes(x = HMSP_Tier, y = OSO_pct, fill = HMSP_Tier)) +
  geom_boxplot(na.rm = T) +
  labs(title = "Boxplot of OSO23 rate by HMSP Tier",
       x = "HMSP Tier",
       y = "OSO Rate") +
  theme_minimal() +
  scale_x_discrete(drop = TRUE) +
  coord_flip() # Flip the coordinates to make it horizontal 
cor(oso23$OSO_pct, oso23$HMSP_Score)
```

### Hypothesis Tests

```{r}
t.test(OSO_pct ~ LEVEL, data = oso23, var.equal = F)
t.test(OSO_pct ~ type_code, data = sub, var.equal = F)
t.test(OSO_pct ~ special, data = oso23, var.equal = F)

sub = oso23 |>
  mutate(ZONE = case_when(
   ZONE %in% c("NORTH", "WEST") ~ "NW",
   ZONE %in% c("SOUTH", "EAST") ~ "SE",
  ))
t.test(OSO_pct ~ ZONE, data = sub, var.equal = F)

sub = oso23 |>
  mutate(nature_code = as.factor(ifelse(
   nature_code %in% c("BOYS' SCHOOL", "GIRLS' SCHOOL"), "SINGLE", nature_code
  )))
t.test(OSO_pct ~ nature_code, data = sub, var.equal = F)

sub = oso23 |>
  mutate(HMSP_Tier = case_when(
   HMSP_Tier %in% c("A", "B") ~ "AB",
   HMSP_Tier %in% c("C", "D", "E") ~ "CDE",
  ))
t.test(OSO_pct ~ HMSP_Tier, data = sub, var.equal = F)
```

### Observations

- Obesity rate of secondary schools seem to be somewhat higher than that of those in primary schools. Try to find some external empirical evidence to explain why this may be the case. Two sample t-test rejects equality of mean obesity rates.

- Obesity rates of schools in the West ad North have somewhat higher rates of obesity compared to the schools in the South and East. Two sample t-test rejects equality of mean obesity rates between NW and SE. This is something directly relevant to our course. You may explore this avenue further by augmenting with sub-zone and planning area analyses. Moran's I (local and global) can be useful here. Also identify regional clusters using Moran's I. HPB will be able to address their efforts directly in these schools.

- Government schools seem to have a higher prevalence of obesity. Try explain why this maybe the case. Two sample t-test rejects equality of mean obesity rates.

- Obesity rates seem to be lower in girl's schools. Interestingly the rates are higher in co-ed schools. t-test for no difference between the mean obesity rate of single gender schools versus co-ed schools were rejected. Research and see whether there is evidence of such outcomes in other countries as well. Also check whether there is anthropological evidence that may support this idea!

- Schools with specialised educational programs such as SAP, IP, and GEP. The obesity rate of those schools seem to be significantly lower compared to the rest of the schools. Two sample t-test rejects equality of mean obesity rates. There is likely a third unobserved variable that is causing this difference. We should try to find this variable. I feel a better comparison that can supersede this variable could be the "elite" status of the school. I suggest you to look for a way to assign this binary variable to the schools based on primary school enrollment and secondary school cut-off points. 

- The fact that schools with a tier A HMSP score has a lower obesity prevalence is an important finding. There is an outlier in this data. Find out what that school is. The schools with a tier B score level seems to be somewhat lower in obesity rates compared to the rest. Further, the rates of obesity in Tier A and B seem to be more consistent with less variability. I think HPB will want to know this to support the program. 

- However, the overall correlation between the HMSP score and the obesity rates seem to be low. Further, the t-test for no difference in the mean obesity rates of schools with high versus low tier scores was not rejected. We need to re-calibrate this numerical result to include only the schools with actual data (without mode imputation). I believe, Ian is currently working in this direction.

Other things to do:

- Identify the outliers in the boxplots.

- Add in the regional indicators from Census 2020.

- Construct a model.
