---
title: "HMSP_EDA"
author: "S F Liu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EDA

Load the libraries.

```{r}
library(sf); library(tmap); suppressMessages(library(spdep))
```

Import data.

```{r}
hmsp = read.csv("hmsp_tier_pri_sec.csv")
head(hmsp)
```

Give spatial awareness.

```{r}
sg = st_read("MySingapura.shp")
sg = st_union(sg)
sg_map = tm_shape(sg) + tm_borders()

(hmsp = st_as_sf(hmsp, coords = c("longitude", "latitude"), crs = 4326))
sg_map + 
  tm_shape(hmsp) + 
  tm_dots(size = 0.1) +
  tm_title("Primary and Secondary Schools, Singapore, 2022")
```

School cluster maps.

```{r}
# School zones
(sch_zones = 
  sg_map +
  tm_shape(hmsp) + 
  tm_dots(
    fill = "ZONE", 
    fill.scale = tm_scale(values = "poly.glasbey"), 
    size = 0.5,
    fill.legend = tm_legend(position = c("right", "bottom"))) +
  tm_title("School Zones, 2022") +
  tm_style("classic"))

# School clusters
sch_clusters = 
  sg_map +
  tm_shape(hmsp) + 
  tm_dots(fill = "CLUSTER") + 
  tm_facets_wrap(by = "ZONE", nrow = 1, ncol = 1)
tmap_animation(sch_clusters, width=800, delay=200)
tmap_animation(sch_clusters, filename="sch_clusters.gif", width=800, delay=200)
```

HMSP tier maps.

```{r}
tmap_mode("view")
hmsp_tier = 
  tm_shape(hmsp) + 
  tm_dots(
    fill = "TIER", 
    fill.scale = tm_scale(values = "powerbi.bloom"), 
    size = 0.5
    )
hmsp_tier
hmsp_score = 
  tm_shape(hmsp) + 
  tm_dots(
    fill = "TIER_SCORE", 
    fill.scale = tm_scale(values = "brewer.blues"), 
    size = 0.5)
hmsp_score
```

# Cluster based analysis (Social distance)

Visualisation of the the clusters

```{r}
tm_shape(hmsp) + tm_dots(fill = "CLUSTER", size = 0.5)
```

Group interaction weights matrix based on school clusters.

```{r}
(cluster_nb = read.gal("clusters.gal"))
plot(cluster_nb, st_geometry(hmsp))

```

Spatial autocorrelation measures.

```{r}
cluster_wlist = nb2listw(cluster_nb)
moran.test(hmsp$TIER_SCORE, cluster_wlist)
moran.plot(hmsp$TIER_SCORE, cluster_wlist)
geary.test(hmsp$TIER_SCORE, cluster_wlist)
```

Do not reject the null of no spatial autocorrelation at 5% level. There is insufficient evidence to suggest a significant spatial spillover effect within school clusters. (Higher order neighbours are not valid here because we used a Group Interaction scheme.) 

Average number of links: 10.88146
p-value for Moran's test: 0.1903

# Graph neighbours based analysis

1. Delauney triangulation neighbours
 
```{r}
hmsp = st_jitter(hmsp) # Pertrurb geometries to remove duplication
coords = st_coordinates(hmsp)
del_nb = tri2nb(coords, row.names = hmsp$PT_ID) 
plot(del_nb, st_geometry(hmsp))
# write.nb.gal(del_nb, "del.gal")
```

Spatial autocorrelation measures.

```{r}
del_wlist = nb2listw(del_nb)
moran.test(hmsp$TIER_SCORE, del_wlist)
moran.plot(hmsp$TIER_SCORE, del_wlist)
geary.test(hmsp$TIER_SCORE, del_wlist)
```

Average number of links: 5.896657 
p-value for Moran's test: 0.2324

2. Sphere of influence neighbours

```{r}
if (require(dbscan, quietly=TRUE)) {
  soi_nb = graph2nb(soi.graph(del_nb, coords))
} else soi_nb = NULL
plot(soi_nb, st_geometry(hmsp))
# write.nb.gal(soi_nb, "soi.gal")
```

Spatial autocorrelation measures.

```{r}
soi_wlist = nb2listw(soi_nb)
moran.test(hmsp$TIER_SCORE, soi_wlist)
moran.plot(hmsp$TIER_SCORE, soi_wlist)
geary.test(hmsp$TIER_SCORE, soi_wlist)
```

Average number of links: 2.613982 
p-value for Moran's test: 0.4771

# K nearest neighbour based analysis

1. One neighbour

```{r}
knn1_nb = knn2nb(knearneigh(coords, k = 1), row.names = hmsp$PT_ID) 
plot(knn1_nb, st_geometry(hmsp))
```

Spatial autocorrelation measures.

```{r}
knn1_wlist = nb2listw(knn1_nb)
moran.test(hmsp$TIER_SCORE, knn1_wlist)
moran.plot(hmsp$TIER_SCORE, knn1_wlist)
geary.test(hmsp$TIER_SCORE, knn1_wlist)
```

Average number of links: 1
p-value for Moran's test: 0.7719

2. Two neighbours

```{r}
knn2_nb = knn2nb(knearneigh(coords, k = 2), row.names = hmsp$PT_ID) 
plot(knn2_nb, st_geometry(hmsp))
```

Spatial autocorrelation measures.

```{r}
knn2_wlist = nb2listw(knn2_nb)
moran.test(hmsp$TIER_SCORE, knn2_wlist)
moran.plot(hmsp$TIER_SCORE, knn2_wlist)
geary.test(hmsp$TIER_SCORE, knn2_wlist)
```

Average number of links: 2
p-value for Moran's test: 0.6085

3. Three neighbours

```{r}
knn3_nb = knn2nb(knearneigh(coords, k = 3), row.names = hmsp$PT_ID) 
plot(knn3_nb, st_geometry(hmsp))
```

Spatial autocorrelation measures.

```{r}
knn3_wlist = nb2listw(knn3_nb)
moran.test(hmsp$TIER_SCORE, knn3_wlist)
moran.plot(hmsp$TIER_SCORE, knn3_wlist)
geary.test(hmsp$TIER_SCORE, knn3_wlist)
```

Average number of links: 3
p-value for Moran's test: 0.4829

To HPB: *Weak evidence* of spatial autocorrelation based on school clusters.