---
title: "Obesity in Primary Schools"
author: "S F Liu"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pkgs = c("dplyr", "ggplot2", "tidyr", "readr", "crayon", "tibble", "knitr", 
         "gt", "sf", "tmap")
invisible(lapply(pkgs, function(pkg) {
  suppressPackageStartupMessages(library(pkg, character.only = TRUE))
}))
rm(pkgs)
if (requireNamespace("rstudioapi", quietly = TRUE)) {
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}
```

# Load Data

```{r}
oso_primary = read_csv("oso_primary.csv", show_col_types = FALSE)
oso_primary = oso_primary |> drop_na(average) 

# Re-classify
oso_primary[,c(3, 4, 15, 17:30)] = 
  lapply(oso_primary[,c(3, 4, 15, 17:30)], as.factor)
```

# Density of OSO 2016 to 2023

```{r}
# Reshape data into long format
long_data = oso_primary |>
  pivot_longer(
    cols = c(OSO16_pct, OSO17_pct, OSO19_pct, OSO20_pct, OSO22_pct, OSO23_pct), 
    names_to = "Variable", 
    values_to = "Value"
    )

# Create density plot
ggplot(long_data, aes(x = Value, fill = Variable, color = Variable)) +
  geom_density(alpha = 0.4) +
  labs(title = "Density Plots of Obesity Rates: 2016 - 2023",
       x = "Obesity Rate",
       y = "Density") +
  theme_minimal()
```

- Density of 2022 lies above the other years due to the COVID effect.

- Average obesity rate is computed without this temporal anomaly.

```{r}
# Create new categorical column 'above_avg'
oso_primary = oso_primary |>
  mutate(
    average = rowMeans(across(c(OSO16_pct, OSO17_pct, OSO19_pct, OSO20_pct, OSO23_pct)), 
                       na.rm = TRUE),
    above_avg = as.factor(ifelse(OSO23_pct > average, "Above Average", "Below Average"))
  )

# Create boxplot comparing OSO23_pct for Above/Below Average groups
ggplot(oso_primary, aes(x = above_avg, y = OSO23_pct, fill = above_avg)) +
  geom_boxplot(alpha = 0.6, outlier.shape = 16, outlier.color = "red") +
  labs(title = "Boxplot of OSO23 Rate for Above and Below Average Groups",
       x = "Status",
       y = "OSO23_pct") +
  scale_fill_manual(values = c("red", "green")) +  # Custom colors
  theme_minimal() +
  coord_flip() # Flip the coordinates to make it horizontal 

t.test(OSO23_pct ~ above_avg, data = oso_primary, var.equal = F)
```

- There is significant evidence of a difference in the mean obesity rates of schools above and below average.

# Maps

Base Map
```{r}
sg = st_read("MySingapura.shp")
sg = st_union(sg)
sg_map = tm_shape(sg) + tm_borders()
oso_primary = st_as_sf(oso_primary, 
                       coords = c("longitude", "latitude"), crs = 4326)
```

## Schools with obesity rates above 16%

```{r}
sub = oso_primary |>
  filter(OSO23_pct > 0.15)
sub$SCHOOL
sg_map + 
  tm_shape(sub) + tm_dots(fill = "OSO23_pct", 
                          fill.legend = tm_legend(position = c("right", "bottom")),
                          fill.scale = tm_scale_intervals(style = "pretty")) +
  tm_title("Primary Schools with Obesity Rates Higher than 15% (2023)")
```


## Schools with above average obesity rates:

Below are the schools whose 2023 obesity rate is above the historic average recorded in their respective schools. This highlights schools with an obesity rate that is generally on the rise.

```{r}
sub = oso_primary |>
  filter(above_avg == "Above Average")
sg_map + 
  tm_shape(sub) + tm_dots(fill = "OSO23_pct", 
                          fill.legend = tm_legend(position = c("right", "bottom")),
                          fill.scale = tm_scale_intervals(style = "pretty")) +
  tm_title("Primary Schools with Rising Obesity Rates")
summary(oso_primary$average)
```

Upon removing the post COVID year, the average obesity rate between 2016 and 2023 is 11.6% (median is 11.7%)

# Obesity Rate in 2023 Compared to Other Aspects

1. Obesity and school type

```{r}
ggplot(oso_primary, aes(x = OSO23_pct, fill = type_code, color = type_code)) +
  geom_density(alpha = 0.4) +  # Adjust transparency for overlap
  labs(title = "Density Plot of Obesity Rate by School Type",
       x = "Obesity Rate 2023",
       y = "Density") +
  theme_minimal()
```

Government schools typically has a higher obesity rate.

2. Obesity and school type

```{r}
ggplot(oso_primary, aes(x = OSO23_pct, fill = zone_code, color = zone_code)) +
  geom_density(alpha = 0.4) +  # Adjust transparency for overlap
  labs(title = "Density Plot of Obesity Rate by Zone",
       x = "Obesity Rate 2023",
       y = "Density") +
  theme_minimal()
```

Obesity rate in the South is typically lower.

3. Obesity and gender mix

```{r}
ggplot(oso_primary, aes(x = OSO23_pct, fill = nature_code, color = nature_code)) +
  geom_density(alpha = 0.4) +  # Adjust transparency for overlap
  labs(title = "Density Plot of Obesity Rate by Gender Mix",
       x = "Obesity Rate 2023",
       y = "Density") +
  theme_minimal()
```

Co-ed schools typically have a higher rate of obesity.

4. Obesity and special program

```{r}
ggplot(oso_primary, aes(x = OSO23_pct, fill = special, color = special)) +
  geom_density(alpha = 0.4) +  # Adjust transparency for overlap
  labs(title = "Density Plot of Obesity Rate by Special Program Availability",
       x = "Obesity Rate 2023",
       y = "Density") +
  theme_minimal()
```

5. Obesity and popularity

```{r}
ggplot(oso_primary, aes(x = OSO23_pct, fill = Popular, color = Popular)) +
  geom_density(alpha = 0.4) +  # Adjust transparency for overlap
  labs(title = "Density Plot of Obesity Rate by Popularity",
       x = "Obesity Rate 2023",
       y = "Density") +
  theme_minimal()
```

A school is defined as popular if all the available spaces are filled by Phase 2C. Based on this definition, schools that are less popular seem to have a higher obesity prevalence than the other schools.